### 2024-07-31 paper/paper.tex
first commit

\documentclass[conference]{IEEEtran} 
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{array}
\usepackage{multirow}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\begin{document}

\title{An In-Depth Analysis of Student Prompts Using CharlieBot \\
{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
should not be used}
\thanks{Identify applicable funding agency here. If none, delete this.}
}

\author{\IEEEauthorblockN{1\textsuperscript{st} Rodrigo Prestes Machado}
\IEEEauthorblockA{\textit{dept. Informática} \\
\textit{Instituto Federal do Rio Grande do Sul}\\
Porto Alegre, Brazil \\
0000-0003-0428-6387}
\and
\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
\and
\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
\textit{name of organization (of Aff.)}\\
City, Country \\
email address or ORCID}
}

\maketitle

\begin{abstract}
This document is a model and instructions for \LaTeX.
This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
or Math in Paper Title or Abstract.
\end{abstract}

\begin{IEEEkeywords}
component, formatting, style, styling, insert
\end{IEEEkeywords}

\section{Introduction}

%What is the problem to be solved?

% Are there any existing solutions?

% Which is the best?

% What is its main limitation?

% What do you hope to achieve?

\section{Background}

\begin{table*}[htbp]
\caption{Categories and Descriptions}
    \begin{center}
    \begin{tabular}{|p{3cm}|p{3cm}|p{5cm}|}
    \hline
    \textbf{Critical Thinking} & \textbf{Category} & \textbf{Description} \\
    \hline
    \multirow{6}{3cm}{Tends to promote} & Debugging Help & Prompts that seek
    help to identify, fix errors, or understand the provided code snippet. \\
    & Conceptual Questions & Prompts that are more about understanding concepts
    or algorithms rather than specific code. \\
    & Student Correction & Prompts where the student corrects the bot. \\
    \hline
    \multirow{6}{3cm}{Tends not to promote} & Code Snippet & Prompts that ask
    for a specific part of the code, like a function or a segment. \\
    & Complete Solution & Prompts that request an entire solution or a complete
    code snippet. \\
    & Multiple Question Exercise & Prompts where the user wants to solve a
    multiple-question exercise. \\
    \hline
    \multirow{3}{3cm}{Neutral} & Language Change & Prompts that request a change
    of idiom. \\
    & Uncategorized & Prompts that do not fit into any of the above categories. \\
    \hline
    \end{tabular}
    \label{tab:categories}
    \end{center}
\end{table*}

\section{Method}

Initially, we identified four initial categories were identified for classifying
student messages with the bot, as outlined in the article "Coding with AI: How
Are Tools Like ChatGPT Being Used by Students in Foundational Programming
Courses." These initial categories are as follows: (1) Debugging Help: Messages
that seek assistance in identifying and fixing errors or understanding the
provided code snippet; (2) Code Snippet: Messages requesting a specific part
of the code, such as a function or a segment; (3) Conceptual Questions: Messages
focused on understanding concepts or algorithms rather than specific code;
and (4) Complete Solution: Messages requesting an entire solution or a complete
code snippet.
\\
Upon further exploration of the data, it was determined that four additional
categories could be added, resulting in a total of eight categories: (1)
Multiple Question Exercise: Messages where the user seeks to solve a multiple-question
exercise; (2) Student Correction: Messages where the student corrects the bot;
(3) Language Change: Messages requesting a change of language; and (4)
Uncategorized: Messages that do not fit into any of the aforementioned
categories.
\\
Subsequently, these eight categories were input into Claude.ai, accompanied by
detailed instructions on the desired structure of the input and the expected
format of the output. For each conversation in the CharlieBot log, a .csv file
was downloaded, and Claude was tasked with classifying the user messages into
one of the eight categories.

The classifications made by Claude were then manually reviewed. In cases where
there was disagreement with the classification, manual adjustments were made,
and the accuracy of Claude's classification was recorded. Each conversation
with CharlieBot was saved into a .csv file containing the original conversation,
along with two additional columns: "Classification" to denote the category and
"AI" to indicate Claude's accuracy.

As new categorizations were completed, an example was submitted to Claude's
knowledge base to enhance accuracy. Initially, Claude began with an example
representing 1\% of the total memory of its knowledge base. As classifications
were progressively completed, more examples were provided to Claude to increase
the percentage of its knowledge base occupied by relevant examples.

Upon completion of all categorizations, all resulting .csv files were consolidated
for final counts and analysis using the Pandas library. This methodology
facilitated an efficient categorization of student messages, thereby enabling
a comprehensive analysis of how tools such as ChatGPT are utilized in foundational
programming courses.

\section{Result and Discussion}

Bellow is the distribution of the messages in the categories:

\centerline{\includegraphics[scale=0.6]{figures/figure1.png}}

Approximately 64\% of the messages pertain to categories associated with
critical thinking, corroborating with the findings of
\cite{10.1007/978-3-031-64299-9_20}. In contrast, around 28\% of the messages
indicate a preference among students for ready-made answers.

\section{Conclusion}

\bibliographystyle{IEEEtran}
\bibliography{references}

\vspace{12pt}
\color{red}
IEEE conference templates contain guidance text for composing and formatting conference papers. Please ensure that all template text is removed from your conference paper prior to submission to the conference. Failure to remove the template text from your paper may result in your paper not being published.

\end{document}

### 2024-07-31 .devcontainer/postCreate.sh

@@ -1,0 +1,5 @@
+#!/bin/sh
+
+pip install pandas matplotlib textblob
+sudo tlmgr install times courier cite algorithms xcolor cm-super texlive-fonts-recommended
+sudo tlmgr install multirow

### 2024-08-01 paper/paper.tex
basic updates

@@
 
-\title{An In-Depth Analysis of Student Prompts Using CharlieBot \\
+\title{An In-Depth Analysis of Student Prompts With CharlieBot \\
 {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 
+According \cite{LO2024105100} further research about the use of AI tools in
+education is needed to understand how these tools can be used to improve
+students' learning experience. In particular, the study suggested that future
+research on students' study habits.
+
 % Are there any existing solutions?
@@
 \section{Background}
+
+The paper \cite{10343037} examined students from different demographic groups
+using AI tools like ChatGPT, comparing perceptions and impacts
+among these groups. The results showed that AI tools had a positive impact on
+academic performance, especially for students with less prior experience.
 
@@
 
-\section{Result and Discussion}
+\section{Results and Discussion}
 
-Bellow is the distribution of the messages in the categories:
-
-\centerline{\includegraphics[scale=0.6]{figures/figure1.png}}
+The graph \ref{fig:graph1} shows the distribution of the messages in the
+categories:
+\begin{figure}[h!]
+    \centering
+    \includegraphics[scale=0.7]{figures/figure1.png}
+    \caption{Classification of messages.}
+    \label{fig:graph1}
+\end{figure}
 
@@
 
-\vspace{12pt}
-\color{red}
-IEEE conference templates contain guidance text for composing and formatting conference papers.
-Please ensure that all template text is removed from your conference paper prior to submission to the conference.
-Failure to remove the template text from your paper may result in your paper not being published.
-
 \end{document}

### 2024-08-01 paper/paper.tex
update

@@
 \begin{table*}[htbp]
-\caption{Categories and Descriptions}
+\caption{Categories of Analysis}
     \begin{center}
-    \begin{tabular}{|p{3cm}|p{3cm}|p{5cm}|}
+    \begin{tabular}{|p{3cm}|p{5cm}|p{3cm}|}
     \hline
-    \textbf{Critical Thinking} & \textbf{Category} & \textbf{Description} \\
+    \textbf{Category} & \textbf{Description} & \textbf{Critical Thinking} \\
     \hline
-    \multirow{6}{3cm}{Tends to promote} & Debugging Help & Prompts that seek
-    help to identify, fix errors, or understand the provided code snippet. \\
-    & Conceptual Questions & Prompts that are more about understanding concepts
-    or algorithms rather than specific code. \\
-    & Student Correction & Prompts where the student corrects the bot. \\
+    Debugging Help & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \multirow{6}{3cm}{Tends to promote.} \\
+    Conceptual Questions & Prompts that are more about understanding concepts or algorithms rather than specific code. & \\
+    Student Correction & Prompts where the student corrects the bot. & \\
     \hline
-    \multirow{6}{3cm}{Tends not to promote} & Code Snippet & Prompts that ask
-    for a specific part of the code, like a function or a segment. \\
-    & Complete Solution & Prompts that request an entire solution or a complete
-    code snippet. \\
-    & Multiple Question Exercise & Prompts where the user wants to solve a
-    multiple-question exercise. \\
+    Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. & \multirow{6}{3cm}{Tends not to promote.} \\
+    Complete Solution & Prompts that request an entire solution or a complete code snippet. & \\
+    Multiple Question & Prompts where the user wants to solve a multiple-question exercise. & \\
     \hline
-    \multirow{3}{3cm}{Neutral} & Language Change & Prompts that request a change
-    of idiom. \\
-    & Uncategorized & Prompts that do not fit into any of the above categories. \\
+    Language Change & Prompts that request a change of idiom to convey a message more effectively. & \multirow{3}{3cm}{Neutral.} \\
+    Uncategorized & Prompts that do not fit into any of the above categories. & \\
     \hline
@@
 
+The table \ref{tab:categories} shows the categories of analysis used in this
+study. The categories were based on the work of
+\cite{10.1007/978-3-031-64299-9_20}, with the addition of four new categories.
+Also, we group the categories to understand the impact on critical thinking.
+
 \section{Method}
 
-Initially, we identified four initial categories were identified for classifying
-student messages with the bot, as outlined in the article "Coding with AI: How
-Are Tools Like ChatGPT Being Used by Students in Foundational Programming
-Courses." These initial categories are as follows: (1) Debugging Help: Messages
-that seek assistance in identifying and fixing errors or understanding the
-provided code snippet; (2) Code Snippet: Messages requesting a specific part
-of the code, such as a function or a segment; (3) Conceptual Questions: Messages
-focused on understanding concepts or algorithms rather than specific code;
-and (4) Complete Solution: Messages requesting an entire solution or a complete
-code snippet.
+CharlieBot is a ChatGPT-based bot that provides assistance to students in
+Java programming courses in UC3M. CharlieBot logs all conversations with
+students, which were used for this study. The goal was to categorize student
+messages to understand how students interact with the bot and identify patterns
+in their study habits.
 \\
-Upon further exploration of the data, it was determined that four additional
-categories could be added, resulting in a total of eight categories: (1)
-Multiple Question Exercise: Messages where the user seeks to solve a multiple-question
-exercise; (2) Student Correction: Messages where the student corrects the bot;
-(3) Language Change: Messages requesting a change of language; and (4)
-Uncategorized: Messages that do not fit into any of the aforementioned
-categories.
+The categories outlined in \cite{10.1007/978-3-031-64299-9_20} served as the
+foundation for classifying student messages. However, it was identified that
+some categories were missing, necessitating the addition of four new categories.
+Table \ref{tab:categories} presents these categories along with their
+descriptions.
 \\
@@
 detailed instructions on the desired structure of the input and the expected
-format of the output. For each conversation in the CharlieBot log, a .csv file
+format of the output. For each conversation with the CharlieBot log, a .csv file
 was downloaded, and Claude was tasked with classifying the user messages into

### 2024-08-05 paper/paper.tex
first draft

@@
 
-\title{An In-Depth Analysis of Student Prompts With CharlieBot \\
+\title{Analysis of Programming Students' Prompts to Identify Patterns of
+Critical Thinking in Interactions with Generative AI \\
 {\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
@@
 
-According \cite{LO2024105100} further research about the use of AI tools in
+According \cite{Kwan24} further research about the use of AI tools in
 education is needed to understand how these tools can be used to improve
@@
 research on students' study habits.
+
+Problems
+
+The literature suggests a concern regarding the impact of ChatGPT on students'
+critical thinking abilities. Vargas \cite{Vargas2023} indicated that
+ChatGPT use might lead to an overreliance on the tool, potentially diminishing
+students' critical thinking skills. Similarly, Cai \cite{cai23} identified
+overdependence and reduced intellectual engagement as significant drawbacks of
+ChatGPT-supported learning. Chan \cite{chan23} also noted that this
+reliance could result in a decrease in critical thinking, as students might make
+decisions based solely on the information provided by ChatGPT. Collectively,
+these studies highlight a common theme of reduced critical engagement and
+overreliance on ChatGPT in educational contexts.
 
@@
 
-The paper \cite{10343037} examined students from different demographic groups
+The paper \cite{Kwan24} examined students from different demographic groups
 using AI tools like ChatGPT, comparing perceptions and impacts
@@
 academic performance, especially for students with less prior experience.
+
+\subsection{Critical Thinking}
+
+Critical thinking is a fundamental skill for students to develop, as it enables
+them to analyze, evaluate, and synthesize information. This skill is essential
+for problem-solving and decision-making, making it a crucial aspect of
+educational success, specifically in programming courses. However, the use of AI
+tools like ChatGPT has raised concerns about the potential impact on students'
+critical thinking abilities \cite{Vargas2023} \cite{cai23} \cite{chan23}.
+
+In the other hand, the study of \cite{zhang24} suggested that incorporating
+ChatGPT could stimulate students' critical thinking, probably because they
+had to assess the correctness of ChatGPT's answers.
+
+Regarding self-regulation, \cite{cai23} noted that ChatGPT allowed students to
+regulate their learning pace.
+
+\cite{wu24} found that ChatGPT-supported learning not only promoted students'
+self-regulation but also increased their self-efficacy.
+
+Finally, some students asserted that the use of ChatGPT enhanced the depth of
+their thinking and understanding \cite{chan23}.
+
+\subsection{CharlieBot}
+
+CharlieBot is a ChatGPT-based bot that provides assistance to students in Java
+programming courses at UC3M. The bot uses RAG(retrieval augmented generation) to
+generate responses to student queries, providing explanations, code snippets,
+and debugging help. The goal of CharlieBot is to support students in their
+learning process by providing personalized assistance and feedback.
 
@@
 study. The categories were based on the work of
-\cite{10.1007/978-3-031-64299-9_20}, with the addition of four new categories.
+\cite{Ghimire24} with the addition of four new categories.
 Also, we group the categories to understand the impact on critical thinking.
@@
 \\
-The categories outlined in \cite{10.1007/978-3-031-64299-9_20} served as the
-foundation for classifying student messages. However, it was identified that
-some categories were missing, necessitating the addition of four new categories.
-Table \ref{tab:categories} presents these categories along with their
-descriptions.
+The categories outlined in \cite{Ghimire24} served as the
+foundation for classifying student messages (Debugging Help, Conceptual
+Questions, Code Snippet and Complete Solution). However, it was identified that
+some categories were missing, necessitating the addition of new categories
+(Student Correction, Multiple Question, Language Change and Uncategorized). Table
+\ref{tab:categories} presents these categories along with their descriptions.
 \\
@@
 critical thinking, corroborating with the findings of
-\cite{10.1007/978-3-031-64299-9_20}. In contrast, around 28\% of the messages
+\cite{Ghimire24}. In contrast, around 28\% of the messages
 indicate a preference among students for ready-made answers.

### 2024-08-12 paper/paper.tex
update

@@
 \and
-\IEEEauthorblockN{2\textsuperscript{nd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
+\IEEEauthorblockN{2\textsuperscript{nd} Carlos Alario Hoyos}
+\IEEEauthorblockA{\textit{Dep. Ingeniería Telemática} \\
+\textit{Universidad Carlos III de Madrid}\\
+Madrid, Spain \\
+0000-0002-3082-0814}
 \and
-\IEEEauthorblockN{3\textsuperscript{rd} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{4\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{5\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
-\and
-\IEEEauthorblockN{6\textsuperscript{th} Given Name Surname}
-\IEEEauthorblockA{\textit{dept. name of organization (of Aff.)} \\
-\textit{name of organization (of Aff.)}\\
-City, Country \\
-email address or ORCID}
+\IEEEauthorblockN{3\textsuperscript{rd} Carlos Delgado Kloos}
+\IEEEauthorblockA{\textit{Dep. Ingeniería Telemática} \\
+\textit{Universidad Carlos III de Madrid}\\
+Madrid, Spain \\
+0000-0003-4093-3705}
 }
@@
 
-According \cite{Kwan24} further research about the use of AI tools in
-education is needed to understand how these tools can be used to improve
-students' learning experience. In particular, the study suggested that future
-research on students' study habits.
+Recent advancements in Generative Artificial Intelligence (GenAI) have opened
+up new possibilities in education. In programming courses, students can utilize
+GenAI tools to improve their understanding, receiving assistance, feedback, and
+detailed explanations.
 
-Problems
+The study by Chan and Hu et al. \cite{chan23} revealed that both undergraduate
+and postgraduate students exhibit positive attitudes toward the use of GenAI
+in teaching and learning, noting that it enhances the depth of their thinking
+and understanding. Additionally, a systematic review by Lo, Hew, and Jong
+\cite{Lo24} demonstrated that students could effectively learn from ChatGPT,
+leading to improved understanding and academic achievement. Cai et al.
+\cite{cai23} further observed that ChatGPT enables students to regulate their
+learning pace, promoting self-regulation.
 
-The literature suggests a concern regarding the impact of ChatGPT on students'
-critical thinking abilities. Vargas \cite{Vargas2023} indicated that
-ChatGPT use might lead to an overreliance on the tool, potentially diminishing
-students' critical thinking skills. Similarly, Cai \cite{cai23} identified
-overdependence and reduced intellectual engagement as significant drawbacks of
-ChatGPT-supported learning. Chan \cite{chan23} also noted that this
-reliance could result in a decrease in critical thinking, as students might make
-decisions based solely on the information provided by ChatGPT. Collectively,
-these studies highlight a common theme of reduced critical engagement and
-overreliance on ChatGPT in educational contexts.
+In the other hand, researchers have concerns about the impact of these
+tools on students. The systematic review of Vargas-Murillo et al.
+\cite{Murillo23} indicated that ChatGPT use might lead to an overreliance on the
+tool. Chan et al. \cite{chan23} also noted that this reliance could result in a
+decrease in critical thinking, as students might make decisions based solely on
+the information provided by ChatGPT.
+
+The student's confidence is warranted, as shown by Puryear and Sprint
+\cite{Puryear22}, who demonstrated that GitHub Copilot can generate solutions
+for student assignments with accuracy rates ranging from 68\% to 95\%. However,
+this raises concerns that students might rely too heavily on GenAI tools,
+potentially neglecting a deeper understanding of the underlying concepts.
+Cai et al. \cite{cai23} further identified overdependence and reduced
+intellectual engagement as significant drawbacks of using ChatGPT in learning.
+
+Regardless of teachers' preferences or beliefs, preliminary surveys by Dickey,
+Bejarano, and Garg \cite{Dickey24} revealed that at least 54.5\% of students are
+already using Generative AI (GenAI) for homework. This highlights the need to
+understand how students are interacting with these tools and how they can be
+used to enhance learning. Furthermore, a review by Lo, Hew, and Jong \cite{Lo24}
+emphasized the necessity for extended studies and objective measures to gather
+more robust evidence on the use of GenAI tools in education.
 
@@
 
+To deal with this issue, some researchers have proposed different pedagogical
+strategies. Denny et al. \cite{Denny24} introduced the concept of \textit{Prompt
+Problems}, where students solve programming exercises by formulating natural
+language prompts. Prasad and Aamod \cite{Prasad24} proposed a self-regulated
+learning framework using GenAI to solve programming problems. Lauren and Watta
+\cite{Lauren23} explored the integration of generative AI with evidence-based
+learning strategies in computer science and engineering courses.
+
 % Which is the best?
+
+Effective strategies are crucial to determinate the best approaches and training
+educators to use GenAI tools effectively can be time-consuming and
+resource-intensive. Therefore, understanding how students interact with these
+tools is essential. By analyzing these interaction patterns, we can assess the
+impact of GenAI tools on students' learning outcomes.
+This knowledge can then be used to refine the design of GenAI tools, ensuring
+they better support pedagogical strategies and enhance overall learning
+experiences.
 
@@
 
+It is important to consider how GenAI tools affect students across different
+demographic groups, knowledge areas, and levels of programming experience.
+Understanding these variations can provide insights into the impact of GenAI
+tools on their learning outcomes, helping to tailor educational approaches that
+address the specific needs of diverse student populations.
+
 % What do you hope to achieve?
+
+This study aims to analyze the interactions between students and CharlieBot,
+an ChatGPT-based bot developed by UC3M that leverages Retrieval-Augmented
+Generation (RAG) to enhance its contextual understanding of Java programming.
 
@@
 
-The paper \cite{Kwan24} examined students from different demographic groups
+The paper \cite{Lo24} From the cognitive perspective, students were able to
+learn from ChatGPT, which increased their understanding and achievement.
+However, concerns were raised that the growing use of AI tools might lead to
+a decline in critical thinking among students. Teachers and stakeholders should
+continue to investigate pedagogical approaches that leverage ChatGPT to enhance
+students' understanding and critical thinking. For example, teachers can
+instruct students to fact-check and validate information produced by ChatGPT.
+
+examined students from different demographic groups
 using AI tools like ChatGPT, comparing perceptions and impacts
@@
 self-regulation but also increased their self-efficacy.
-
-Finally, some students asserted that the use of ChatGPT enhanced the depth of
-their thinking and understanding \cite{chan23}.
 

### 2024-08-19 paper/paper.tex
updates

@@
 \IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
+% The preceding line is only needed to identify funding in the first footnote.
+% If that is unneeded, please comment it out.
 \usepackage{cite}
@@
 
-\title{Analysis of Programming Students' Prompts to Identify Patterns of
-Critical Thinking in Interactions with Generative AI \\
-{\footnotesize \textsuperscript{*}Note: Sub-titles are not captured in Xplore and
-should not be used}
-\thanks{Identify applicable funding agency here. If none, delete this.}
+\title{Analysis of Programming Students' Prompts to Identify Interactions
+Patterns with Generative AI \\
+{\footnotesize \textsuperscript{}}
+\thanks{The authors would like to thank the Federal Institute of Education,
+Science, and Technology of Rio Grande do Sul (IFRS) for the partial financial
+support provided for the execution of this research.}
 }
@@
 \cite{Lo24} demonstrated that students could effectively learn from ChatGPT,
-leading to improved understanding and academic achievement. Cai et al.
-\cite{cai23} further observed that ChatGPT enables students to regulate their
-learning pace, promoting self-regulation.
+leading to improved understanding and academic achievement. Besides of that,
+it was also observed  that ChatGPT enables students to regulate their learning
+pace, promoting self-regulation \cite{Baha24} \cite{cai23}.
 
@@
 learning framework using GenAI to solve programming problems. Lauren and Watta
-\cite{Lauren23} explored the integration of generative AI with evidence-based
+\cite{Lauren23} explored the integration of GenAI with evidence-based
 learning strategies in computer science and engineering courses.
@@
 
-Effective strategies are crucial to determinate the best approaches and training
-educators to use GenAI tools effectively can be time-consuming and
-resource-intensive. Therefore, understanding how students interact with these
-tools is essential. By analyzing these interaction patterns, we can assess the
-impact of GenAI tools on students' learning outcomes.
-This knowledge can then be used to refine the design of GenAI tools, ensuring
-they better support pedagogical strategies and enhance overall learning
-experiences.
+The study of effective educational strategies in use of GenAI tools in education
+context are essential for determining the best training for educators. However,
+as these tools gain popularity among students \cite{Dickey24} the urgency to
+equip educators with best practices may lag behind their rapid adoption.
+Therefore, analyzing students' interaction patterns with GenAI tools is
+essential for understanding how these tools are being used and how educators can
+leverage them to enhance learning. In addition, these interaction patterns can
+create new opportunities to redesign GenAI tools to better support pedagogical
+strategies and improve overall learning.
 
@@
 
-It is important to consider how GenAI tools affect students across different
-demographic groups, knowledge areas, and levels of programming experience.
-Understanding these variations can provide insights into the impact of GenAI
-tools on their learning outcomes, helping to tailor educational approaches that
-address the specific needs of diverse student populations.
+Nevertheless, it is important to consider how GenAI tools affect students across
+different demographic groups, knowledge areas, cultural backgrounds \cite{catalan21}
+\cite{neo22} and students' prior experience. Thus, this works is limited to a
+specific group of students and a specific GenAI tool. The results may not be
+generalizable to other populations or tools.
 
@@
 tools like ChatGPT has raised concerns about the potential impact on students'
-critical thinking abilities \cite{Vargas2023} \cite{cai23} \cite{chan23}.
+critical thinking abilities \cite{Murillo23} \cite{cai23} \cite{chan23}.
 

### 2024-08-26 paper/paper.tex
update

@@
 
-\subsection{CharlieBot}
+\subsection{The CharlieBot}
 
-CharlieBot is a ChatGPT-based bot that provides assistance to students in Java
-programming courses at UC3M. The bot uses RAG(retrieval augmented generation) to
-generate responses to student queries, providing explanations, code snippets,
-and debugging help. The goal of CharlieBot is to support students in their
-learning process by providing personalized assistance and feedback.
+CharlieBot is a ChatGPT 3.5 based bot that provides assistance to students in a
+Java programming courses at UC3M. The bot uses the content of UC3M Java
+programming course at EdX to supply a RAG(Retrieval-Augmented Generation) system
+and thus try to respond more accurately to students' questions \cite{Sun24}.
+The main goal of CharlieBot is support students in their learning process by
+providing personalized assistance and feedback.
 
@@
 
-The table \ref{tab:categories} shows the categories of analysis used in this
-study. The categories were based on the work of
-\cite{Ghimire24} with the addition of four new categories.
-Also, we group the categories to understand the impact on critical thinking.
-
 \section{Method}
 
-CharlieBot is a ChatGPT-based bot that provides assistance to students in
-Java programming courses in UC3M. CharlieBot logs all conversations with
-students, which were used for this study. The goal was to categorize student
-messages to understand how students interact with the bot and identify patterns
-in their study habits.
-\\
-The categories outlined in \cite{Ghimire24} served as the
-foundation for classifying student messages (Debugging Help, Conceptual
-Questions, Code Snippet and Complete Solution). However, it was identified that
-some categories were missing, necessitating the addition of new categories
-(Student Correction, Multiple Question, Language Change and Uncategorized). Table
-\ref{tab:categories} presents these categories along with their descriptions.
-\\
-Subsequently, these eight categories were input into Claude.ai, accompanied by
-detailed instructions on the desired structure of the input and the expected
-format of the output. For each conversation with the CharlieBot log, a .csv file
-was downloaded, and Claude was tasked with classifying the user messages into
-one of the eight categories.
+The study was conducted in two phases: categorization and analysis. The
+categorization phase involved classifying student messages into eight
+categories. The table \ref{tab:categories} shows the categories of analysis
+and their description. The initial categories were based on the work
+of \cite{Ghimire24}, however, four new categories were added due to the need to
+give more detail to the classification. Beside of that, the categories were
+grouped to understand the impact on critical thinking.
 
-The classifications made by Claude were then manually reviewed. In cases where
-there was disagreement with the classification, manual adjustments were made,
-and the accuracy of Claude's classification was recorded. Each conversation
-with CharlieBot was saved into a .csv file containing the original conversation,
-along with two additional columns: "Classification" to denote the category and
-"AI" to indicate Claude's accuracy.
-
-As new categorizations were completed, an example was submitted to Claude's
-knowledge base to enhance accuracy. Initially, Claude began with an example
-representing 1\% of the total memory of its knowledge base. As classifications
-were progressively completed, more examples were provided to Claude to increase
-the percentage of its knowledge base occupied by relevant examples.
-
-Upon completion of all categorizations, all resulting .csv files were consolidated
-for final counts and analysis using the Pandas library. This methodology
-facilitated an efficient categorization of student messages, thereby enabling
-a comprehensive analysis of how tools such as ChatGPT are utilized in foundational
-programming courses.
+The analysis phase involved to understand the way that the students interact
+with CharlieBot. The students messages were collected from the logs of the bot
+and categorized manually by the authors. A total of YYY messages, of the first
+semester of 2024, were analyzed. The participants were students of the YYY
+semester of the Computer Science at UC3M where Java programming is taught.
 
@@
 
+The results of this study suggest that students interact with CharlieBot in a
+manner that promotes critical thinking. The majority of the messages analyzed
+were related to debugging help, conceptual questions, and student correction,
+indicating that students are actively engaging with the bot to understand
+concepts and solve problems. This is a positive outcome, as it suggests that
+CharlieBot is being used as a tool to enhance learning rather than a crutch for
+students to rely on.
+
 \bibliographystyle{IEEEtran}

### 2024-08-26 paper/paper.tex
update

@@
 
-\section{Background}
+\section{Related Work}
+
+The recent advancements in Generative Artificial Intelligence (GenAI) and its
+use in programming courses have raised possibilities and concerns among
+educators and researchers. The study of \cite{Yilmaz23} found that the main
+benefit of using ChatGPT in programming learning is to provide providing fast
+and mostly correct answers to questions, improving thinking skills, facilitating
+debugging, and increasing self-confidence. In contrast, the study the same
+appoint limitations such ask laziness, being unable to answer some questions,
+or giving incomplete/incorrect answers, causing professional anxiety in students.
 

### 2024-08-28 CSEDU2025/Paper.tex
update format

--- 2024-08-26 paper/paper.tex
@@
-\documentclass[conference]{IEEEtran}
-\IEEEoverridecommandlockouts
-% The preceding line is only needed to identify funding in the first footnote.
-% If that is unneeded, please comment it out.
-\usepackage{cite}
-\usepackage{amsmath,amssymb,amsfonts}
-\usepackage{algorithmic}
-\usepackage{graphicx}
-\usepackage{textcomp}
-\usepackage{xcolor}
-\usepackage{array}
+\documentclass[a4paper,twoside]{article}
+
+\usepackage{epsfig}
+\usepackage{subcaption}
+\usepackage{calc}
+\usepackage{amssymb}
+\usepackage{amstext}
+\usepackage{amsmath}
+\usepackage{amsthm}
+\usepackage{multicol}
+\usepackage{pslatex}
+\usepackage{algorithm2e}
+\usepackage[bottom]{footmisc}
+% Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
+\usepackage{natbib}
 \usepackage{multirow}
-\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
-    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
+\usepackage{SCITEPRESS}
+
 \begin{document}
@@
 \title{Analysis of Programming Students' Prompts to Identify Interactions
-Patterns with Generative AI \\
-{\footnotesize \textsuperscript{}}
-\thanks{The authors would like to thank the Federal Institute of Education,
-Science, and Technology of Rio Grande do Sul (IFRS) for the partial financial
-support provided for the execution of this research.}
+Patterns with Generative AI}
+
+\author{\authorname{Rodrigo Prestes Machado\sup{1}\orcidAuthor{0000-0003-0428-6387},
+Carlos Alario Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814} and
+Carlos Delgado Kloos\sup{2}\orcidAuthor{0000-0003-4093-3705}}
+\affiliation{\sup{1}Department of Informatics, Federal Institute of Education,
+Science and Technology, Porto Alegre, Brazil}
+\affiliation{\sup{2}Department Telematics Engineering, Carlos III University,
+Madrid, Spain}
+\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario,cdk\}@it.uc3m.es}
 }
 
-\author{\IEEEauthorblockN{1\textsuperscript{st} Rodrigo Prestes Machado}
-\IEEEauthorblockA{\textit{dept. Informática} \\
-\textit{Instituto Federal do Rio Grande do Sul}\\
-Porto Alegre, Brazil \\
-0000-0003-0428-6387}
-\and
-\IEEEauthorblockN{2\textsuperscript{nd} Carlos Alario Hoyos}
-\IEEEauthorblockA{\textit{Dep. Ingeniería Telemática} \\
-\textit{Universidad Carlos III de Madrid}\\
-Madrid, Spain \\
-0000-0002-3082-0814}
-\and
-\IEEEauthorblockN{3\textsuperscript{rd} Carlos Delgado Kloos}
-\IEEEauthorblockA{\textit{Dep. Ingeniería Telemática} \\
-\textit{Universidad Carlos III de Madrid}\\
-Madrid, Spain \\
-0000-0003-4093-3705}
-}
+\keywords{Generative Artificial Intelligence, Programming Education, Chatbots}
 
-\maketitle
+\abstract{The abstract should summarize the contents of the paper and should
+contain at least 70 and at most 200 words. The text must be set to 9-point font
+size.}
 
-\begin{abstract}
-This document is a model and instructions for \LaTeX.
-This and the IEEEtran.cls file define the components of your paper [title, text, heads, etc.]. *CRITICAL: Do Not Use Symbols, Special Characters, Footnotes, 
-or Math in Paper Title or Abstract.
-\end{abstract}
+\twocolumn\maketitle\normalsize\setcounter{footnote}{0}
 
-\begin{IEEEkeywords}
-component, formatting, style, styling, insert
-\end{IEEEkeywords}
-
-\section{Introduction}
+\section{\uppercase{Introduction}}
+\label{sec:introduction}
 
@@
 
-The study by Chan and Hu et al. \cite{chan23} revealed that both undergraduate
-and postgraduate students exhibit positive attitudes toward the use of GenAI
-in teaching and learning, noting that it enhances the depth of their thinking
-and understanding. Additionally, a systematic review by Lo, Hew, and Jong
-\cite{Lo24} demonstrated that students could effectively learn from ChatGPT,
-leading to improved understanding and academic achievement. Besides of that,
-it was also observed  that ChatGPT enables students to regulate their learning
-pace, promoting self-regulation \cite{Baha24} \cite{cai23}.
+The study of \cite{chan23} revealed that both undergraduate and postgraduate
+students exhibit positive attitudes toward the use of GenAI in teaching and
+learning, noting that it enhances the depth of their thinking and understanding.
+Additionally, a systematic review conducted by \cite{Lo24} demonstrated that
+students could effectively learn from ChatGPT, leading to improved understanding
+and academic achievement. Besides of that, it was also observed that ChatGPT
+enables students to regulate their learning pace, promoting self-regulation
+\citep{Baha24} \citep{cai23}.
 
 In the other hand, researchers have concerns about the impact of these
-tools on students. The systematic review of Vargas-Murillo et al.
-\cite{Murillo23} indicated that ChatGPT use might lead to an overreliance on the
-tool. Chan et al. \cite{chan23} also noted that this reliance could result in a
-decrease in critical thinking, as students might make decisions based solely on
-the information provided by ChatGPT.
+tools on students. The systematic review of \cite{Murillo23} indicated that
+ChatGPT use might lead to an overreliance on the tool. \cite{chan23} also noted
+that this reliance could result in a decrease in critical thinking, as students
+might make decisions based solely on the information provided by ChatGPT.
 
-The student's confidence is warranted, as shown by Puryear and Sprint
-\cite{Puryear22}, who demonstrated that GitHub Copilot can generate solutions
-for student assignments with accuracy rates ranging from 68\% to 95\%. However,
-this raises concerns that students might rely too heavily on GenAI tools,
-potentially neglecting a deeper understanding of the underlying concepts.
-Cai et al. \cite{cai23} further identified overdependence and reduced
-intellectual engagement as significant drawbacks of using ChatGPT in learning.
+The student's confidence is warranted, as shown by \cite{Puryear22}, who
+demonstrated that GitHub Copilot can generate solutions for student assignments
+with accuracy rates ranging from 68\% to 95\%. However, this raises concerns that
+students might rely too heavily on GenAI tools, potentially neglecting a deeper
+understanding of the underlying concepts. \cite{cai23} further identified
+overdependence and reduced intellectual engagement as significant drawbacks of
+using ChatGPT in learning.
 
-Regardless of teachers' preferences or beliefs, preliminary surveys by Dickey,
-Bejarano, and Garg \cite{Dickey24} revealed that at least 54.5\% of students are
-already using Generative AI (GenAI) for homework. This highlights the need to
-understand how students are interacting with these tools and how they can be
-used to enhance learning. Furthermore, a review by Lo, Hew, and Jong \cite{Lo24}
-emphasized the necessity for extended studies and objective measures to gather
-more robust evidence on the use of GenAI tools in education.
+Regardless of teachers' preferences or beliefs, preliminary surveys
+performed by \cite{Dickey24} revealed that at least 54.5\% of students are
+already using GenAI for homework. This highlights the need to understand how
+students are interacting with these tools and how they can be used to enhance
+learning. Furthermore, a review by \cite{Lo24} emphasized the necessity for
+extended studies and objective measures to gather more robust evidence on the
+use of GenAI tools in education.
 
@@
 To deal with this issue, some researchers have proposed different pedagogical
-strategies. Denny et al. \cite{Denny24} introduced the concept of \textit{Prompt
+strategies \cite{Denny24} introduced the concept of \textit{Prompt
 Problems}, where students solve programming exercises by formulating natural
-language prompts. Prasad and Aamod \cite{Prasad24} proposed a self-regulated
-learning framework using GenAI to solve programming problems. Lauren and Watta
-\cite{Lauren23} explored the integration of GenAI with evidence-based
-learning strategies in computer science and engineering courses.
+language prompts. \cite{Prasad24} proposed a self-regulated learning framework
+using GenAI to solve programming problems. \cite{Lauren23} explored the
+integration of GenAI with evidence-based learning strategies in computer science
+and engineering courses.
 
@@
 The study of effective educational strategies in use of GenAI tools in education
-context are essential for determining the best training for educators. However,
+context are important for determining the best training for educators. However,
 as these tools gain popularity among students \cite{Dickey24} the urgency to
@@
 Therefore, analyzing students' interaction patterns with GenAI tools is
-essential for understanding how these tools are being used and how educators can
+important for understanding how these tools are being used and how educators can
 leverage them to enhance learning. In addition, these interaction patterns can
 create new opportunities to redesign GenAI tools to better support pedagogical
-strategies and improve overall learning.
+strategies without repressing the development of abstract, critical and creative
+thinking.
 
@@
 Nevertheless, it is important to consider how GenAI tools affect students across
-different demographic groups, knowledge areas, cultural backgrounds \cite{catalan21}
-\cite{neo22} and students' prior experience. Thus, this works is limited to a
-specific group of students and a specific GenAI tool. The results may not be
-generalizable to other populations or tools.
+different demographic groups, knowledge areas, cultural backgrounds
+\cite{catalan21} \cite{neo22} and students' prior experience. Thus, this works
+is limited to a specific group of students and a specific GenAI tool. The
+results may not be generalizable to other populations or tools.
 
@@
 This study aims to analyze the interactions between students and CharlieBot,
-an ChatGPT-based bot developed by UC3M that leverages Retrieval-Augmented
+an ChatGPT 3.5 based bot developed by UC3M that leverages Retrieval-Augmented
 Generation (RAG) to enhance its contextual understanding of Java programming.
 
-\section{Background}
+\section{\uppercase{Related Work}}
 
-The paper \cite{Lo24} From the cognitive perspective, students were able to
+The advancements in GenAI and its use in programming courses have raised
+possibilities and concerns among educators and researchers. The study of
+\cite{Yilmaz23} found that the main benefit of using ChatGPT in programming
+learning is to provide providing fast and mostly correct answers to questions,
+improving thinking skills, facilitating debugging, and increasing
+self-confidence. In contrast, the study the same appoint limitations such ask
+laziness, being unable to answer some questions, or giving incomplete/incorrect
+answers, causing professional anxiety in students.
+
+From a cognitive perspective, \cite{Lo24} shown that students were able to
 learn from ChatGPT, which increased their understanding and achievement.
-However, concerns were raised that the growing use of AI tools might lead to
-a decline in critical thinking among students. Teachers and stakeholders should
+However, concerns were raised that the growing use of AI tools might lead to a
+decline in critical thinking among students. Teachers and stakeholders should
 continue to investigate pedagogical approaches that leverage ChatGPT to enhance
@@
 
-examined students from different demographic groups
-using AI tools like ChatGPT, comparing perceptions and impacts
-among these groups. The results showed that AI tools had a positive impact on
-academic performance, especially for students with less prior experience.
+Examined students from different demographic groups using AI tools like
+ChatGPT, comparing perceptions and impacts among these groups. The results
+showed that AI tools had a positive impact on academic performance, especially
+for students with less prior experience.
 
@@
 
-\subsection{The CharlieBot}
+\subsection{CharlieBot}
 
@@
 programming course at EdX to supply a RAG(Retrieval-Augmented Generation) system
-and thus try to respond more accurately to students' questions \cite{Sun24}.
-The main goal of CharlieBot is support students in their learning process by
+and thus try to respond more accurately to students' questions \cite{Sun24}. The
+main goal of CharlieBot is support students in their learning process by
 providing personalized assistance and feedback.
 
-\begin{table*}[htbp]
-\caption{Categories of Analysis}
-    \begin{center}
-    \begin{tabular}{|p{3cm}|p{5cm}|p{3cm}|}
-    \hline
-    \textbf{Category} & \textbf{Description} & \textbf{Critical Thinking} \\
-    \hline
-    Debugging Help & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \multirow{6}{3cm}{Tends to promote.} \\
-    Conceptual Questions & Prompts that are more about understanding concepts or algorithms rather than specific code. & \\
-    Student Correction & Prompts where the student corrects the bot. & \\
-    \hline
-    Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. & \multirow{6}{3cm}{Tends not to promote.} \\
-    Complete Solution & Prompts that request an entire solution or a complete code snippet. & \\
-    Multiple Question & Prompts where the user wants to solve a multiple-question exercise. & \\
-    \hline
-    Language Change & Prompts that request a change of idiom to convey a message more effectively. & \multirow{3}{3cm}{Neutral.} \\
-    Uncategorized & Prompts that do not fit into any of the above categories. & \\
-    \hline
-    \end{tabular}
-    \label{tab:categories}
-    \end{center}
-\end{table*}
-
-\section{Method}
+\section{\uppercase{Method}}
 
@@
 categorization phase involved classifying student messages into eight
-categories. The table \ref{tab:categories} shows the categories of analysis
-and their description. The initial categories were based on the work
-of \cite{Ghimire24}, however, four new categories were added due to the need to
+categories. The table \ref{tab:categories} shows the categories of analysis and
+their description. The initial categories were based on the work of
+\cite{Ghimire24}, however, four new categories were added due to the need to
 give more detail to the classification. Beside of that, the categories were
@@
 
+\begin{table*}[htbp]
+  \caption{Categories of Analysis}
+      \begin{center}
+      \begin{tabular}{|p{4cm}|p{6cm}|p{3cm}|}
+      \hline
+      \textbf{Category} & \textbf{Description} & \textbf{Critical Thinking} \\
+      \hline
+      Debugging Help & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \multirow{7}{3cm}{Tends to promote.} \\
+      Conceptual Questions & Prompts that are more about understanding concepts or algorithms rather than specific code. & \\
+      Student Correction & Prompts where the student corrects the bot. & \\
+      \hline
+      Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. & \multirow{6}{3cm}{Tends not to promote.} \\
+      Complete Solution & Prompts that request an entire solution or a complete code snippet. & \\
+      Multiple Question & Prompts where the user wants to solve a multiple-question exercise. & \\
+      \hline
+      Language Change & Prompts that request a change of idiom to convey a message more effectively. & \multirow{4}{3cm}{Neutral.} \\
+      Uncategorized & Prompts that do not fit into any of the above categories. & \\
+      \hline
+      \end{tabular}
+      \label{tab:categories}
+      \end{center}
+  \end{table*}
+
 \section{Results and Discussion}
 
-The graph \ref{fig:graph1} shows the distribution of the messages in the
+The Figure \ref{fig:graph1} shows the distribution of the messages in the
 categories:
+
 \begin{figure}[h!]
     \centering
-    \includegraphics[scale=0.7]{figures/figure1.png}
+    \includegraphics[scale=0.55]{img/figure1.png}
     \caption{Classification of messages.}
@@
 
-\bibliographystyle{IEEEtran}
-\bibliography{references}
+\section*{ACKNOWLEDGEMENTS}
+
+The authors would like to thank the Federal Institute of Education, Science and
+Technology (IFRS) for the partial financial support provided for the execution
+of this research.
+
+\bibliographystyle{apalike}
+{\small
+\bibliography{References}}
+
+\section*{\uppercase{Appendix}}
+
+If any, the appendix should appear directly after the
+references without numbering, and not on a new page. To do so please use the
+following command: \textit{$\backslash$section*\{APPENDIX\}}
 

### 2024-09-16 CSEDU2025/Paper.tex
classification

@@
 
+% Spacing
+We analyzed the time intervals between the user's interactions with the
+CharlieBot. Our analysis showed that 12.93\% messages thar combatilized 40\% of
+the conversations had response times longer than 60 minutes. In the case way,
+1.29\% of the messages and 12.00\% of the conversations  containing messages
+with delays exceeding 24 hours.
+
+% Categories Time
+
 \section{Conclusion}

### 2024-09-16 CSEDU2025/Paper.tex
update

--- 2024-08-28 CSEDU2025/Paper.tex
@@
 
+% Spacing
+Spacing, a well-established learning strategy, was examined by analyzing
+ the time intervals between users’ interactions with CharlieBot. Although
+ only 12\% of the conversations exhibited message delays exceeding 24 hours,
+40\% involved response times longer than 60 minutes. These findings suggest
+that students typically do not extend conversations across multiple days.
+However, the data also reveal that a substantial number of conversations
+feature one or more pauses exceeding an hour, a practice known to enhance
+knowledge retention.
+
+% Categories Time
+
 \section{Conclusion}

### 2024-09-19 CSEDU2025/Paper.tex
update

@@
 
-Regardless of teachers' preferences or beliefs, preliminary surveys
-performed by \cite{Dickey24} revealed that at least 54.5\% of students are
-already using GenAI for homework. This highlights the need to understand how
-students are interacting with these tools and how they can be used to enhance
-learning. Furthermore, a review by \cite{Lo24} emphasized the necessity for
-extended studies and objective measures to gather more robust evidence on the
-use of GenAI tools in education.
+Regardless of teachers' preferences or beliefs, a preliminary surveys performed
+by \cite{Dickey24} revealed that at least 54.5\% of students are already using
+GenAI for homework. This highlights the need to understand how students are
+interacting with these tools and how they can be used to enhance learning.
+Furthermore, a review by \cite{Lo24} emphasized the necessity for extended
+studies and objective measures to gather more robust evidence on the use of
+GenAI tools in education.
 
@@
 
-\subsection{Critical Thinking}
+% Critical thinking is a fundamental skill for students to develop, as it enables
+% them to analyze, evaluate, and synthesize information. This skill is essential
+% for problem-solving and decision-making, making it a crucial aspect of
+% educational success, specifically in programming courses. However, the use of AI
+% tools like ChatGPT has raised concerns about the potential impact on students'
+% critical thinking abilities \cite{Murillo23} \cite{cai23} \cite{chan23}.
 
-Critical thinking is a fundamental skill for students to develop, as it enables
-them to analyze, evaluate, and synthesize information. This skill is essential
-for problem-solving and decision-making, making it a crucial aspect of
-educational success, specifically in programming courses. However, the use of AI
-tools like ChatGPT has raised concerns about the potential impact on students'
-critical thinking abilities \cite{Murillo23} \cite{cai23} \cite{chan23}.
+% In the other hand, the study of \cite{zhang24} suggested that incorporating
+% ChatGPT could stimulate students' critical thinking, probably because they
+% had to assess the correctness of ChatGPT's answers.
 
-In the other hand, the study of \cite{zhang24} suggested that incorporating
-ChatGPT could stimulate students' critical thinking, probably because they
-had to assess the correctness of ChatGPT's answers.
+% Regarding self-regulation, \cite{cai23} noted that ChatGPT allowed students to
+% regulate their learning pace.
 
-Regarding self-regulation, \cite{cai23} noted that ChatGPT allowed students to
-regulate their learning pace.
+% \cite{wu24} found that ChatGPT-supported learning not only promoted students'
+% self-regulation but also increased their self-efficacy.
 
-\cite{wu24} found that ChatGPT-supported learning not only promoted students'
-self-regulation but also increased their self-efficacy.
+\begin{table*}[htbp]
+  \caption{Categories of Analysis}
+      \begin{center}
+      \begin{tabular}{|p{4cm}|p{9cm}|}
+      \hline
+      \textbf{Category} & \textbf{Description} \\
+      \hline
+      Debugging Help & * Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
+      Conceptual Question & * Prompts that are more about understanding concepts or algorithms rather than specific code. \\
+      Student Correction & * Prompts where the student corrects the bot. \\
+      \hline
+      Code Snippet & * Prompts that ask for a specific part of the code, like a function or a segment. \\
+      Complete Solution & * Prompts that request an entire solution or a complete code snippet. \\
+      Multiple Question & * Prompts where the user wants to solve a multiple-question exercise. \\
+      \hline
+      Language Change & * Prompts that request a change of idiom to convey a message more effectively. \\
+      Uncategorized & * Prompts that do not fit into any of the above categories. \\
+      \hline
+      \end{tabular}
+      \label{tab:categories}
+      \end{center}
+  \end{table*}
+
+\subsection{Metacognition}
+
+The use of GenAI tools can help students learning by providing rapid assistance
+and feedback. However, it is important to teach students how to use these tools
+properly to promote learning. In this sense, it is important that students are
+aware of the cognitive processes.
+
+Metacognition is defined as the awareness about the cognitive processes.
+\cite{flavell79} defined a model of cognitive monitoring in four classes of
+phenomena that can act or interact with each other, which are: Metacognitive
+Knowledge, Metacognitive Experience, Metacognitive Task, and Metacognitive
+Strategies. Metacognitive knowledge refers beliefs about what factor or
+variables affect the course and outcome of cognitive enterprise. It is
+subdivided into three majors categories or variables: Person, Task, and
+Strategy. Metacognitive experience refers to the feelings and emotions
+associated with cognitive activities. Metacognitive task refers to the type of
+task or problem that the individual is facing. Metacognitive strategies refer to
+the methods or techniques that individuals use to control their cognitive
+processes.
 
@@
 
-\begin{table*}[htbp]
-  \caption{Categories of Analysis}
-      \begin{center}
-      \begin{tabular}{|p{4cm}|p{6cm}|p{3cm}|}
-      \hline
-      \textbf{Category} & \textbf{Description} & \textbf{Critical Thinking} \\
-      \hline
-      Debugging Help & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \multirow{7}{3cm}{Tends to promote.} \\
-      Conceptual Questions & Prompts that are more about understanding concepts or algorithms rather than specific code. & \\
-      Student Correction & Prompts where the student corrects the bot. & \\
-      \hline
-      Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. & \multirow{6}{3cm}{Tends not to promote.} \\
-      Complete Solution & Prompts that request an entire solution or a complete code snippet. & \\
-      Multiple Question & Prompts where the user wants to solve a multiple-question exercise. & \\
-      \hline
-      Language Change & Prompts that request a change of idiom to convey a message more effectively. & \multirow{4}{3cm}{Neutral.} \\
-      Uncategorized & Prompts that do not fit into any of the above categories. & \\
-      \hline
-      \end{tabular}
-      \label{tab:categories}
-      \end{center}
-  \end{table*}
-
 \section{Results and Discussion}
@@
     \centering
-    \includegraphics[scale=0.55]{img/figure1.png}
+    \includegraphics[scale=0.6]{img/figure1.png}
     \caption{Classification of messages.}
@@
 
-Approximately 64\% of the messages pertain to categories associated with
-critical thinking, corroborating with the findings of
-\cite{Ghimire24}. In contrast, around 28\% of the messages
-indicate a preference among students for ready-made answers.
+% Approximately 64\% of the messages pertain to categories associated with
+% critical thinking, corroborating with the findings of
+% \cite{Ghimire24}. In contrast, around 28\% of the messages
+% indicate a preference among students for ready-made answers.
+
+% Uncategorized
+
+The Figure \ref{fig:graph1} shows that around 5\% of the prompts
+written by students were not categorized. These messages range from
+greetings to the Bot, such as a simple "\textit{Hola}," to
+contextual statements like "\textit{Estoy repasando orientación a
+objetos.}"
+
+% Multiple Choice
+\subsubsection*{Sequence of Multiple Choice}
+
+Around 16.9\% of the messages were classified as Multiple Choice. These prompts
+typically involve students sending one or more exercises to CharlieBot, asking
+for the solution.
+
+We identified a sequence of Multiple Choice patterns in some of the
+conversations. From the point of view of Metacognitive Knowledge, a possible
+interpretation of this complex phenomenon is that students may believe in their
+inability to answer several questions. They may also assume that the questions
+share a common theme, and solving them together may help consolidate knowledge.
+Additionally, they might understand that obtaining the answers quickly is more
+efficient.
+
+Regarding Metacognitive Experiences, students may experience feelings such as
+cognitive overload, a sense of urgency, and/or confidence in ready-made answers,
+which can often lead to relief and a reduction of frustration.
+
+Concerning the Metacognitive about the Task, one interpretation may be that the
+student simply wants to complete the task quickly to alleviate the workload,
+and/or obtain a correct answer without focusing on deeper learning at
+this moment.
+
+In relation to Metacognitive Strategies, students are likely focusing on saving
+time and/or possibly lacking planning skills. They may also perceive that
+solving the questions on their own requires too much cognitive effort, but this
+pattern might reveal a lack of progress monitoring and understanding of
+concepts. After obtaining the answers, the metacognitive evaluation process
+could lead to feelings of satisfaction, but further reflection may indicate that
+the strategy was not effective in supporting deep learning. Depending on this
+evaluation, it may influence future decision-making to change the strategy.
+
+% Conceptual Questions
+\subsubsection*{Sequence of Conceptual Question}
+
+Messages classified as conceptual questions represent around 32\% of all
+messages. Regarding Metacognitive Knowledge, about these sequences of prompts
+related with conceptual questions, one possible interpretation may be that the
+students are aware they have not fully mastered a concept and are trying to
+understand it, demonstrating a good level of self-awareness. In this sense, the
+student realizes that the problem requires an understanding of concepts in order
+to advance in their studies. In this way, the student applies an active learning
+strategy. About Metacognitive Experience, several types of feelings may be
+associated at this moment, such as actively reflecting on the answers,
+uncertainty and/or confusion when trying to understand the concept, frustration
+for not progressing, and satisfaction as their questions are answered in a
+way that facilitates understanding. Regarding the type of metacognition related
+to the Task, the student recognized the difficult and may be trying to achieve a
+deeper understanding of the evolved concepts to later apply in practice. With
+respect to Metacognitive Strategies, the student demonstrates a metacognitive
+planning process when identifying their knowledge gaps. They may start with more
+general questions to obtain an overview of the concept and then ask more
+specific questions as they gain more clarity. Because the questions are in
+sequence, the student probably can adequately monitor their progress.
 

### 2024-09-20 CSEDU2025/Paper.tex
update

@@
 % \cite{wu24} found that ChatGPT-supported learning not only promoted students'
-% self-regulation but also increased their self-efficacy.
-
 \begin{table*}[htbp]
-  \caption{Categories of Analysis}
+  \caption{Categories - adapted from \cite{Ghimire24}}
   \begin{center}
-      \begin{tabular}{|p{4cm}|p{9cm}|}
+    \renewcommand{\arraystretch}{1.5} % Increase the spacing between rows
+    \begin{tabular}{|p{3,5cm}|p{11cm}|}
       \hline
@@
       \hline
-      Debugging Help & * Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
-      Conceptual Question & * Prompts that are more about understanding concepts or algorithms rather than specific code. \\
-      Student Correction & * Prompts where the student corrects the bot. \\
+      Debugging Help & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
+      Conceptual Question & Prompts that are more about understanding concepts than specific code. \\
+      Student Correction & Prompts where the student corrects the bot. \\
       \hline
-      Code Snippet & * Prompts that ask for a specific part of the code, like a function or a segment. \\
-      Complete Solution & * Prompts that request an entire solution or a complete code snippet. \\
-      Multiple Question & * Prompts where the user wants to solve a multiple-question exercise. \\
+      Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. \\
+      Complete Solution & Prompts that request an entire solution or a complete code snippet. \\
+      Multiple Question & Prompts where the user wants to solve a multiple-question exercise (Quiz). \\
       \hline
-      Language Change & * Prompts that request a change of idiom to convey a message more effectively. \\
-      Uncategorized & * Prompts that do not fit into any of the above categories. \\
+      Language Change & Prompts that request a change of idiom. \\
+      Uncategorized & Prompts that do not fit into any of the above categories. \\
       \hline
@@
 
-The study was conducted in two phases: categorization and analysis. The
-categorization phase involved classifying student messages into eight
+The study was conducted in three phases: data collection, categorization, and
+analysis. During the first semester of 2024, student interactions with
+CharlieBot were collected as part of the Java programming course at UC3M.
+The categorization phase involved classifying student messages into eight
 categories. The table \ref{tab:categories} shows the categories of analysis and
@@
 \cite{Ghimire24}, however, four new categories were added due to the need to
-give more detail to the classification. Beside of that, the categories were
-grouped to understand the impact on critical thinking.
-
-The analysis phase involved to understand the way that the students interact
-with CharlieBot. The students messages were collected from the logs of the bot
-and categorized manually by the authors. A total of YYY messages, of the first
-semester of 2024, were analyzed. The participants were students of the YYY
-semester of the Computer Science at UC3M where Java programming is taught.
-
-\section{Results and Discussion}
+give more detail to the classification. The analysis phase involved to
+understand the way that the students interact with CharlieBot.
+\section{\uppercase{Results and Discussion}}
 
@@
     \centering
-    \includegraphics[scale=0.6]{img/figure1.png}
+    \includegraphics[scale=0.62]{img/figure1.png}
     \caption{Classification of messages.}
@@
 \end{figure}
+
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.52]{img/figure2.png}
+  \caption{Examples of conversations.}
+  \label{fig:graph2}
+\end{figure*}
 
@@
 
+% Debugging help
+\subsubsection*{Sequence of Debugging Help}
+
+Around 30.9\% of the messages were classified as Debugging Help. Besides of
+that, we identified a pattern of a sequences of Debugging Help in the
+conversations. These prompts typically start with a message requesting a
+Complete Solution or a Code Snippet. After that, the students normally chance
+from passive to more active learning strategy.
+
+In terms of Metacognitive Knowledge, the student probably recognizes that they
+do not have the necessary tools or knowledge to debug the code on their own.
+Concerning the understanding of the task, they may know that debugging is not
+just a matter of superficial corrections but requires understanding the
+underlying logic. Thus, the student apparently believes that the sequence of
+questions related to debugging will help them understand the codes.
+
+In terms of Metacognitive Experiences, some feelings may arise at this moment,
+such as frustration or uncertainty in the face of errors, confusion during the
+process, and also growing confidence or satisfaction as they better understand
+the code and possible corrections.
+
+With respect to Metacognitive related with the objective of the Task, the
+student is probably focused on understanding the code in practical terms,
+identifying errors, and learning how to better debug in the future.
+
+As for Metacognitive associated with the Strategy, the student has requested
+external help and realizes that dividing the debugging task into several
+questions can result in a better understanding of the code. They can monitor
+their progress throughout the process, adjusting their questions as necessary,
+evaluate the effectiveness of the responses received, and reflect on what was
+learned during the process.
+
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.55]{img/figure3.png}
+  \caption{Interaction times per categories.}
+  \label{fig:graph3}
+\end{figure*}
+
 % Conceptual Questions
@@
 Messages classified as conceptual questions represent around 32\% of all
-messages. Regarding Metacognitive Knowledge, about these sequences of prompts
-related with conceptual questions, one possible interpretation may be that the
-students are aware they have not fully mastered a concept and are trying to
-understand it, demonstrating a good level of self-awareness. In this sense, the
-student realizes that the problem requires an understanding of concepts in order
-to advance in their studies. In this way, the student applies an active learning
-strategy. About Metacognitive Experience, several types of feelings may be
-associated at this moment, such as actively reflecting on the answers,
-uncertainty and/or confusion when trying to understand the concept, frustration
-for not progressing, and satisfaction as their questions are answered in a
-way that facilitates understanding. Regarding the type of metacognition related
-to the Task, the student recognized the difficult and may be trying to achieve a
-deeper understanding of the evolved concepts to later apply in practice. With
-respect to Metacognitive Strategies, the student demonstrates a metacognitive
-planning process when identifying their knowledge gaps. They may start with more
-general questions to obtain an overview of the concept and then ask more
-specific questions as they gain more clarity. Because the questions are in
-sequence, the student probably can adequately monitor their progress.
+messages. We have found a pattern of sequences of Conceptual Questions in some
+conversations. Some conversations are entirely composed of messages classified
+as Conceptual Questions. In other cases, the sequence of Conceptual questions
+starts with a request for a Complete Solution or a Code Snippet, followed by
+a progression of Conceptual Questions. In both cases, the student is actively
+engaging with the bot to understand the concepts and compromised with more
+deep learning.
+
+Regarding Metacognitive Knowledge, these sequences of prompts classified as
+Conceptual Question, can be interpreted that the students are aware they have
+not fully mastered a concept and are trying to understand it, demonstrating a
+good level of self-awareness. In this sense, the student realizes that the
+problem requires an understanding of concepts in order to advance in their
+studies. In this way, the student applies an active learning strategy.
+
+About Metacognitive Experience, several types of feelings may be associated at
+this moment, such as actively reflecting on the answers, uncertainty and/or
+confusion when trying to understand the concept, frustration for not
+progressing, and satisfaction as their questions are answered in a way that
+facilitates understanding.
+
+Regarding the type of Metacognition related to objective of the Task, the
+students may recognize the difficult and trying to achieve a deeper
+understanding of the evolved concepts to later apply in practice.
+
+With respect to Metacognitive Strategies, the student demonstrates a sence of
+planning to identify their knowledge gaps. They may start with more general
+questions to obtain an overview of the concept and then ask more specific
+questions as they gain more clarity. Because the questions are in sequence, the
+student probably can adequately monitor their progress.
+
+\subsection*{Lack of Interleaving}
+
+% Interleaving
+Interleaving, a technique that involves alternating the conversation among
+different types of categories.
+
+% Eglington, L. G., & Kang, S. H. (2017). Interleaved presentation benefits
+% science category learning. Journal of Applied Research in Memory and Cognition,
+% 6, 475-485
+
+\subsection*{Spacing}
 
 % Spacing
-Spacing, a well-established learning strategy, was examined by analyzing
-the time intervals between users’ interactions with CharlieBot. Although
-only 12\% of the conversations exhibited message delays exceeding 24 hours,
-40\% involved response times longer than 60 minutes. These findings suggest
-that students typically do not extend conversations across multiple days.
-However, the data also reveal that a substantial number of conversations
-feature one or more pauses exceeding an hour, a practice known to enhance
-knowledge retention.
-
-% Categories Time
+Spacing, a well-established learning strategy \cite{Carvalho20}, was examined
+by analyzing the time intervals between users’ interactions with CharlieBot.
+Although only 7.4\% of the conversations exhibited message delays exceeding 24
+hours, 26.19\% involved response times longer than 60 minutes. These findings
+suggest that students typically do not extend conversations across multiple days.
+However, the data also reveal that some conversations had one or more pauses
+exceeding an hour, a practice known to enhance knowledge retention.
 

### 2024-09-24 CSEDU2025/Paper.tex
update

@@
 
+% Since most of the messages are located in categories where students are 
+% notoriously more active in their learning (conceptual questions 32.1% and 
+% Debugging Help 30.9%), it seems that students are trying to use the AI in a 
+% way that promotes their learning. However, the pattern of conversations with 
+% sequential blocks and also little spacing may suggest that there is a gap in 
+% knowledge of study strategies.
+
 % Uncategorized
@@
 
+To highlight the low incidence of interleaving, we conducted an analysis to
+identify blocks of four or more consecutive messages classified within the same
+category. The results showed that approximately 36\% of the messages are
+distributed across 22 blocks. Specifically, the Debugging Help category contains
+10 blocks of four or more messages, Conceptual Question has seven blocks,
+Multiple Choice has three blocks, while Code Snippet and Complete Solution each
+have one block.
+
+Interestingly, when expanding the analysis to blocks of three consecutive
+messages, the percentage of messages grouped within blocks of the same category
+increases to about 45% of the total. It is also worth noting that three
+categories – Student Correction, Language Change, and Uncategorized – did not
+form any sequential blocks. The figure \ref{fig:graph2} shows 15 conversations,
+one clear example of blocks can be observed in conversation eight which contains
+two blocks of Multiple Choice questions. Another example is conversation 10,
+with the entire conversation forming a one block of Conceptual Questions.
+
 \subsection*{Spacing}
@@
 exceeding an hour, a practice known to enhance knowledge retention.
+
+% Another interesting study technique is spacing. Only 7.4% of the conversations
+% exhibited message delays exceeding 24 hours and 26.19% involved response times 
+% longer than 60 minutes. However, Carlos told me that this may be due to some 
+% limitation of the bot, as it only maintains a conversation while the browser 
+ % tab is open.
+
+ % There is a big difference (I haven't looked at it statistically yet) in the 
+%  interaction time with the Bot when a student asks conceptual questions. 
+%  Perhaps this actually makes sense, since it is a category where, from the 
+%  point of view of the metacognitive objective, the student is more concerned 
+%  with understanding a concept and deeper learning.
 

### 2024-09-25 CSEDU2025/Paper.tex
update

@@
 
-\subsection*{Lack of Interleaving}
+\subsection*{Interleaving}
 
 % Interleaving
-Interleaving, a technique that involves alternating the conversation among
-different types of categories.
 
-% Eglington, L. G., & Kang, S. H. (2017). Interleaved presentation benefits
-% science category learning. Journal of Applied Research in Memory and Cognition,
-% 6, 475-485
+Intervealing is a study strategy that involves mixing or alternating different
+topics or problems within a single learning session. This
+approach has been shown to be useful for promoting long-term learning and
+improving the application of knowledge in other situations \cite{Rivers21}.
 
-To highlight the low incidence of interleaving, we conducted an analysis to
-identify blocks of four or more consecutive messages classified within the same
-category. The results showed that approximately 36\% of the messages are
-distributed across 22 blocks. Specifically, the Debugging Help category contains
-10 blocks of four or more messages, Conceptual Question has seven blocks,
-Multiple Choice has three blocks, while Code Snippet and Complete Solution each
-have one block.
+When analyzing the classification of messages in conversations, we identified
+the presence of sequences of messages belonging to the same category, that it
+could suggest a low occurrence of interleaving. To better assess the extent of
+these sequences, we calculated the percentage of messages that were part of
+repeated blocks.
 
-Interestingly, when expanding the analysis to blocks of three consecutive
-messages, the percentage of messages grouped within blocks of the same category
-increases to about 45% of the total. It is also worth noting that three
-categories – Student Correction, Language Change, and Uncategorized – did not
-form any sequential blocks. The figure \ref{fig:graph2} shows 15 conversations,
-one clear example of blocks can be observed in conversation eight which contains
-two blocks of Multiple Choice questions. Another example is conversation 10,
-with the entire conversation forming a one block of Conceptual Questions.
+Considering blocks with at least three messages, we observed that 56.2\% of the
+students' prompts were grouped into some sequence. When analyzing blocks with a
+minimum of 4 messages, we found that 47.5\% of the prompts were part of
+continuous sequences. These results indicate a moderate incidence of
+interleaving among the students' messages.
+
+To illustrate the presence of blocks the Figure \ref{fig:graph2} presents
+examples of conversations and their classification. In Conversation 2, the
+student alternates messages among Code Snippet, Debugging Help and Conceptual
+Question, showing an tendency of interleaving between practical and theoretical
+perspectives. In contrast, in Conversation 10, the student focused on a single
+block of Conceptual Question. Although the focus was on understanding a
+theoretical concept, this conversation tended not to promote interleaving, due
+to the absence of alternation between different types of tasks.
 

### 2024-09-26 CSEDU2025/Paper.tex
update paper

@@
 
-\title{Analysis of Programming Students' Prompts to Identify Interactions
-Patterns with Generative AI}
+\title{Analysis of Programming Students' Prompts with Generative AI}
 
 \author{\authorname{Rodrigo Prestes Machado\sup{1}\orcidAuthor{0000-0003-0428-6387},
-Carlos Alario Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814} and
+Carlos Alario Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814},
+Patricia Callejo Pinardo\sup{2}\orcidAuthor{0000-0001-6124-6213},
+Iria Manuela Estévez Ayres\sup{2}\orcidAuthor{0000-0002-1047-5398},
 Carlos Delgado Kloos\sup{2}\orcidAuthor{0000-0003-4093-3705}}
@@
 Science and Technology, Porto Alegre, Brazil}
-\affiliation{\sup{2}Department Telematics Engineering, Carlos III University,
+\affiliation{\sup{2}Department of Telematics Engineering, Carlos III University,
 Madrid, Spain}
-\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario,cdk\}@it.uc3m.es}
+\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario,pcallejo, ayres,cdk\}@it.uc3m.es}
 }
@@
 
-%What is the problem to be solved?
+% %What is the problem to be solved?
 
@@
 
-% Are there any existing solutions?
+% % Are there any existing solutions?
 
@@
 
-% Which is the best?
+% % Which is the best?
 
@@
 
-% What is its main limitation?
+% % What is its main limitation?
 
@@
 
-% What do you hope to achieve?
+% % What do you hope to achieve?
 
@@
 Generation (RAG) to enhance its contextual understanding of Java programming.
+We created three research questions to guide our analysis:
 
-\section{\uppercase{Related Work}}
+RQ1: What was the distribution pattern of the classified messages?
 
-The advancements in GenAI and its use in programming courses have raised
-possibilities and concerns among educators and researchers. The study of
-\cite{Yilmaz23} found that the main benefit of using ChatGPT in programming
-learning is to provide providing fast and mostly correct answers to questions,
-improving thinking skills, facilitating debugging, and increasing
-self-confidence. In contrast, the study the same appoint limitations such ask
-laziness, being unable to answer some questions, or giving incomplete/incorrect
-answers, causing professional anxiety in students.
+RQ2: Was there a time gap between student submissions? If so, which category
+showed the largest spacing between submissions?
 
-From a cognitive perspective, \cite{Lo24} shown that students were able to
-learn from ChatGPT, which increased their understanding and achievement.
-However, concerns were raised that the growing use of AI tools might lead to a
-decline in critical thinking among students. Teachers and stakeholders should
-continue to investigate pedagogical approaches that leverage ChatGPT to enhance
-students' understanding and critical thinking. For example, teachers can
-instruct students to fact-check and validate information produced by ChatGPT.
+RQ3: Do students’ conversations with the bot alternate between practical and
+theoretical approaches, thereby promoting interleaving?
 
-Examined students from different demographic groups using AI tools like
-ChatGPT, comparing perceptions and impacts among these groups. The results
-showed that AI tools had a positive impact on academic performance, especially
-for students with less prior experience.
+\section{\uppercase{Theoretical Framework}}
+
+\subsection{Related Work}
+
+% The advancements in GenAI and its use in programming courses have raised
+% possibilities and concerns among educators and researchers. The study of
+% \cite{Yilmaz23} found that the main benefit of using ChatGPT in programming
+% learning is to provide providing fast and mostly correct answers to questions,
+% improving thinking skills, facilitating debugging, and increasing
+% self-confidence. In contrast, the study the same appoint limitations such ask
+% laziness, being unable to answer some questions, or giving incomplete/incorrect
+% answers, causing professional anxiety in students.
+
+% From a cognitive perspective, \cite{Lo24} shown that students were able to
+% learn from ChatGPT, which increased their understanding and achievement.
+% However, concerns were raised that the growing use of AI tools might lead to a
+% decline in critical thinking among students. Teachers and stakeholders should
+% continue to investigate pedagogical approaches that leverage ChatGPT to enhance
+% students' understanding and critical thinking. For example, teachers can
+% instruct students to fact-check and validate information produced by ChatGPT.
+
+% Examined students from different demographic groups using AI tools like
+% ChatGPT, comparing perceptions and impacts among these groups. The results
+% showed that AI tools had a positive impact on academic performance, especially
+% for students with less prior experience.
 
@@
 % regulate their learning pace.
-
-% \cite{wu24} found that ChatGPT-supported learning not only promoted students'
 \begin{table*}[htbp]
@@
   \begin{center}
-    \renewcommand{\arraystretch}{1.5} % Increase the spacing between rows
-    \begin{tabular}{|p{3,5cm}|p{11cm}|}
+    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
+    \begin{tabular}{p{3cm} p{12cm}} % Remove vertical bars
       \hline
@@
       Student Correction & Prompts where the student corrects the bot. \\
-      \hline
       Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. \\
       Complete Solution & Prompts that request an entire solution or a complete code snippet. \\
-      Multiple Question & Prompts where the user wants to solve a multiple-question exercise (Quiz). \\
-      \hline
+      Multiple Question & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
       Language Change & Prompts that request a change of idiom. \\
@@
 
-The use of GenAI tools can help students learning by providing rapid assistance
-and feedback. However, it is important to teach students how to use these tools
-properly to promote learning. In this sense, it is important that students are
-aware of the cognitive processes.
+The use of generative artificial intelligence tools can support the important
+mechanism of learning regulation by students \citep{larsen18}, since these
+technologies frequently generate automatic and often accurate responses
+\cite{Puryear22}, it becomes essential for students to use these tools more
+strategically to promote conscious and active learning.
 
-Metacognition is defined as the awareness about the cognitive processes.
-\cite{flavell79} defined a model of cognitive monitoring in four classes of
-phenomena that can act or interact with each other, which are: Metacognitive
-Knowledge, Metacognitive Experience, Metacognitive Task, and Metacognitive
-Strategies. Metacognitive knowledge refers beliefs about what factor or
-variables affect the course and outcome of cognitive enterprise. It is
-subdivided into three majors categories or variables: Person, Task, and
-Strategy. Metacognitive experience refers to the feelings and emotions
-associated with cognitive activities. Metacognitive task refers to the type of
-task or problem that the individual is facing. Metacognitive strategies refer to
-the methods or techniques that individuals use to control their cognitive
-processes.
+The Theory of Metacognition addresses awareness of mental processes.
+\cite{flavell79} proposed a model of metacognitive monitoring that includes four
+interrelated phenomena: Metacognitive Knowledge, Metacognitive Experience,
+ Metacognitive Goals, and Metacognitive Actions. These processes do not occur
+in isolation, but they influence each other, altering cognitive progress over
+time.
+
+Metacognitive knowledge includes beliefs about variables that affect the
+outcomes of cognitive activities. It is divided into three types: beliefs about
+personal abilities, perceived difficulty in the task, and previously used
+strategies. Metacognitive Experience refers to the feelings that arise
+before, during, and after cognitive activity, such as frustration, confusion,
+and satisfaction, among others. Metacognitive goals are the key to regulating
+thought as they relate to the goals the individual seeks to achieve, directly
+influencing the actions taken. For example, if a student’s goal is to complete
+a task quickly, they may adopt a more passive approach to learning. Lastly,
+Metacognitive Actions involve the planning, monitoring, and evaluation of
+strategies used to achieve the goals. In terms of planning, students can
+determine how to approach a task, such as spacing their study sessions
+\citep{Ouhao18, Carvalho20}, interleaving topics \citep{Rivers21}, utilizing
+retrieval practice \citep{larsen18}, and other strategies.
+
+\section{\uppercase{Material and Method}}
+
+This section outlines the tools and procedures used in this study. CharlieBot
+served as the GenAI tool for educational purposes, and the method was structured
+into three distinct phases, each designed to systematically assess its
+interaction with students.
 
@@
 
-CharlieBot is a ChatGPT 3.5 based bot that provides assistance to students in a
-Java programming courses at UC3M. The bot uses the content of UC3M Java
-programming course at EdX to supply a RAG(Retrieval-Augmented Generation) system
-and thus try to respond more accurately to students' questions \cite{Sun24}. The
-main goal of CharlieBot is support students in their learning process by
-providing personalized assistance and feedback.
+CharlieBot is an educational bot built on ChatGPT 3.5 and enhanced with
+Retrieval-Augmented Generation (RAG) capabilities to improve the precision of
+responses provided to students \cite{Sun24}. The RAG feature was developed using
+materials from a MOOC, including text, exercises, video transcripts, and other
+resources.
 
-\section{\uppercase{Method}}
+\subsection{Method}
 
-The study was conducted in three phases: data collection, categorization, and
-analysis. During the first semester of 2024, student interactions with
-CharlieBot were collected as part of the Java programming course at UC3M.
-The categorization phase involved classifying student messages into eight
-categories. The table \ref{tab:categories} shows the categories of analysis and
-their description. The initial categories were based on the work of
-\cite{Ghimire24}, however, four new categories were added due to the need to
-give more detail to the classification. The analysis phase involved to
-understand the way that the students interact with CharlieBot.
+The study was conducted in three phases: data collection,
+categorization, and analysis. During the data collection phase, students
+enrolled in a second semester Java course at a public university in Spain were
+introduced to CharlieBot and allowed to use it freely, without being subjected
+to a specific methodology for its use. Data were collected anonymously, ensuring
+that the records of interactions could not be linked to any individual student
+or their academic performance.
+
+In the categorization phase, the students' messages were manually classified
+into eight distinct categories for analysis. Initially, \cite{Ghimire24}
+proposed four categories: Debugging Help, Conceptual Question, Code Snippet,
+and Complete Solution. However, the data collected indicated the need for
+additional categories, resulting in the inclusion of four more: Multiple Choice
+(Quiz), Student Corrections, Language Change, and Uncategorized. Table
+\ref{tab:categories} presents these categories along with their respective
+descriptions.
+
+The analysis phase involved the use of Python scripts in the Jupyter Notebook
+environment to extract relevant information from previously classified
+data. The datasets and code snippets used in this process are available at: YYY.
+
 \section{\uppercase{Results and Discussion}}
+
+The results of the analysis are presented in this section, addressing the
+research questions posed in the introduction.
 
@@
 
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.52]{img/figure2.png}
-  \caption{Examples of conversations.}
-  \label{fig:graph2}
-\end{figure*}
+\subsection{Classification of Messages}
+
+The Figure \ref{fig:graph1} shows that around 5\% of the prompts
+written by students were not categorized. These messages range from
+greetings to the Bot, such as a simple "\textit{Hola}," to
+contextual statements like "\textit{Estoy repasando orientación a
+objetos.}"
 
@@
 
-% Uncategorized
-
-The Figure \ref{fig:graph1} shows that around 5\% of the prompts
-written by students were not categorized. These messages range from
-greetings to the Bot, such as a simple "\textit{Hola}," to
-contextual statements like "\textit{Estoy repasando orientación a
-objetos.}"
-
-% Multiple Choice
-\subsubsection*{Sequence of Multiple Choice}
-
-Around 16.9\% of the messages were classified as Multiple Choice. These prompts
-typically involve students sending one or more exercises to CharlieBot, asking
-for the solution.
-
-We identified a sequence of Multiple Choice patterns in some of the
-conversations. From the point of view of Metacognitive Knowledge, a possible
-interpretation of this complex phenomenon is that students may believe in their
-inability to answer several questions. They may also assume that the questions
-share a common theme, and solving them together may help consolidate knowledge.
-Additionally, they might understand that obtaining the answers quickly is more
-efficient.
-
-Regarding Metacognitive Experiences, students may experience feelings such as
-cognitive overload, a sense of urgency, and/or confidence in ready-made answers,
-which can often lead to relief and a reduction of frustration.
-
-Concerning the Metacognitive about the Task, one interpretation may be that the
-student simply wants to complete the task quickly to alleviate the workload,
-and/or obtain a correct answer without focusing on deeper learning at
-this moment.
-
-In relation to Metacognitive Strategies, students are likely focusing on saving
-time and/or possibly lacking planning skills. They may also perceive that
-solving the questions on their own requires too much cognitive effort, but this
-pattern might reveal a lack of progress monitoring and understanding of
-concepts. After obtaining the answers, the metacognitive evaluation process
-could lead to feelings of satisfaction, but further reflection may indicate that
-the strategy was not effective in supporting deep learning. Depending on this
-evaluation, it may influence future decision-making to change the strategy.
-
-% Debugging help
-\subsubsection*{Sequence of Debugging Help}
-
-Around 30.9\% of the messages were classified as Debugging Help. Besides of
-that, we identified a pattern of a sequences of Debugging Help in the
-conversations. These prompts typically start with a message requesting a
-Complete Solution or a Code Snippet. After that, the students normally chance
-from passive to more active learning strategy.
-
-In terms of Metacognitive Knowledge, the student probably recognizes that they
-do not have the necessary tools or knowledge to debug the code on their own.
-Concerning the understanding of the task, they may know that debugging is not
-just a matter of superficial corrections but requires understanding the
-underlying logic. Thus, the student apparently believes that the sequence of
-questions related to debugging will help them understand the codes.
-
-In terms of Metacognitive Experiences, some feelings may arise at this moment,
-such as frustration or uncertainty in the face of errors, confusion during the
-process, and also growing confidence or satisfaction as they better understand
-the code and possible corrections.
-
-With respect to Metacognitive related with the objective of the Task, the
-student is probably focused on understanding the code in practical terms,
-identifying errors, and learning how to better debug in the future.
-
-As for Metacognitive associated with the Strategy, the student has requested
-external help and realizes that dividing the debugging task into several
-questions can result in a better understanding of the code. They can monitor
-their progress throughout the process, adjusting their questions as necessary,
-evaluate the effectiveness of the responses received, and reflect on what was
-learned during the process.
-
 \begin{figure*}[htbp]
@@
 
-% Conceptual Questions
-\subsubsection*{Sequence of Conceptual Question}
+\subsection{Spacing}
 
-Messages classified as conceptual questions represent around 32\% of all
-messages. We have found a pattern of sequences of Conceptual Questions in some
-conversations. Some conversations are entirely composed of messages classified
-as Conceptual Questions. In other cases, the sequence of Conceptual questions
-starts with a request for a Complete Solution or a Code Snippet, followed by
-a progression of Conceptual Questions. In both cases, the student is actively
-engaging with the bot to understand the concepts and compromised with more
-deep learning.
+% Spacing
+Spacing, a well-established learning strategy \citep{Carvalho20}, was examined
+by analyzing the time intervals between users’ interactions with CharlieBot.
+Although only 7.4\% of the conversations exhibited message delays exceeding 24
+hours, 26.19\% involved response times longer than 60 minutes. These findings
+suggest that students typically do not extend conversations across multiple days.
+However, the data also reveal that some conversations had one or more pauses
+exceeding an hour, a practice known to enhance knowledge retention.
 
-Regarding Metacognitive Knowledge, these sequences of prompts classified as
-Conceptual Question, can be interpreted that the students are aware they have
-not fully mastered a concept and are trying to understand it, demonstrating a
-good level of self-awareness. In this sense, the student realizes that the
-problem requires an understanding of concepts in order to advance in their
-studies. In this way, the student applies an active learning strategy.
+% Another interesting study technique is spacing. Only 7.4% of the conversations
+% exhibited message delays exceeding 24 hours and 26.19% involved response times 
+% longer than 60 minutes. However, Carlos told me that this may be due to some 
+% limitation of the bot, as it only maintains a conversation while the browser 
+ % tab is open.
 
-About Metacognitive Experience, several types of feelings may be associated at
-this moment, such as actively reflecting on the answers, uncertainty and/or
-confusion when trying to understand the concept, frustration for not
-progressing, and satisfaction as their questions are answered in a way that
-facilitates understanding.
+ % There is a big difference (I haven't looked at it statistically yet) in the 
+%  interaction time with the Bot when a student asks conceptual questions. 
+%  Perhaps this actually makes sense, since it is a category where, from the 
+%  point of view of the metacognitive objective, the student is more concerned 
+%  with understanding a concept and deeper learning.
 
-Regarding the type of Metacognition related to objective of the Task, the
-students may recognize the difficult and trying to achieve a deeper
-understanding of the evolved concepts to later apply in practice.
-
-With respect to Metacognitive Strategies, the student demonstrates a sence of
-planning to identify their knowledge gaps. They may start with more general
-questions to obtain an overview of the concept and then ask more specific
-questions as they gain more clarity. Because the questions are in sequence, the
-student probably can adequately monitor their progress.
-
-\subsection*{Interleaving}
+\subsection{Interleaving}
 
@@
 
-\subsection*{Spacing}
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.52]{img/figure2.png}
+  \caption{Examples of conversations.}
+  \label{fig:graph2}
+\end{figure*}
 
-% Spacing
-Spacing, a well-established learning strategy \cite{Carvalho20}, was examined
-by analyzing the time intervals between users’ interactions with CharlieBot.
-Although only 7.4\% of the conversations exhibited message delays exceeding 24
-hours, 26.19\% involved response times longer than 60 minutes. These findings
-suggest that students typically do not extend conversations across multiple days.
-However, the data also reveal that some conversations had one or more pauses
-exceeding an hour, a practice known to enhance knowledge retention.
+% % Conceptual Questions
+% \subsubsection*{Sequence of Conceptual Question}
 
-% Another interesting study technique is spacing. Only 7.4% of the conversations
-% exhibited message delays exceeding 24 hours and 26.19% involved response times 
-% longer than 60 minutes. However, Carlos told me that this may be due to some 
-% limitation of the bot, as it only maintains a conversation while the browser 
- % tab is open.
+% Messages classified as conceptual questions represent around 32\% of all
+% messages. We have found a pattern of sequences of Conceptual Questions in some
+% conversations. Some conversations are entirely composed of messages classified
+% as Conceptual Questions. In other cases, the sequence of Conceptual questions
+% starts with a request for a Complete Solution or a Code Snippet, followed by
+% a progression of Conceptual Questions. In both cases, the student is actively
+% engaging with the bot to understand the concepts and compromised with more
+% deep learning.
 
- % There is a big difference (I haven't looked at it statistically yet) in the 
-%  interaction time with the Bot when a student asks conceptual questions. 
-%  Perhaps this actually makes sense, since it is a category where, from the 
-%  point of view of the metacognitive objective, the student is more concerned 
-%  with understanding a concept and deeper learning.
+% Regarding Metacognitive Knowledge, these sequences of prompts classified as
+% Conceptual Question, can be interpreted that the students are aware they have
+% not fully mastered a concept and are trying to understand it, demonstrating a
+% good level of self-awareness. In this sense, the student realizes that the
+% problem requires an understanding of concepts in order to advance in their
+% studies. In this way, the student applies an active learning strategy.
+
+% About Metacognitive Experience, several types of feelings may be associated at
+% this moment, such as actively reflecting on the answers, uncertainty and/or
+% confusion when trying to understand the concept, frustration for not
+% progressing, and satisfaction as their questions are answered in a way that
+% facilitates understanding.
+
+% Regarding the type of Metacognition related to objective of the Task, the
+% students may recognize the difficult and trying to achieve a deeper
+% understanding of the evolved concepts to later apply in practice.
+
+% With respect to Metacognitive Strategies, the student demonstrates a sence of
+% planning to identify their knowledge gaps. They may start with more general
+% questions to obtain an overview of the concept and then ask more specific
+% questions as they gain more clarity. Because the questions are in sequence, the
+% student probably can adequately monitor their progress.
+
+% % Multiple Choice
+% \subsubsection*{Sequence of Multiple Choice}
+
+% Around 16.9\% of the messages were classified as Multiple Choice. These prompts
+% typically involve students sending one or more exercises to CharlieBot, asking
+% for the solution.
+
+% We identified a sequence of Multiple Choice patterns in some of the
+% conversations. From the point of view of Metacognitive Knowledge, a possible
+% interpretation of this complex phenomenon is that students may believe in their
+% inability to answer several questions. They may also assume that the questions
+% share a common theme, and solving them together may help consolidate knowledge.
+% Additionally, they might understand that obtaining the answers quickly is more
+% efficient.
+
+% Regarding Metacognitive Experiences, students may experience feelings such as
+% cognitive overload, a sense of urgency, and/or confidence in ready-made answers,
+% which can often lead to relief and a reduction of frustration.
+
+% Concerning the Metacognitive about the Task, one interpretation may be that the
+% student simply wants to complete the task quickly to alleviate the workload,
+% and/or obtain a correct answer without focusing on deeper learning at
+% this moment.
+
+% In relation to Metacognitive Strategies, students are likely focusing on saving
+% time and/or possibly lacking planning skills. They may also perceive that
+% solving the questions on their own requires too much cognitive effort, but this
+% pattern might reveal a lack of progress monitoring and understanding of
+% concepts. After obtaining the answers, the metacognitive evaluation process
+% could lead to feelings of satisfaction, but further reflection may indicate that
+% the strategy was not effective in supporting deep learning. Depending on this
+% evaluation, it may influence future decision-making to change the strategy.
+
+% % Debugging help
+% \subsubsection*{Sequence of Debugging Help}
+
+% Around 30.9\% of the messages were classified as Debugging Help. Besides of
+% that, we identified a pattern of a sequences of Debugging Help in the
+% conversations. These prompts typically start with a message requesting a
+% Complete Solution or a Code Snippet. After that, the students normally chance
+% from passive to more active learning strategy.
+
+% In terms of Metacognitive Knowledge, the student probably recognizes that they
+% do not have the necessary tools or knowledge to debug the code on their own.
+% Concerning the understanding of the task, they may know that debugging is not
+% just a matter of superficial corrections but requires understanding the
+% underlying logic. Thus, the student apparently believes that the sequence of
+% questions related to debugging will help them understand the codes.
+
+% In terms of Metacognitive Experiences, some feelings may arise at this moment,
+% such as frustration or uncertainty in the face of errors, confusion during the
+% process, and also growing confidence or satisfaction as they better understand
+% the code and possible corrections.
+
+% With respect to Metacognitive related with the objective of the Task, the
+% student is probably focused on understanding the code in practical terms,
+% identifying errors, and learning how to better debug in the future.
+
+% As for Metacognitive associated with the Strategy, the student has requested
+% external help and realizes that dividing the debugging task into several
+% questions can result in a better understanding of the code. They can monitor
+% their progress throughout the process, adjusting their questions as necessary,
+% evaluate the effectiveness of the responses received, and reflect on what was
+% learned during the process.
 

### 2024-09-30 CSEDU2025/Paper.tex
update paper

@@
 
-\title{Analysis of Programming Students' Prompts with Generative AI}
+\title{
+  Temporary: An Analysis of Programming Students’ Interaction Prompts with
+  Generative AI: A Metacognitive Perspective
+}
 
@@
 
-\keywords{Generative Artificial Intelligence, Programming Education, Chatbots}
+\keywords{Generative Artificial Intelligence, Metacognition, Programming Education, Chatbots}
 
@@
 
-% %What is the problem to be solved?
-
-Recent advancements in Generative Artificial Intelligence (GenAI) have opened
-up new possibilities in education. In programming courses, students can utilize
-GenAI tools to improve their understanding, receiving assistance, feedback, and
+Recent advancements in Generative Artificial Intelligence (GenAI) have opened up
+new possibilities in education. In programming courses, students can utilize
+GenAI tools to improve their understanding, receive feedback, and obtain
 detailed explanations.
 
-The study of \cite{chan23} revealed that both undergraduate and postgraduate
-students exhibit positive attitudes toward the use of GenAI in teaching and
-learning, noting that it enhances the depth of their thinking and understanding.
-Additionally, a systematic review conducted by \cite{Lo24} demonstrated that
-students could effectively learn from ChatGPT, leading to improved understanding
-and academic achievement. Besides of that, it was also observed that ChatGPT
-enables students to regulate their learning pace, promoting self-regulation
-\citep{Baha24} \citep{cai23}.
+The study guided by \cite{chan23} revealed that both undergraduate and
+postgraduate students exhibit positive attitudes toward the use of GenAI in
+teaching and learning, noting that it enhances the depth of their thinking and
+understanding. Additionally, a systematic review conducted by \cite{Lo24}
+demonstrated that students could effectively learn from ChatGPT, leading to
+improved understanding and academic achievement. Besides that, it was also
+observed that ChatGPT enables students to regulate their learning pace
+\citep{Baha24} and can potentially support students’ self-regulated learning,
+especially for those with previous technical and disciplinary knowledge
+\citep{Xia23}.
 
-In the other hand, researchers have concerns about the impact of these
-tools on students. The systematic review of \cite{Murillo23} indicated that
+On the other hand, researchers have concerns about the impact of these tools on
+students. The systematic review conducted by \cite{Murillo23} indicated that
 ChatGPT use might lead to an overreliance on the tool. \cite{chan23} also noted
@@
 
-The student's confidence is warranted, as shown by \cite{Puryear22}, who
-demonstrated that GitHub Copilot can generate solutions for student assignments
-with accuracy rates ranging from 68\% to 95\%. However, this raises concerns that
-students might rely too heavily on GenAI tools, potentially neglecting a deeper
-understanding of the underlying concepts. \cite{cai23} further identified
-overdependence and reduced intellectual engagement as significant drawbacks of
-using ChatGPT in learning.
+The student’s confidence is justified, as demonstrated by \cite{Puryear22}, who
+found that GitHub Copilot can generate solutions for student assignments with
+accuracy rates ranging from 68\% to 95\%. However, this also raises concerns
+that students may become overly reliant on GenAI tools, potentially bypassing
+the opportunity to develop a deeper understanding of underlying concepts.
+\cite{cai23} further highlighted overdependence and diminished intellectual
+engagement as significant drawbacks associated with the use of ChatGPT in
+learning.
 
@@
 by \cite{Dickey24} revealed that at least 54.5\% of students are already using
-GenAI for homework. This highlights the need to understand how students are
-interacting with these tools and how they can be used to enhance learning.
-Furthermore, a review by \cite{Lo24} emphasized the necessity for extended
-studies and objective measures to gather more robust evidence on the use of
-GenAI tools in education.
+GenAI for homework. This highlights the need to increase the understanding of
+how students are interacting with these tools and how they can be used to
+enhance learning. Furthermore, a review organized by \cite{Lo24} emphasized the
+necessity for extended studies and objective measures to gather more robust
+evidence on the use of GenAI tools in education.
 
-% % Are there any existing solutions?
+The primary objective of this study is to analyze the interactions between
+students in a Java programming course and an educational bot powered by
+ChatGPT 3.5, enhanced with Retrieval-Augmented Generation (RAG) capabilities,
+within the context of two Metacognitive strategies: Interleaving and Spacing.
+To achieve this objective, we formulated three research questions:
 
-To deal with this issue, some researchers have proposed different pedagogical
-strategies \cite{Denny24} introduced the concept of \textit{Prompt
-Problems}, where students solve programming exercises by formulating natural
-language prompts. \cite{Prasad24} proposed a self-regulated learning framework
-using GenAI to solve programming problems. \cite{Lauren23} explored the
-integration of GenAI with evidence-based learning strategies in computer science
-and engineering courses.
-
-% % Which is the best?
-
-The study of effective educational strategies in use of GenAI tools in education
-context are important for determining the best training for educators. However,
-as these tools gain popularity among students \cite{Dickey24} the urgency to
-equip educators with best practices may lag behind their rapid adoption.
-Therefore, analyzing students' interaction patterns with GenAI tools is
-important for understanding how these tools are being used and how educators can
-leverage them to enhance learning. In addition, these interaction patterns can
-create new opportunities to redesign GenAI tools to better support pedagogical
-strategies without repressing the development of abstract, critical and creative
-thinking.
+\begin{enumerate}
+  \item What was the distribution pattern of the classified messages?
+  \item Was there a space between student prompts? If so, which category promote the largest spacing between submissions?
+  \item Do students' conversations with the bot alternate between practical and theoretical approaches, thereby promoting interleaving?
+\end{enumerate}
 
@@
 
-Nevertheless, it is important to consider how GenAI tools affect students across
-different demographic groups, knowledge areas, cultural backgrounds
-\cite{catalan21} \cite{neo22} and students' prior experience. Thus, this works
-is limited to a specific group of students and a specific GenAI tool. The
-results may not be generalizable to other populations or tools.
+However, it is crucial to consider how GenAI tools impact students across
+various demographic groups, academic disciplines, cultural backgrounds, and
+levels of prior experience \citep{catalan21} \citep{neo22}. Consequently, this
+study is limited to a specific group of students and focuses on the use of a
+single GenAI tool. As a result, the findings may not be generalizable to other
+populations or tools.
 
-% % What do you hope to achieve?
+%% Are there any existing solutions?
+% To deal with this issue, some researchers have proposed different pedagogical
+% strategies \cite{Denny24} introduced the concept of \textit{Prompt
+% Problems}, where students solve programming exercises by formulating natural
+% language prompts. \cite{Prasad24} proposed a self-regulated learning framework
+% using GenAI to solve programming problems. \cite{Lauren23} explored the
+% integration of GenAI with evidence-based learning strategies in computer science
+% and engineering courses.
 
-This study aims to analyze the interactions between students and CharlieBot,
-an ChatGPT 3.5 based bot developed by UC3M that leverages Retrieval-Augmented
-Generation (RAG) to enhance its contextual understanding of Java programming.
-We created three research questions to guide our analysis:
+%% Which is the best?
 
-RQ1: What was the distribution pattern of the classified messages?
-
-RQ2: Was there a time gap between student submissions? If so, which category
-showed the largest spacing between submissions?
-
-RQ3: Do students’ conversations with the bot alternate between practical and
-theoretical approaches, thereby promoting interleaving?
+% The study of effective educational strategies in use of GenAI tools in education
+% context are important for determining the best training for educators. However,
+% as these tools gain popularity among students \cite{Dickey24} the urgency to
+% equip educators with best practices may lag behind their rapid adoption.
+% Therefore, analyzing students' interaction patterns with GenAI tools is
+% important for understanding how these tools are being used and how educators can
+% leverage them to enhance learning. In addition, these interaction patterns can
+% create new opportunities to redesign GenAI tools to better support pedagogical
+% strategies without repressing the development of abstract, critical and creative
+% thinking.
 
@@
 \subsection{Related Work}
+
+TODO: Work in progress
 
@@
 % regulate their learning pace.
+
 \begin{table*}[htbp]
@@
 before, during, and after cognitive activity, such as frustration, confusion,
-and satisfaction, among others. Metacognitive goals are the key to regulating
+and satisfaction, among others. Metacognitive Goals are the key to regulating
 thought as they relate to the goals the individual seeks to achieve, directly
@@
 and Complete Solution. However, the data collected indicated the need for
-additional categories, resulting in the inclusion of four more: Multiple Choice
-(Quiz), Student Corrections, Language Change, and Uncategorized. Table
-\ref{tab:categories} presents these categories along with their respective
-descriptions.
+additional categories, leading to the inclusion of four more: Multiple Choice
+(Quiz), Student Corrections, Language Change, and Uncategorized. In this study,
+students were free to use the bot as they wished, which contributed to the
+emergence of these additional categories. Table \ref{tab:categories} presents
+these categories along with their respective descriptions.
 
-The analysis phase involved the use of Python scripts in the Jupyter Notebook
-environment to extract relevant information from previously classified
-data. The datasets and code snippets used in this process are available at: YYY.
+The analysis phase involved the use of Python scripts to extract information
+from previously classified data. The datasets and code snippets used in this
+process are available at: YYY.
 
@@
 
-The results of the analysis are presented in this section, addressing the
-research questions posed in the introduction.
+A total of 343 students' messages were categorized across 42 conversations, with
+an average of 8.17 messages per conversation. This data offers insights into how
+users engage with CharlieBot and the types of prompts they submit. The analysis
+results are presented in this section, addressing the research questions
+outlined in the introduction.
 
@@
 
-\subsection{Classification of Messages}
-
-The Figure \ref{fig:graph1} shows that around 5\% of the prompts
-written by students were not categorized. These messages range from
-greetings to the Bot, such as a simple "\textit{Hola}," to
-contextual statements like "\textit{Estoy repasando orientación a
-objetos.}"
-
-% Approximately 64\% of the messages pertain to categories associated with
-% critical thinking, corroborating with the findings of
-% \cite{Ghimire24}. In contrast, around 28\% of the messages
-% indicate a preference among students for ready-made answers.
-
-% Since most of the messages are located in categories where students are 
-% notoriously more active in their learning (conceptual questions 32.1% and 
-% Debugging Help 30.9%), it seems that students are trying to use the AI in a 
-% way that promotes their learning. However, the pattern of conversations with 
-% sequential blocks and also little spacing may suggest that there is a gap in 
-% knowledge of study strategies.
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.55]{img/figure3.png}
-  \caption{Interaction times per categories.}
-  \label{fig:graph3}
-\end{figure*}
-
-\subsection{Spacing}
-
-% Spacing
-Spacing, a well-established learning strategy \citep{Carvalho20}, was examined
-by analyzing the time intervals between users’ interactions with CharlieBot.
-Although only 7.4\% of the conversations exhibited message delays exceeding 24
-hours, 26.19\% involved response times longer than 60 minutes. These findings
-suggest that students typically do not extend conversations across multiple days.
-However, the data also reveal that some conversations had one or more pauses
-exceeding an hour, a practice known to enhance knowledge retention.
-
-% Another interesting study technique is spacing. Only 7.4% of the conversations
-% exhibited message delays exceeding 24 hours and 26.19% involved response times 
-% longer than 60 minutes. However, Carlos told me that this may be due to some 
-% limitation of the bot, as it only maintains a conversation while the browser 
- % tab is open.
-
- % There is a big difference (I haven't looked at it statistically yet) in the 
-%  interaction time with the Bot when a student asks conceptual questions. 
-%  Perhaps this actually makes sense, since it is a category where, from the 
-%  point of view of the metacognitive objective, the student is more concerned 
-%  with understanding a concept and deeper learning.
-
-\subsection{Interleaving}
-
-% Interleaving
-
-Intervealing is a study strategy that involves mixing or alternating different
-topics or problems within a single learning session. This
-approach has been shown to be useful for promoting long-term learning and
-improving the application of knowledge in other situations \cite{Rivers21}.
-
-When analyzing the classification of messages in conversations, we identified
-the presence of sequences of messages belonging to the same category, that it
-could suggest a low occurrence of interleaving. To better assess the extent of
-these sequences, we calculated the percentage of messages that were part of
-repeated blocks.
-
-Considering blocks with at least three messages, we observed that 56.2\% of the
-students' prompts were grouped into some sequence. When analyzing blocks with a
-minimum of 4 messages, we found that 47.5\% of the prompts were part of
-continuous sequences. These results indicate a moderate incidence of
-interleaving among the students' messages.
-
-To illustrate the presence of blocks the Figure \ref{fig:graph2} presents
-examples of conversations and their classification. In Conversation 2, the
-student alternates messages among Code Snippet, Debugging Help and Conceptual
-Question, showing an tendency of interleaving between practical and theoretical
-perspectives. In contrast, in Conversation 10, the student focused on a single
-block of Conceptual Question. Although the focus was on understanding a
-theoretical concept, this conversation tended not to promote interleaving, due
-to the absence of alternation between different types of tasks.
-
 \begin{figure*}[htbp]
@@
 
-% % Conceptual Questions
+\subsection{Classification of Messages}
+
+About the first research question: \textit{What was the distribution pattern of
+the classified messages?.}
+
+According to Figure \ref{fig:graph1} messages classified as
+Conceptual Question account for 32.1\% of the total. Often, the sequence of
+Conceptual Question begins with a request for a Complete Solution or Code
+Snippet, which then evolves into a series of Conceptual Question, as seen in
+conversation 2 of Figure \ref{fig:graph2}. In other cases, the conversation
+consists entirely of messages classified as Conceptual Question, such as in
+conversation 10 of Figure \ref{fig:graph2}. This type of message indicates that
+students recognize they have not yet fully mastered a concept and are seeking to
+better understand it. In this context, the student realizes that solving the
+problem requires understanding the concepts, thereby applying an active learning
+strategy.
+
+As shown in Figure \ref{fig:graph1}, approximately 30.9\% of the messages were
+classified as Debugging Help. These prompts often begin with a request for a
+Complete Solution or Code Snippet. Following this, students typically shift
+from a passive to a more active learning strategy, as illustrated by the start
+of conversation 12 in Figure \ref{fig:graph2}. Debugging Help messages represent
+a form of active learning, where students likely focus on gaining a practical
+understanding of the code, identifying errors, and improving their debugging
+skills for future tasks.
+
+Since most of the messages are located in categories where students are
+notoriously engaged in more active in their learning (Conceptual Question and
+Debugging Help) corroborating with the findings of \cite{Ghimire24} which showed
+that the prompts are also localized in the same categories
+
+Requests for the bot to generate a Code Snippet or Complete Solution accounted
+for 10.5\% of student messages. These prompts reflected a desire for more
+direct answers; however, it was observed that, after receiving these solutions,
+many students transitioned to a more active approach in seeking to understand
+the code or the underlying concepts. Thus, it can be inferred that prompts
+within the Code Snippet or Complete Solution category were often used as a
+starting point for more in-depth study.
+
+Multiple-choice exercise resolutions accounted for 16.9\% of the responses.
+These prompts typically involved students submitting one or more exercises to
+CharlieBot and requesting solutions. We identified sequences of multiple-choice
+questions, as illustrated by conversations 8, 9, and 15 in \ref{fig:graph2}.
+This less active learning approach can occur for various reasons, ranging from
+the belief that reviewing solutions to similar questions enhances learning, to
+issues such as negligence, poor planning, or other possible factors.
+
+Student corrections to the bot’s responses accounted for 2.6\% of the
+interactions. In most cases, these corrections occurred when the student already
+had the correct answer to an exercise and identified an error in the bot’s
+response, as illustrated in conversations 4 and 9 in Figure \ref{fig:graph2}.
+These corrections highlight that, although the bot provides quick feedback, it
+is not always accurate, and its errors can be difficult for students to
+identify.
+
+The Figure \ref{fig:graph1} also shows that around 5\% of the prompts written by
+students were not categorized. These messages range from greetings to the Bot,
+such as a simple "\textit{Hola}," to contextual statements like
+"\textit{Estoy repasando orientación a objetos.}"
+
+\subsection{Spacing}
+
+% Spacing
+
+About the second research question: \textit{Was there a space between student
+prompts? If so, which category promotes the largest spacing between
+submissions?}
+
+Spacing, a well-established learning strategy \citep{Carvalho20}, was analyzed
+by examining the time intervals between users' interactions with CharlieBot.
+Only 9\% of the conversations showed message delays exceeding 24 hours,
+indicating that students generally do not extend interactions across multiple
+days. However, this students' behavior is directly influenced by the
+implementation of the initial version of CharlieBot, which was designed to keep
+the chat session active only while the student kept the browser tab open and
+avoided reloading the page.
+
+On the other hand, about 26.2\% of the conversations, or roughly a quarter,
+included pauses longer than 60 minutes, a practice known to help with knowledge
+retention. However, this data only shows that students paused their use of the
+AI, which could mean they were reflecting on the topic independently or engaging
+in unrelated activities, making it uncertain in what context the students were
+applying spacing.
+
+There is a noticeable difference in the interaction time with the bot when a
+student asks a conceptual question. This was probably due to the fact that, in
+this category, the student is more focused on understanding a concept and
+engaging in deeper learning, which aligns with the metacognitive goal of
+promoting comprehension and long-term retention. The Figure \ref{fig:graph3}
+shows the average interaction time per category.
+
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.55]{img/figure3.png}
+  \caption{Average interaction time per category.}
+  \label{fig:graph3}
+\end{figure*}
+
+\subsection{Interleaving}
+
+% Interleaving
+
+Regarding the third research question: \textit{Do students' conversations with
+the bot alternate between practical and theoretical approaches, thereby
+promoting interleaving?}
+
+Intervealing is a learning strategy that involves mixing or alternating
+different topics or problems within a single learning session. This approach has
+been shown to be useful for promoting long-term learning and improving the
+application of knowledge in other situations \citep{Rivers21}.
+
+When analyzing the classification of messages in conversations, we identified
+the presence of sequences of messages belonging to the same category, that it
+could suggest a low occurrence of interleaving. To illustrate the presence of
+blocks the Figure \ref{fig:graph2} presents examples of conversations and their
+classification. In Conversation 2, the student alternates messages among Code
+Snippet, Debugging Help and Conceptual Question, showing an tendency of
+interleaving from practical to theoretical perspective. In contrast, in
+Conversation 10, the student focused on a single block of Conceptual Question.
+Although the focus was on understanding a theoretical concept, this conversation
+tended not to promote interleaving, due to the absence of alternation between
+different types of tasks.
+
+To better assess the extent of these sequences, we calculated the percentage of
+messages that were part of repeated blocks. Considering blocks with at least
+three messages, we observed that 56.2\% of the students' prompts were grouped
+into some sequence. When analyzing blocks with a minimum of four messages, we
+found that 47.5\% of the prompts were part of continuous sequences.
+
+However, if we consider that sequential messages from a certain point indicate
+that the student is \textit{blocked} (or focused) on a task, the remaining space
+for interleaving is 43.8\%, assuming a block starting from at least three
+consecutive messages, and 52.5\% for blocks starting from four consecutive
+messages. Thus, the data suggest that interleaving occurs on students'
+conversations with CharlieBot but in which situations and with what type of
+students profile is a study that needs to be further investigated.
+
+\section{Conclusion and Future Work}
+
+TODO: Work in progress
+
+* Although the data indicates some use of interleaving and spacing strategies,
+we believe that students can be guided to adopt a more deliberate approach in
+utilizing GenAI, enabling them to more effectively harness these tools to
+enhance their learning.
+
+* Due to the anonymity of the interactions, it's unclear whether
+students were intentionally using interleaving and spacing strategies or if
+these patterns emerged naturally. Future research should investigate whether
+students were consciously applying.
+
+* To provide teachers with valuable data for analyzing how students are using
+the tools, we are developing an AI model to automatically classify student
+messages. This will enable educators to better understand student behavior and
+offer more targeted support where needed.
+
+\section*{ACKNOWLEDGEMENTS}
+
+The authors would like to thank the Federal Institute of Education, Science and
+Technology (IFRS) for the partial financial support provided for the execution
+of this research.
+
+\bibliographystyle{apalike}
+{\small
+\bibliography{References}}
+
+\section*{\uppercase{\textit{Appendix}}}
+
+If any, the appendix should appear directly after the
+references without numbering, and not on a new page. To do so please use the
+following command: \textit{$\backslash$section*\{APPENDIX\}}
+
+\end{document}
+
+% % Conceptual Question
 % \subsubsection*{Sequence of Conceptual Question}
 
-% Messages classified as conceptual questions represent around 32\% of all
-% messages. We have found a pattern of sequences of Conceptual Questions in some
+% Messages classified as conceptual question represent around 32\% of all
+% messages. We have found a pattern of sequences of Conceptual Question in some
 % conversations. Some conversations are entirely composed of messages classified
-% as Conceptual Questions. In other cases, the sequence of Conceptual questions
+% as Conceptual Question. In other cases, the sequence of Conceptual question
 % starts with a request for a Complete Solution or a Code Snippet, followed by
-% a progression of Conceptual Questions. In both cases, the student is actively
+% a progression of Conceptual Question. In both cases, the student is actively
 % engaging with the bot to understand the concepts and compromised with more
@@
 % learned during the process.
-
-\section{Conclusion}
-
-The results of this study suggest that students interact with CharlieBot in a
-manner that promotes critical thinking. The majority of the messages analyzed
-were related to debugging help, conceptual questions, and student correction,
-indicating that students are actively engaging with the bot to understand
-concepts and solve problems. This is a positive outcome, as it suggests that
-CharlieBot is being used as a tool to enhance learning rather than a crutch for
-students to rely on.
-
-\section*{ACKNOWLEDGEMENTS}
-
-The authors would like to thank the Federal Institute of Education, Science and
-Technology (IFRS) for the partial financial support provided for the execution
-of this research.
-
-\bibliographystyle{apalike}
-{\small
-\bibliography{References}}
-
-\section*{\uppercase{Appendix}}
-
-If any, the appendix should appear directly after the
-references without numbering, and not on a new page. To do so please use the
-following command: \textit{$\backslash$section*\{APPENDIX\}}
-
-\end{document}

### 2024-09-30 CSEDU2025/Paper.tex
update

@@
 The study guided by \cite{chan23} revealed that both undergraduate and
-postgraduate students exhibit positive attitudes toward the use of GenAI in
-teaching and learning, noting that it enhances the depth of their thinking and
-understanding. Additionally, a systematic review conducted by \cite{Lo24}
+postgraduate students exhibit positive attitudes towards the use of GenAI in
+teaching and learning, noting that it improves the depth of their thinking and
+understanding. Furthermore, a systematic review conducted by \cite{Lo24}
 demonstrated that students could effectively learn from ChatGPT, leading to
@@
 
-On the other hand, researchers have concerns about the impact of these tools on
+However, researchers are concerned about the impact of these tools on
 students. The systematic review conducted by \cite{Murillo23} indicated that
-ChatGPT use might lead to an overreliance on the tool. \cite{chan23} also noted
-that this reliance could result in a decrease in critical thinking, as students
-might make decisions based solely on the information provided by ChatGPT.
+the use of ChatGPT could lead to an overreliance on the tool. \cite{chan23} also noted
+that this dependency could result in a decrease in critical thinking, as students
+could make decisions based solely on the information provided by ChatGPT.
 
@@
 accuracy rates ranging from 68\% to 95\%. However, this also raises concerns
-that students may become overly reliant on GenAI tools, potentially bypassing
-the opportunity to develop a deeper understanding of underlying concepts.
-\cite{cai23} further highlighted overdependence and diminished intellectual
+that students may become overly dependent on GenAI tools, potentially bypassing
+the opportunity to develop a deeper understanding of the underlying concepts.
+\cite{cai23} further highlighted the overdependence and diminished intellectual
 engagement as significant drawbacks associated with the use of ChatGPT in
@@
 GenAI for homework. This highlights the need to increase the understanding of
-how students are interacting with these tools and how they can be used to
+how students interact with these tools and how they can be used to
 enhance learning. Furthermore, a review organized by \cite{Lo24} emphasized the
-necessity for extended studies and objective measures to gather more robust
+need for extended studies and objective measures to gather more robust
 evidence on the use of GenAI tools in education.
@@
 within the context of two Metacognitive strategies: Interleaving and Spacing.
-To achieve this objective, we formulated three research questions:
+To achieve this objective, we formulate three research questions:
 
@@
 
-However, it is crucial to consider how GenAI tools impact students across
+However, it is crucial to consider how GenAI tools impact students in
 various demographic groups, academic disciplines, cultural backgrounds, and
-levels of prior experience \citep{catalan21} \citep{neo22}. Consequently, this
+levels of previous experience \citep{catalan21} \citep{neo22}. Consequently, this
 study is limited to a specific group of students and focuses on the use of a
@@
 
-TODO: Work in progress
+TODO: Work in progress.
 
@@
 About the first research question: \textit{What was the distribution pattern of
-the classified messages?.}
+the classified messages?}
 
@@
 better understand it. In this context, the student realizes that solving the
-problem requires understanding the concepts, thereby applying an active learning
+problem requires understanding of concepts, thus applying an active learning
 strategy.
@@
 classified as Debugging Help. These prompts often begin with a request for a
-Complete Solution or Code Snippet. Following this, students typically shift
-from a passive to a more active learning strategy, as illustrated by the start
+Complete Solution or Code Snippet. Following this, students typically switch from
+passive to more active learning strategy, as illustrated by the start
 of conversation 12 in Figure \ref{fig:graph2}. Debugging Help messages represent
@@
 Since most of the messages are located in categories where students are
-notoriously engaged in more active in their learning (Conceptual Question and
-Debugging Help) corroborating with the findings of \cite{Ghimire24} which showed
-that the prompts are also localized in the same categories
+notoriously engaged in being more active in their learning (conceptual question 
+and problem solving help), corroborating the findings of \cite{Ghimire24} which 
+showed that the questions are also localized in the same categories.
 
@@
 many students transitioned to a more active approach in seeking to understand
-the code or the underlying concepts. Thus, it can be inferred that prompts
+the code or the underlying concepts. Thus, it can be inferred that the prompts
 within the Code Snippet or Complete Solution category were often used as a
-starting point for more in-depth study.
+starting point for a more in-depth study.
 
-Multiple-choice exercise resolutions accounted for 16.9\% of the responses.
+Multiple Choice exercise resolutions represented 16.9\% of the responses.
 These prompts typically involved students submitting one or more exercises to
-CharlieBot and requesting solutions. We identified sequences of multiple-choice
+CharlieBot and requesting solutions. We identified sequences of multiple choice
 questions, as illustrated by conversations 8, 9, and 15 in \ref{fig:graph2}.
@@
 
-Student corrections to the bot’s responses accounted for 2.6\% of the
+Student corrections to bot responses accounted for 2.6\% of the
 interactions. In most cases, these corrections occurred when the student already

### 2024-10-01 CSEDU2025/Paper.tex

@@
 \title{
-  Temporary: An Analysis of Programming Students’ Interaction Prompts with
-  Generative AI: A Metacognitive Perspective
+  Examining Programming Students’ Interaction with Generative AI: A
+  Metacognitive Perspective
 }
@@
 Patricia Callejo Pinardo\sup{2}\orcidAuthor{0000-0001-6124-6213},
-Iria Manuela Estévez Ayres\sup{2}\orcidAuthor{0000-0002-1047-5398},
+Iria Estévez-Ayres\sup{2}\orcidAuthor{0000-0002-1047-5398},
 Carlos Delgado Kloos\sup{2}\orcidAuthor{0000-0003-4093-3705}}
-\affiliation{\sup{1}Department of Informatics, Federal Institute of Education,
-Science and Technology, Porto Alegre, Brazil}
-\affiliation{\sup{2}Department of Telematics Engineering, Carlos III University,
-Madrid, Spain}
+\affiliation{\sup{1}Department of Informatics, Instituto Federal de Educação,
+Ciência e Tecnologia, Porto Alegre, Brazil}
+\affiliation{\sup{2}Department of Telematics Engineering, Universidad Carlos III
+de Madrid, Leganés (Madrid), Spain}
 \email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario,pcallejo, ayres,cdk\}@it.uc3m.es}
@@
 
-\abstract{The abstract should summarize the contents of the paper and should
-contain at least 70 and at most 200 words. The text must be set to 9-point font
-size.}
+\abstract{This study investigates the interaction patterns of programming
+students with a generative AI-powered chatbot (CharlieBot) in the context of a
+Java programming course. Specifically, it explores how students utilize the bot,
+enhanced with Retrieval-Augmented Generation (RAG), along with two metacognitive
+strategies: interleavBaha24Sun24ing and spacing. The methodology analyzed and
+classified student conversations, focusing on types of interactions such as
+conceptual questions and debugging help. The results indicate that most
+interactions are focused on active learning strategies, although the deliberate use
+of interleaving and spacing is limited. The study highlights the need for more
+guidance on the strategic use of AI tools to improve learning outcomes.}
 
@@
 
-Recent advancements in Generative Artificial Intelligence (GenAI) have opened up
+Recent advances in Generative Artificial Intelligence (GenAI) have opened up
 new possibilities in education. In programming courses, students can utilize
@@
 students. The systematic review conducted by \cite{Murillo23} indicated that
-the use of ChatGPT could lead to an overreliance on the tool. \cite{chan23} also noted
-that this dependency could result in a decrease in critical thinking, as students
-could make decisions based solely on the information provided by ChatGPT.
+the use of ChatGPT could lead to an overreliance on the tool. \cite{chan23} also
+noted that this dependency could result in a decrease in critical thinking, as
+students could make decisions based solely on the information provided by
+ChatGPT.
 
@@
 
-Regardless of teachers' preferences or beliefs, a preliminary surveys performed
+Regardless of teachers' preferences or beliefs, preliminary surveys conducted
 by \cite{Dickey24} revealed that at least 54.5\% of students are already using
@@
   \item What was the distribution pattern of the classified messages?
-  \item Was there a space between student prompts? If so, which category promote the largest spacing between submissions?
+  \item Was there a space between the student prompts? If so, which category promotes the largest spacing between submissions?
   \item Do students' conversations with the bot alternate between practical and theoretical approaches, thereby promoting interleaving?
@@
 (Quiz), Student Corrections, Language Change, and Uncategorized. In this study,
-students were free to use the bot as they wished, which contributed to the
+students were free to use the bot as they wanted, which contributed to the
 emergence of these additional categories. Table \ref{tab:categories} presents
@@
 
-A total of 343 students' messages were categorized across 42 conversations, with
+A total of 343 student messages were categorized in 42 conversations, with
 an average of 8.17 messages per conversation. This data offers insights into how
-users engage with CharlieBot and the types of prompts they submit. The analysis
-results are presented in this section, addressing the research questions
+users engage with CharlieBot and the types of prompts they submit. The results of
+the analysis are presented in this section, addressing the research questions
 outlined in the introduction.
 
-The Figure \ref{fig:graph1} shows the distribution of the messages in the
+Figure \ref{fig:graph1} shows the distribution of the messages in the
 categories:
@@
 conversation 10 of Figure \ref{fig:graph2}. This type of message indicates that
-students recognize they have not yet fully mastered a concept and are seeking to
+students recognize that they have not yet fully mastered a concept and are seeking to
 better understand it. In this context, the student realizes that solving the
@@
 of conversation 12 in Figure \ref{fig:graph2}. Debugging Help messages represent
-a form of active learning, where students likely focus on gaining a practical
+a form of active learning, in which students likely focus on gaining a practical
 understanding of the code, identifying errors, and improving their debugging
@@
 interactions. In most cases, these corrections occurred when the student already
-had the correct answer to an exercise and identified an error in the bot’s
+had the correct answer to an exercise and identified an error in the bot
 response, as illustrated in conversations 4 and 9 in Figure \ref{fig:graph2}.
 These corrections highlight that, although the bot provides quick feedback, it
-is not always accurate, and its errors can be difficult for students to
+is not always accurate and its errors can be difficult for students to
 identify.
 
-The Figure \ref{fig:graph1} also shows that around 5\% of the prompts written by
+Figure \ref{fig:graph1} also shows that around 5\% of the prompts written by
 students were not categorized. These messages range from greetings to the Bot,
@@
 
-About the second research question: \textit{Was there a space between student
+About the second research question: \textit{Was there a space between the student
 prompts? If so, which category promotes the largest spacing between
@@
 Spacing, a well-established learning strategy \citep{Carvalho20}, was analyzed
-by examining the time intervals between users' interactions with CharlieBot.
-Only 9\% of the conversations showed message delays exceeding 24 hours,
-indicating that students generally do not extend interactions across multiple
-days. However, this students' behavior is directly influenced by the
-implementation of the initial version of CharlieBot, which was designed to keep
-the chat session active only while the student kept the browser tab open and
-avoided reloading the page.
-
-On the other hand, about 26.2\% of the conversations, or roughly a quarter,
-included pauses longer than 60 minutes, a practice known to help with knowledge
-retention. However, this data only shows that students paused their use of the
-AI, which could mean they were reflecting on the topic independently or engaging
-in unrelated activities, making it uncertain in what context the students were
-applying spacing.
+by examining the time intervals between user interactions with CharlieBot.
+Approximately 26.2\% of the conversations, or roughly a quarter, included pauses
+longer than 60 minutes. However, this data only indicates that students paused
+their use of the AI, which could suggest that they were either reflecting on the
+topic independently or engaging in unrelated activities. Furthermore, 16.7\%
+of the conversations exhibited pauses of 12 hours and 9\% showed delays
+exceeding 24 hours between messages, indicating that some degree of spacing
+occurred. However, it is important to note that this behavior may have been
+influenced by the design of the initial version of CharlieBot, which only kept
+the chat session active as long as the student kept the browser tab open and
+avoided reloading the page. This adds uncertainty about the precise context in
+which students applied spacing.
 
@@
 engaging in deeper learning, which aligns with the metacognitive goal of
-promoting comprehension and long-term retention. The Figure \ref{fig:graph3}
+promoting comprehension and long-term retention. Figure \ref{fig:graph3}
 shows the average interaction time per category.
@@
 When analyzing the classification of messages in conversations, we identified
-the presence of sequences of messages belonging to the same category, that it
+the presence of sequences of messages belonging to the same category, which
 could suggest a low occurrence of interleaving. To illustrate the presence of
-blocks the Figure \ref{fig:graph2} presents examples of conversations and their
+blocks, Figure \ref{fig:graph2} presents examples of conversations and their
 classification. In Conversation 2, the student alternates messages among Code
-Snippet, Debugging Help and Conceptual Question, showing an tendency of
-interleaving from practical to theoretical perspective. In contrast, in
+Snippet, Debugging Help, and Conceptual Question, showing a tendency to
+interleave from a practical to theoretical perspective. In contrast, in
 Conversation 10, the student focused on a single block of Conceptual Question.
@@
 tended not to promote interleaving, due to the absence of alternation between
-different types of tasks.
+different types of task.
 
@@
 messages that were part of repeated blocks. Considering blocks with at least
-three messages, we observed that 56.2\% of the students' prompts were grouped
+three messages, we observed that 56.2\% of the student's prompts were grouped
 into some sequence. When analyzing blocks with a minimum of four messages, we
@@
 messages. Thus, the data suggest that interleaving occurs on students'
-conversations with CharlieBot but in which situations and with what type of
-students profile is a study that needs to be further investigated.
+conversations with CharlieBot, but in which situations and with what type of
+student profile is a study that needs to be further investigated.
 
@@
 
-* Although the data indicates some use of interleaving and spacing strategies,
-we believe that students can be guided to adopt a more deliberate approach in
+* The results indicate that most interactions are focused on active learning
+strategies, although deliberate use of interleaving and spacing is limited. We
+believe that students can be guided to adopt a more deliberate approach in
 utilizing GenAI, enabling them to more effectively harness these tools to

### 2024-10-03 CSEDU2025/Paper.tex
update paper

@@
 
-\keywords{Generative Artificial Intelligence, Metacognition, Programming Education, Chatbots}
+\keywords{
+  Programming Education, Generative Artificial Intelligence, Chatbots and
+  Metacognition
+}
 
-\abstract{This study investigates the interaction patterns of programming
-students with a generative AI-powered chatbot (CharlieBot) in the context of a
-Java programming course. Specifically, it explores how students utilize the bot,
-enhanced with Retrieval-Augmented Generation (RAG), along with two metacognitive
-strategies: interleavBaha24Sun24ing and spacing. The methodology analyzed and
-classified student conversations, focusing on types of interactions such as
-conceptual questions and debugging help. The results indicate that most
-interactions are focused on active learning strategies, although the deliberate use
-of interleaving and spacing is limited. The study highlights the need for more
-guidance on the strategic use of AI tools to improve learning outcomes.}
+\abstract{
+This study examines the interactions of programming students with a chatbot,
+CharlieBot, powered by OpenAI's GPT-3.5 and enhanced with the
+Retrieval-Augmented Generation (RAG) technique, within the context of a Java
+programming course. It investigates how students use the bot while employing two
+metacognitive strategies: interleaving and spacing. Using the logs of student
+interactions with the bot, the methodology involved classifying and analyzing
+student prompts into eight distinct categories. The findings reveal that while
+most interactions emphasize active learning strategies, students demonstrate
+limited conscious application of interleaving and spacing. This study
+underscores the necessity for increased guidance on the strategic use of AI
+tools to enhance learning outcomes.
+}
 
@@
 
-Recent advances in Generative Artificial Intelligence (GenAI) have opened up
-new possibilities in education. In programming courses, students can utilize
-GenAI tools to improve their understanding, receive feedback, and obtain
-detailed explanations.
+Recent advancements in Generative Artificial Intelligence (GenAI) have created
+new opportunities in both professional programming \citep{Peng23}
+\citep{Pandey24} and education \citep{Puryear22}. In programming courses,
+students can employ GenAI tools to enhance their understanding, receive
+personalized feedback, and access detailed explanations. For example, GitHub
+Copilot, a tool that is integrated into development environments, assists by
+providing real-time suggestions and accelerating code writing, which could help
+streamline the learning process \citep{Denny23}. Educational chatbots, on the
+other hand, try to guide students through coding challenges, answer questions,
+and attempt to foster a deeper comprehension of programming concepts
+\citep{Labadze23}.
 
@@
 postgraduate students exhibit positive attitudes towards the use of GenAI in
-teaching and learning, noting that it improves the depth of their thinking and
-understanding. Furthermore, a systematic review conducted by \cite{Lo24}
+teaching and learning, noting that it may improve the depth of their thinking
+and understanding. Furthermore, a systematic review conducted by \cite{Lo24}
 demonstrated that students could effectively learn from ChatGPT, leading to
-improved understanding and academic achievement. Besides that, it was also
-observed that ChatGPT enables students to regulate their learning pace
-\citep{Baha24} and can potentially support students’ self-regulated learning,
-especially for those with previous technical and disciplinary knowledge
-\citep{Xia23}.
+improved understanding and academic achievement \citep{Callejo24}. Besides that,
+it was also observed that ChatGPT enables students to regulate their learning
+pace \citep{Baha24} and can potentially support students’ self-regulated
+learning, especially for those with previous technical and disciplinary
+knowledge \citep{Xia23}.
 
@@
 
-The primary objective of this study is to analyze the interactions between
-students in a Java programming course and an educational bot powered by
-ChatGPT 3.5, enhanced with Retrieval-Augmented Generation (RAG) capabilities,
-within the context of two Metacognitive strategies: Interleaving and Spacing.
-To achieve this objective, we formulate three research questions:
+In response to the growing need for deeper insights into the use of GenAI tools
+in educational settings, this study aims to help address this gap by analyzing
+the interactions between students in a Java programming course and an
+educational bot powered by gpt-3.5 from OpenAI, enhanced with
+Retrieval-Augmented Generation (RAG) technique. Specifically, it focuses on
+examining these interactions within the context of the metacognitive strategies
+employed by the students. To achieve this objective, we formulate three research
+questions:
 
-\begin{enumerate}
-  \item What was the distribution pattern of the classified messages?
-  \item Was there a space between the student prompts? If so, which category promotes the largest spacing between submissions?
-  \item Do students' conversations with the bot alternate between practical and theoretical approaches, thereby promoting interleaving?
-\end{enumerate}
+\begin{itemize}
+  \item RQ1 - What was the distribution pattern of the classified messages?
+  \item RQ2 - Was there a space between the student prompts? If so, which
+  category promotes the largest spacing between submissions?
+  \item RQ3 - Do students' conversations with the bot alternate between
+  practical and theoretical approaches, thereby promoting interleaving?
+\end{itemize}
 
@@
 
-However, it is crucial to consider how GenAI tools impact students in
+It is important to consider how GenAI tools impact students in
 various demographic groups, academic disciplines, cultural backgrounds, and
-levels of previous experience \citep{catalan21} \citep{neo22}. Consequently, this
-study is limited to a specific group of students and focuses on the use of a
-single GenAI tool. As a result, the findings may not be generalizable to other
+levels of previous experience \citep{catalan21} \citep{neo22}. Consequently,
+this study is limited to a specific group of students and focuses on the use of
+a single GenAI tool. As a result, the findings may not be generalizable to other
 populations or tools.
+
+This paper is organized as follows: Section 2 presents the theoretical
+framework, Section 3 describes the material and methods used in the study,
+Section 4 presents the results and discussion, and Section 5 concludes the
+paper and outlines future work.
 
@@
 
-\subsection{Related Work}
-
-TODO: Work in progress.
-
-% The advancements in GenAI and its use in programming courses have raised
-% possibilities and concerns among educators and researchers. The study of
-% \cite{Yilmaz23} found that the main benefit of using ChatGPT in programming
-% learning is to provide providing fast and mostly correct answers to questions,
-% improving thinking skills, facilitating debugging, and increasing
-% self-confidence. In contrast, the study the same appoint limitations such ask
-% laziness, being unable to answer some questions, or giving incomplete/incorrect
-% answers, causing professional anxiety in students.
-
-% From a cognitive perspective, \cite{Lo24} shown that students were able to
-% learn from ChatGPT, which increased their understanding and achievement.
-% However, concerns were raised that the growing use of AI tools might lead to a
-% decline in critical thinking among students. Teachers and stakeholders should
-% continue to investigate pedagogical approaches that leverage ChatGPT to enhance
-% students' understanding and critical thinking. For example, teachers can
-% instruct students to fact-check and validate information produced by ChatGPT.
-
-% Examined students from different demographic groups using AI tools like
-% ChatGPT, comparing perceptions and impacts among these groups. The results
-% showed that AI tools had a positive impact on academic performance, especially
-% for students with less prior experience.
-
-% Critical thinking is a fundamental skill for students to develop, as it enables
-% them to analyze, evaluate, and synthesize information. This skill is essential
-% for problem-solving and decision-making, making it a crucial aspect of
-% educational success, specifically in programming courses. However, the use of AI
-% tools like ChatGPT has raised concerns about the potential impact on students'
-% critical thinking abilities \cite{Murillo23} \cite{cai23} \cite{chan23}.
-
-% In the other hand, the study of \cite{zhang24} suggested that incorporating
-% ChatGPT could stimulate students' critical thinking, probably because they
-% had to assess the correctness of ChatGPT's answers.
-
-% Regarding self-regulation, \cite{cai23} noted that ChatGPT allowed students to
-% regulate their learning pace.
+This section presents the theoretical framework that underpins the study. First,
+we introduce the concept of metacognition used in this study. Next, we discuss
+related work that explores the use of metacognitive strategies and scaffolds
+in educational settings.
 
@@
 
-The use of generative artificial intelligence tools can support the important
-mechanism of learning regulation by students \citep{larsen18}, since these
-technologies frequently generate automatic and often accurate responses
-\cite{Puryear22}, it becomes essential for students to use these tools more
-strategically to promote conscious and active learning.
-
-The Theory of Metacognition addresses awareness of mental processes.
-\cite{flavell79} proposed a model of metacognitive monitoring that includes four
-interrelated phenomena: Metacognitive Knowledge, Metacognitive Experience,
- Metacognitive Goals, and Metacognitive Actions. These processes do not occur
-in isolation, but they influence each other, altering cognitive progress over
-time.
+Given that generative AI often produces automatic and highly accurate responses
+\citep{Puryear22}, it is essential for students to use these tools within an
+active learning process to enhance their learning outcomes. This active
+learning process can be further understood through the lens of Metacognition, which
+focuses on the awareness of one’s mental processes. \cite{flavell79} proposed a
+model of metacognitive monitoring that includes four interrelated phenomena:
+Metacognitive Knowledge, Metacognitive Experience, Metacognitive Goals, and
+Metacognitive Actions. These processes do not occur in isolation, but they
+influence each other, altering cognitive progress over time.
 
@@
 
+Spacing and interleaving are two metacognitive strategies that have been shown to
+enhance learning outcomes. Spacing refers to the practice of distributing study
+sessions over time, which has been shown to improve long-term retention and
+understanding of the material \citep{Carvalho20}. Interleaving involves mixing
+different topics or problems within a single learning session, which has been
+shown to enhance long-term learning and the application of knowledge in other
+situations \citep{Rivers21}.
+
+\subsection{Related Work}
+
+Studies on the use of metacognitive strategies indicate that these practices can
+be applied to teaching the conscious use of generative AI. \cite{Zheng19}
+investigated the effects of metacognitive scaffolds on group behavior,
+performance, and cognitive load in computer-supported collaborative
+environments. The results showed positive impacts on metacognitive
+behavior and group performance without increasing cognitive load.
+
+\cite{LiWei23} proposed a metacognition-based collaborative approach to enhance
+performance in collaborative programming. The findings indicated that this
+approach influenced computational thinking, critical thinking, and
+metacognitive awareness.
+
+Similarly, \cite{Wang23} explored how metacognitive instruction affects
+computational thinking, critical thinking, and metacognitive skills in
+collaborative programming, observing improvements in problem-solving and
+collaboration.
+
+These studies demonstrate the relevance of metacognition in enhancing
+performance and skills in collaborative environments without imposing additional
+cognitive load. Based on this, it is possible to suggest that metacognitive
+strategies may also be useful in teaching a more conscious and critical use of
+generative AI tools.
+
 \section{\uppercase{Material and Method}}
@@
 CharlieBot is an educational bot built on ChatGPT 3.5 and enhanced with
-Retrieval-Augmented Generation (RAG) capabilities to improve the precision of
+Retrieval-Augmented Generation (RAG) technique to improve the precision of
 responses provided to students \cite{Sun24}. The RAG feature was developed using
@@
 Since most of the messages are located in categories where students are
-notoriously engaged in being more active in their learning (conceptual question
-and problem solving help), corroborating the findings of \cite{Ghimire24} which
+notoriously engaged in being more active in their learning (Conceptual Question
+and Debugging Help), corroborating the findings of \cite{Ghimire24} which
 showed that the questions are also localized in the same categories.

### 2024-10-04 CSEDU2025/Paper.tex
update paper, related work

@@
 
+\cite{Margulieux24} conducted a study on how undergraduate students in
+introductory programming courses used generative AI to solve programming
+problems in a naturalistic setting. The research focused on examining the
+relationship between AI usage and students’ self-regulation strategies,
+self-efficacy, and fear of failure in programming. Additionally, the study
+explored how these variables interacted with learner characteristics, perceived
+usefulness of AI, and academic performance. Findings revealed that students
+with higher self-efficacy, lower fear of failure, or higher prior grades
+tended to use AI less frequently or later in the problem-solving process and
+perceived it as less useful compared to their peers. However, no significant
+relationship was found between students’ self-regulation strategies and their
+use of AI.
+
+\cite{Khusnul24} investigates the efficacy of a meta-learning approach in
+improving metacognitive and creative skills. This study confirms the
+effectiveness of meta-learning strategies and elucidates the relationship
+between meta-learning and metacognition. This research suggests that
+meta-learning can improve metacognitive abilities, providing valuable insights
+into educational technology and course design in higher education settings.
+
+Reconceptualizing AI Literacy: The Importance of Metacognitive Thinking in an Artificial Intelligence (AI)-Enabled Workforce
+
 \section{\uppercase{Material and Method}}
@@
 
-About the first research question: \textit{What was the distribution pattern of
-the classified messages?}
+About the first research question: \textit{RQ1: What was the distribution
+pattern of the classified messages?}
 
@@
 Snippet, which then evolves into a series of Conceptual Question, as seen in
-conversation 2 of Figure \ref{fig:graph2}. In other cases, the conversation
+conversation 2 of Figure \ref{fig:graph2}, which illustrates some examples of
+students' conversations with the bot. In other cases, the conversation
 consists entirely of messages classified as Conceptual Question, such as in
 conversation 10 of Figure \ref{fig:graph2}. This type of message indicates that
-students recognize that they have not yet fully mastered a concept and are seeking to
-better understand it. In this context, the student realizes that solving the
-problem requires understanding of concepts, thus applying an active learning
-strategy.
+students recognize that they have not yet fully mastered a concept and are
+seeking to better understand it. In this context, the student realizes that
+solving the problem requires understanding of concepts, thus applying an active
+learning strategy.
 
@@
 students were not categorized. These messages range from greetings to the Bot,
-such as a simple "\textit{Hola}," to contextual statements like
-"\textit{Estoy repasando orientación a objetos.}"
+such as a simple "\textit{Hello}," to contextual statements like
+"\textit{I'm reviewing object orientation}"
 
@@
 
-About the second research question: \textit{Was there a space between the student
-prompts? If so, which category promotes the largest spacing between
+About the second research question: \textit{RQ2: Was there a space between the
+student prompts? If so, which category promotes the largest spacing between
 submissions?}
@@
 
-Regarding the third research question: \textit{Do students' conversations with
-the bot alternate between practical and theoretical approaches, thereby
+Regarding the third research question: \textit{RQ3: Do students' conversations
+with the bot alternate between practical and theoretical approaches, thereby
 promoting interleaving?}
@@
 strategies, although deliberate use of interleaving and spacing is limited. We
-believe that students can be guided to adopt a more deliberate approach in
-utilizing GenAI, enabling them to more effectively harness these tools to
-enhance their learning.
+believe that students can be guided \cite{Zheng19} \cite{Wang23} \cite{LiWei23}
+\cite{Khusnul24} to adopt a more deliberate approach in utilizing GenAI,
+enabling them to more effectively harness these tools to enhance their learning.
 

### 2024-10-09 CSEDU2025/Paper.tex
perspective interleaving

@@
 limited conscious application of interleaving and spacing. This study
-underscores the necessity for increased guidance on the strategic use of AI
-tools to enhance learning outcomes.
+underscores the need for more guidance on the strategic use of AI
+tools to improve learning outcomes.
 }
@@
 
-Recent advancements in Generative Artificial Intelligence (GenAI) have created
+Recent advances in Generative Artificial Intelligence (GenAI) have created
 new opportunities in both professional programming \citep{Peng23}
 \citep{Pandey24} and education \citep{Puryear22}. In programming courses,
-students can employ GenAI tools to enhance their understanding, receive
+students can employ GenAI tools to improve their understanding, receive
 personalized feedback, and access detailed explanations. For example, GitHub
-Copilot, a tool that is integrated into development environments, assists by
+Copilot, a tool that is integrated into development environments, assists in
 providing real-time suggestions and accelerating code writing, which could help
@@
 other hand, try to guide students through coding challenges, answer questions,
-and attempt to foster a deeper comprehension of programming concepts
+and try to foster a deeper understanding of programming concepts
 \citep{Labadze23}.
 
-The study guided by \cite{chan23} revealed that both undergraduate and
-postgraduate students exhibit positive attitudes towards the use of GenAI in
-teaching and learning, noting that it may improve the depth of their thinking
-and understanding. Furthermore, a systematic review conducted by \cite{Lo24}
-demonstrated that students could effectively learn from ChatGPT, leading to
-improved understanding and academic achievement \citep{Callejo24}. Besides that,
-it was also observed that ChatGPT enables students to regulate their learning
-pace \citep{Baha24} and can potentially support students’ self-regulated
-learning, especially for those with previous technical and disciplinary
-knowledge \citep{Xia23}.
+The study conducted by \cite{chan23} revealed that both undergraduate and
+postgraduate students have positive attitudes towards the use of GenAI.
+Furthermore, a systematic review organized by \cite{Lo24} demonstrated that
+students can effectively learn from ChatGPT, resulting in improved comprehension
+and academic achievement \citep{Callejo24}. Furthermore, it was observed that
+ChatGPT allows students to regulate their learning pace \citep{Baha24} and can
+potentially support self-regulated learning, particularly for students with
+prior technical and disciplinary knowledge \citep{Xia23}.
 
-However, researchers are concerned about the impact of these tools on
-students. The systematic review conducted by \cite{Murillo23} indicated that
-the use of ChatGPT could lead to an overreliance on the tool. \cite{chan23} also
-noted that this dependency could result in a decrease in critical thinking, as
-students could make decisions based solely on the information provided by
-ChatGPT.
+However, researchers have also expressed concerns about the impact of these
+tools on students. The systematic review by \cite{Murillo23} indicated that the
+use of ChatGPT could lead to overreliance on the tool. \cite{chan23} noted that
+overdependence on ChatGPT could lead to a decrease in critical thinking, as
+students can make decisions based solely on the information provided by the tool.
+As a potential consequence, \cite{Eason23} observed in a study that when
+programming students lost access to ChatGPT-4, those who had previously relied
+on it saw their performance drop by 17\%. In contrast, students who had never
+used the tool were unaffected and outperformed their peers.
 
-The student’s confidence is justified, as demonstrated by \cite{Puryear22}, who
-found that GitHub Copilot can generate solutions for student assignments with
-accuracy rates ranging from 68\% to 95\%. However, this also raises concerns
-that students may become overly dependent on GenAI tools, potentially bypassing
-the opportunity to develop a deeper understanding of the underlying concepts.
-\cite{cai23} further highlighted the overdependence and diminished intellectual
-engagement as significant drawbacks associated with the use of ChatGPT in
-learning.
+The confidence of students in these tools is well founded, as shown by
+\cite{Puryear22}, who found that GitHub Copilot can generate solutions for
+student assignments with precision rates ranging from 68\% to 95\%. However,
+\cite{Boudouaia24} observed that using ChatGPT without a structured learning
+approach did not provide significant improvement over traditional self-directed
+learning methods in programming. This raises concerns that students may become
+overly dependent on generative AI tools, potentially hindering their learning
+progress and missing opportunities to develop a deeper understanding of
+fundamental concepts.
 
 Regardless of teachers' preferences or beliefs, preliminary surveys conducted
-by \cite{Dickey24} revealed that at least 54.5\% of students are already using
-GenAI for homework. This highlights the need to increase the understanding of
-how students interact with these tools and how they can be used to
-enhance learning. Furthermore, a review organized by \cite{Lo24} emphasized the
-need for extended studies and objective measures to gather more robust
-evidence on the use of GenAI tools in education.
+by \cite{Dickey24} indicate that more than 54,5\% of students are already using
+GenAI for homework, likely due to their perception of this technology as
+advantageous and intuitive \citep{Boudouaia24}. All of these studies highlight
+the need to increase the understanding of how students interact with these tools
+and how they can be used to improve learning. Furthermore, a review organized by
+\cite{Lo24} emphasized the need for extended studies and objective measures to
+gather more robust evidence on the use of GenAI tools in education.
 
-In response to the growing need for deeper insights into the use of GenAI tools
+In response to the growing need for deeper insight into the use of GenAI tools
 in educational settings, this study aims to help address this gap by analyzing
 the interactions between students in a Java programming course and an
-educational bot powered by gpt-3.5 from OpenAI, enhanced with
+educational bot powered by gpt-3.5 from OpenAI, enhanced with the
 Retrieval-Augmented Generation (RAG) technique. Specifically, it focuses on
 examining these interactions within the context of the metacognitive strategies
-employed by the students. To achieve this objective, we formulate three research
+employed by students. To achieve this objective, we formulate three research
 questions:
@@
 
-% % What is its main limitation?
+% What is its main limitation?
 
@@
 active learning process to enhance their learning outcomes. This active
-learning process can be further understood through the lens of Metacognition, which
-focuses on the awareness of one’s mental processes. \cite{flavell79} proposed a
-model of metacognitive monitoring that includes four interrelated phenomena:
-Metacognitive Knowledge, Metacognitive Experience, Metacognitive Goals, and
-Metacognitive Actions. These processes do not occur in isolation, but they
-influence each other, altering cognitive progress over time.
+learning process can be further understood through the lens of Metacognition,
+which focuses on the awareness of one’s mental processes. \cite{flavell79}
+proposed a model of metacognitive monitoring that includes four interrelated
+phenomena: Metacognitive Knowledge, Metacognitive Experience, Metacognitive
+Goals, and Metacognitive Actions. These processes do not occur in isolation, but
+they influence each other, altering cognitive progress over time.
 
@@
 situations \citep{Rivers21}.
+
+In this study, we propose a variation of the interleaving concept, which we call
+Perspective Interleaving. Unlike traditional interleaving, which typically
+involves alternating between different topics or problem types, Perspective
+Interleaving focuses on students' ability to shift between distinct viewpoints.
+There are factors that suggest the Perspective Interleaving hypothesis can
+benefit student learning. Even without an explicit alternation between different
+types of problems, the student can constantly change their perspective, which
+relates to the concept of contextual discrimination. Furthermore, Kolb's
+Experiential Learning Theory reinforces this idea, stating that alternating
+between concrete experiences and abstract reflections is essential to sustain
+the learning cycle. Finally, deliberately shifting one's perspective on a
+specific problem can help students reflect on their approach and improve their
+metacognitive strategies.
 

### 2024-10-10 CSEDU2025/Paper.tex
update paper

@@
 students can make decisions based solely on the information provided by the tool.
-As a potential consequence, \cite{Eason23} observed in a study that when
+As a potential consequence, \cite{Bastani24} observed in a study that when
 programming students lost access to ChatGPT-4, those who had previously relied
@@
 advantageous and intuitive \citep{Boudouaia24}. All of these studies highlight
-the need to increase the understanding of how students interact with these tools
-and how they can be used to improve learning. Furthermore, a review organized by
-\cite{Lo24} emphasized the need for extended studies and objective measures to
-gather more robust evidence on the use of GenAI tools in education.
+the need to increase the understanding how students interact with these tools
+and how they can be used to improve learning, as emphasized \cite{Lo24}.
 
@@
     \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{3cm} p{12cm}} % Remove vertical bars
+    \begin{tabular}{p{4cm} p{12cm}} % Remove vertical bars
       \hline
@@
       \hline
-      Debugging Help & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
-      Conceptual Question & Prompts that are more about understanding concepts than specific code. \\
-      Student Correction & Prompts where the student corrects the bot. \\
-      Code Snippet & Prompts that ask for a specific part of the code, like a function or a segment. \\
-      Complete Solution & Prompts that request an entire solution or a complete code snippet. \\
-      Multiple Question & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
-      Language Change & Prompts that request a change of idiom. \\
-      Uncategorized & Prompts that do not fit into any of the above categories. \\
+      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
+      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. \\
+      Student Correction (SC) & Prompts where the student corrects the bot. \\
+      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. \\
+      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. \\
+      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
+      Language Change (LC) & Prompts that request a change of idiom. \\
+      Uncategorized (U) & Prompts that do not fit into any of the above categories. \\
       \hline
@@
 personal abilities, perceived difficulty in the task, and previously used
-strategies. Metacognitive Experience refers to the feelings that arise
-before, during, and after cognitive activity, such as frustration, confusion,
-and satisfaction, among others. Metacognitive Goals are the key to regulating
-thought as they relate to the goals the individual seeks to achieve, directly
+strategies. Metacognitive Experience refers to the feelings that arise before,
+during, and after cognitive activity, such as frustration, confusion,
+satisfaction, and others. Metacognitive Goals are the key to regulating thought
+as they relate to the goals the individual seeks to achieve, directly
 influencing the actions taken. For example, if a student’s goal is to complete
@@
 \citep{Ouhao18, Carvalho20}, interleaving topics \citep{Rivers21}, utilizing
-retrieval practice \citep{larsen18}, and other strategies.
+retrieval practice \citep{larsen18}, and other possible strategies.
 
-Spacing and interleaving are two metacognitive strategies that have been shown to
-enhance learning outcomes. Spacing refers to the practice of distributing study
-sessions over time, which has been shown to improve long-term retention and
-understanding of the material \citep{Carvalho20}. Interleaving involves mixing
-different topics or problems within a single learning session, which has been
-shown to enhance long-term learning and the application of knowledge in other
+Spacing and Interleaving are two metacognitive strategies that have been shown
+to enhance learning outcomes and supports the research questions of this study.
+Spacing refers to the practice of distributing study sessions over time, which
+has been shown to improve long-term retention and understanding of the material
+\citep{Carvalho20}. Interleaving involve mixing different topics or problems
+within a single learning session, which has been shown to also enhance long-term
+learning and the application of student's knowledge in other contexts and
 situations \citep{Rivers21}.
 
-In this study, we propose a variation of the interleaving concept, which we call
-Perspective Interleaving. Unlike traditional interleaving, which typically
-involves alternating between different topics or problem types, Perspective
-Interleaving focuses on students' ability to shift between distinct viewpoints.
-There are factors that suggest the Perspective Interleaving hypothesis can
-benefit student learning. Even without an explicit alternation between different
-types of problems, the student can constantly change their perspective, which
-relates to the concept of contextual discrimination. Furthermore, Kolb's
-Experiential Learning Theory reinforces this idea, stating that alternating
-between concrete experiences and abstract reflections is essential to sustain
-the learning cycle. Finally, deliberately shifting one's perspective on a
-specific problem can help students reflect on their approach and improve their
-metacognitive strategies.
+Since students are often dealing with a single problem during an interaction
+session with the bot, in this study we propose a specialized form of the
+Interleaving strategy, termed the \textit{Perspective Interleaving Hypothesis}.
+Unlike traditional Interleaving, which typically involves alternating between
+different topics or problem types, Perspective Interleaving can be characterized
+by a metacognitive strategy that intentionally switches between angles of a
+study object.
+
+An important factor supporting the hypothesis is that when switching between
+categories, such as an abstraction and a concrete implementation, a student is
+engaging in Discriminative Contrasts \citep{kang12}. Furthermore, the hypothesis
+that this approach enhances student learning may be grounded on Kolb’s
+Experiential Learning Theory, which posits that learning involves a continuous
+and structured cycle of concrete experiences, reflective observation, abstract
+conceptualization, and active experimentation \citep{kolb14}.
 
 \subsection{Related Work}
-
-Studies on the use of metacognitive strategies indicate that these practices can
-be applied to teaching the conscious use of generative AI. \cite{Zheng19}
-investigated the effects of metacognitive scaffolds on group behavior,
-performance, and cognitive load in computer-supported collaborative
-environments. The results showed positive impacts on metacognitive
-behavior and group performance without increasing cognitive load.
-
-\cite{LiWei23} proposed a metacognition-based collaborative approach to enhance
-performance in collaborative programming. The findings indicated that this
-approach influenced computational thinking, critical thinking, and
-metacognitive awareness.
-
-Similarly, \cite{Wang23} explored how metacognitive instruction affects
-computational thinking, critical thinking, and metacognitive skills in
-collaborative programming, observing improvements in problem-solving and
-collaboration.
-
-These studies demonstrate the relevance of metacognition in enhancing
-performance and skills in collaborative environments without imposing additional
-cognitive load. Based on this, it is possible to suggest that metacognitive
-strategies may also be useful in teaching a more conscious and critical use of
-generative AI tools.
 
@@
 explored how these variables interacted with learner characteristics, perceived
-usefulness of AI, and academic performance. Findings revealed that students
-with higher self-efficacy, lower fear of failure, or higher prior grades
-tended to use AI less frequently or later in the problem-solving process and
-perceived it as less useful compared to their peers. However, no significant
-relationship was found between students’ self-regulation strategies and their
-use of AI.
+usefulness of AI, and academic performance. Findings revealed that students with
+higher self-efficacy, lower fear of failure, or higher prior grades tended to
+use AI less frequently or  later in the problem-solving process and perceived it
+as less useful compared to their peers. However, no significant relationship was
+found between students’ self-regulation strategies and their use of AI.
 
-\cite{Khusnul24} investigates the efficacy of a meta-learning approach in
-improving metacognitive and creative skills. This study confirms the
-effectiveness of meta-learning strategies and elucidates the relationship
-between meta-learning and metacognition. This research suggests that
-meta-learning can improve metacognitive abilities, providing valuable insights
-into educational technology and course design in higher education settings.
+\cite{Bastani24} investigated the impact of generative artificial intelligence
+(AI), specifically GPT-4, on human learning, with a focus on mathematics
+education in a high school. The problem addressed was how the use of generative
+AI could affect the acquisition of new skills, which is crucial for long-term
+productivity. The method involved a controlled experiment with approximately
+one thousand students, who were exposed to two GPT-4-based tutors: a simple
+tutor (GPT Base) and another with safeguards designed to promote learning
+(GPT Tutor). The results showed that while access to GPT-4 improved performance
+on practice exercises (48\% with GPT Base and 127\% with GPT Tutor), the removal
+of access to GPT Base led to a 17\% decrease in student performance on exams.
+This indicated that unrestricted use of *GPT Base* could hinder learning.
+However, the GPT Tutor was able to mitigate this negative effect.
 
-Reconceptualizing AI Literacy: The Importance of Metacognitive Thinking in an Artificial Intelligence (AI)-Enabled Workforce
+\cite{Boudouaia24} investigated the impact of ChatGPT-facilitated programming on
+college students’ programming behaviors, performance, and perceptions. The
+problem addressed was whether the use of ChatGPT in programming education could
+influence how students interacted with programming tasks, their learning
+outcomes, and their attitudes toward the technology. The study employed a
+quasi-experimental design involving 82 college students divided into two
+groups: one using ChatGPT-assisted programming (CFP) and the other using
+traditional self-directed programming (SDP). The method combined mixed data
+collection techniques, including behavioral logs, programming performance
+evaluations, and interviews. The results showed that students using the CFP mode
+exhibited more frequent behaviors such as debugging and reading feedback.
+Although CFP students had a higher average score in programming performance,
+there was no statistically significant difference compared to the SDP group
+However, students’ perceptions of ChatGPT improved significantly after the
+intervention, with increased perceived usefulness, ease of use, and intention to
+continue using the tool in the future.
+
+\cite{Sidra24} addressed the growing importance of metacognitive skills in an
+AI-enabled workforce. The problem discussed was how AI systems, especially
+collaborative tools like large language models (LLMs), impacted human
+decision-making and the challenges they posed due to the lack of human-like
+metacognition in AI. The method used identified four main characteristics that
+differentiated human-AI interaction from human-human collaboration and proposed,
+based on the dual-process theory of human thought, that metacognitive thinking
+was crucial for the effective use of AI. The results indicated that human
+workers needed to develop and apply metacognitive skills—such as planning,
+monitoring, and evaluating their thought processes—to mitigate cognitive biases
+and improve decision-making when working with AI. The study concluded by
+recommending training and AI design improvements to enhance workers’
+metacognitive abilities, ensuring more effective collaboration with AI systems.
+
+Learning based on metacognitive strategies has been shown to be effective in
+enhancing students’ performance and skills. \cite{Zheng19} investigated the
+effects of metacognitive scaffolds on group behavior, performance, and cognitive
+load in computer-supported collaborative environments. The results showed
+positive impacts on metacognitive behavior and group performance without
+increasing cognitive load. \cite{LiWei23} proposed a metacognition-based
+collaborative approach to enhance performance in collaborative programming. The
+findings indicated that this approach influenced computational thinking,
+critical thinking, and metacognitive awareness. Similarly, \cite{Wang23}
+explored how metacognitive instruction affects computational thinking, critical
+thinking, and metacognitive skills in collaborative programming, observing
+improvements in problem-solving and collaboration. \cite{Khusnul24} investigates
+the efficacy of a meta-learning approach in improving metacognitive and creative
+skills. This study confirms the effectiveness of meta-learning strategies and
+elucidates the relationship between meta-learning and metacognition. This
+research suggests that meta-learning can improve metacognitive abilities,
+providing valuable insights into educational technology and course design in
+higher education settings.
 
@@
 Retrieval-Augmented Generation (RAG) technique to improve the precision of
-responses provided to students \cite{Sun24}. The RAG feature was developed using
-materials from a MOOC, including text, exercises, video transcripts, and other
-resources.
+responses provided to students \cite{Sun24}. The RAG approach was developed
+using materials from a MOOC, including text, exercises, video transcripts, and
+other resources.
 
@@
 into eight distinct categories for analysis. Initially, \cite{Ghimire24}
-proposed four categories: Debugging Help, Conceptual Question, Code Snippet,
-and Complete Solution. However, the data collected indicated the need for
-additional categories, leading to the inclusion of four more: Multiple Choice
-(Quiz), Student Corrections, Language Change, and Uncategorized. In this study,
-students were free to use the bot as they wanted, which contributed to the
-emergence of these additional categories. Table \ref{tab:categories} presents
-these categories along with their respective descriptions.
+proposed four categories: Debugging Help (DH), Conceptual Question (CQ), Code
+Snippet (CS), and Complete Solution (CSO). However, the data collected indicated
+the need for additional categories, leading to the inclusion of four more:
+Multiple Question (MQ), Student Corrections (SC), Language Change (LC), and
+Uncategorized (U). In this study, students were free to use the bot as they
+wanted, which contributed to the emergence of these additional categories.
+Table \ref{tab:categories} presents these categories along with their respective
+descriptions.
 
 The analysis phase involved the use of Python scripts to extract information
-from previously classified data. The datasets and code snippets used in this
-process are available at: YYY.
+from previously classified data. The datasets and code used in this process are
+available at: https://github.com/rodrigoprestesmachado/uc3m.
 
@@
 There is a noticeable difference in the interaction time with the bot when a
-student asks a conceptual question. This was probably due to the fact that, in
-this category, the student is more focused on understanding a concept and
-engaging in deeper learning, which aligns with the metacognitive goal of
-promoting comprehension and long-term retention. Figure \ref{fig:graph3}
-shows the average interaction time per category.
+student asks a conceptual question. The time gap creates an effect of
+information retrieval when resuming the activity. Therefore, we can consider
+this category as one in which students are more cognitively engaged. Thus, in
+this category, students are likely more focused on understanding, which aligns
+with the metacognitive goal of promoting comprehension and long-term retention.
+Figure \ref{fig:graph3} shows the average interaction time per category.
 
@@
 
-TODO: Work in progress.
+The results suggest that most student interactions with GenAI tools are centered
+on active learning strategies, although the deliberate use of interleaving and
+spacing techniques remains limited. Guiding students toward a more intentional
+application of these strategies could enhance their learning outcomes. The
+anonymity of interactions leaves it unclear whether these strategies were used
+deliberately or emerged naturally, highlighting the need for future research on
+students' conscious application of these methods. To support educators in
+analyzing student behavior and offering more targeted assistance, an AI model
+is being developed to automatically classify student messages, providing
+valuable insights into how these tools are being utilized.
 
-* The results indicate that most interactions are focused on active learning
-strategies, although deliberate use of interleaving and spacing is limited. We
-believe that students can be guided \cite{Zheng19} \cite{Wang23} \cite{LiWei23}
-\cite{Khusnul24} to adopt a more deliberate approach in utilizing GenAI,
-enabling them to more effectively harness these tools to enhance their learning.
+% TODO: Work in progress.
 
-* Due to the anonymity of the interactions, it's unclear whether
-students were intentionally using interleaving and spacing strategies or if
-these patterns emerged naturally. Future research should investigate whether
-students were consciously applying.
+% * The results indicate that most interactions are focused on active learning
+% strategies, although deliberate use of interleaving and spacing is limited. We
+% believe that students can be guided \cite{Zheng19} \cite{Wang23} \cite{LiWei23}
+% \cite{Khusnul24} to adopt a more deliberate approach in utilizing GenAI,
+% enabling them to more effectively harness these tools to enhance their learning.
 
-* To provide teachers with valuable data for analyzing how students are using
-the tools, we are developing an AI model to automatically classify student
-messages. This will enable educators to better understand student behavior and
-offer more targeted support where needed.
+% * Due to the anonymity of the interactions, it's unclear whether
+% students were intentionally using interleaving and spacing strategies or if
+% these patterns emerged naturally. Future research should investigate whether
+% students were consciously applying.
+
+% * To provide teachers with valuable data for analyzing how students are using
+% the tools, we are developing an AI model to automatically classify student
+% messages. This will enable educators to better understand student behavior and
+% offer more targeted support where needed.
 

### 2024-10-10 CSEDU2025/Paper.tex
update

@@
 
-Spacing, a well-established learning strategy \citep{Carvalho20}, was analyzed
-by examining the time intervals between user interactions with CharlieBot.
 Approximately 26.2\% of the conversations, or roughly a quarter, included pauses
@@
 
-\subsection{Interleaving}
+\subsection{Perspective Interleaving}
 
@@
 promoting interleaving?}
-
-Intervealing is a learning strategy that involves mixing or alternating
-different topics or problems within a single learning session. This approach has
-been shown to be useful for promoting long-term learning and improving the
-application of knowledge in other situations \citep{Rivers21}.
 

### 2024-10-16 CSEDU2025/Paper.tex
update data

@@
 Figure \ref{fig:graph1} also shows that around 5\% of the prompts written by
-students were not categorized. These messages range from greetings to the Bot,
-such as a simple "\textit{Hello}," to contextual statements like
+students were not categorized. These messages range from gratitude to the Bot,
+such as a simple "\textit{thanks}," to contextual statements like
 "\textit{I'm reviewing object orientation}"

### 2024-10-29 CSEDU2025/Paper.tex
update paper

--- 2024-10-10 CSEDU2025/Paper.tex
@@
 \usepackage{SCITEPRESS}
+\usepackage{arydshln}
 
@@
 \title{
-  Examining Programming Students’ Interaction with Generative AI: A
-  Metacognitive Perspective
+  Analyzing Student Use of Spacing and Interleaving Strategies in Interactions
+  with GenAI-Powered Chatbots in Programming Courses
 }
@@
 \keywords{
-  Programming Education, Generative Artificial Intelligence, Chatbots and
-  Metacognition
+  Programming, Generative Artificial Intelligence, Chatbots, Spacing,
+  Interleaving and Metacognition
 }
@@
 \abstract{
-This study examines the interactions of programming students with a chatbot,
-CharlieBot, powered by OpenAI's GPT-3.5 and enhanced with the
-Retrieval-Augmented Generation (RAG) technique, within the context of a Java
-programming course. It investigates how students use the bot while employing two
-metacognitive strategies: interleaving and spacing. Using the logs of student
-interactions with the bot, the methodology involved classifying and analyzing
-student prompts into eight distinct categories. The findings reveal that while
-most interactions emphasize active learning strategies, students demonstrate
-limited conscious application of interleaving and spacing. This study
-underscores the need for more guidance on the strategic use of AI
-tools to improve learning outcomes.
+This study examines the interactions of programming students with a chatbot
+powered by OpenAI's GPT-3.5 and enhanced with the Retrieval-Augmented Generation
+(RAG) technique, within the context of a Java programming course. It
+investigates how students use the bot while employing two metacognitive
+strategies: interleaving and spacing. Using the logs of student interactions
+with the bot, the methodology involved classifying and analyzing student prompts
+into eight distinct categories. The findings reveal that while most interactions
+emphasize active learning strategies, students demonstrate limited conscious
+application of interleaving and spacing. This study underscores the need for
+more guidance on the strategic use of AI tools to improve learning outcomes.
 }
@@
 advantageous and intuitive \citep{Boudouaia24}. All of these studies highlight
-the need to increase the understanding how students interact with these tools
+the need to increase the understanding of how students interact with these tools
 and how they can be used to improve learning, as emphasized \cite{Lo24}.
@@
 the interactions between students in a Java programming course and an
-educational bot powered by gpt-3.5 from OpenAI, enhanced with the
+educational bot powered by OpenAI gpt-3.5, enhanced with the
 Retrieval-Augmented Generation (RAG) technique. Specifically, it focuses on
-examining these interactions within the context of the metacognitive strategies
-employed by students. To achieve this objective, we formulate three research
+examining these interactions within the context of two metacognitive strategies:
+spacing and interleaving. To achieve this objective, we formulate three research
 questions:
@@
 \begin{itemize}
-  \item RQ1 - What was the distribution pattern of the classified messages?
-  \item RQ2 - Was there a space between the student prompts? If so, which
-  category promotes the largest spacing between submissions?
-  \item RQ3 - Do students' conversations with the bot alternate between
-  practical and theoretical approaches, thereby promoting interleaving?
+  \item RQ1 - What distribution patterns emerged in the classified student
+  interaction prompts with the educational chatbot?
+  \item RQ2 - Was there spacing between student prompts, and if so, which
+  category led to the most rapid and consistent interactions with the bot?
+  \item RQ3 - Do students' interactions with the bot alternate between study
+  topics to suggest interleaving?
 \end{itemize}
@@
 
-This paper is organized as follows: Section 2 presents the theoretical
+This paper is organized as follows. Section 2 presents the theoretical
 framework, Section 3 describes the material and methods used in the study,
@@
 we introduce the concept of metacognition used in this study. Next, we discuss
-related work that explores the use of metacognitive strategies and scaffolds
+related work exploring the use of metacognitive strategies and scaffolds
 in educational settings.
@@
 
-Given that generative AI often produces automatic and highly accurate responses
-\citep{Puryear22}, it is essential for students to use these tools within an
+Given that generative AI often produces highly accurate and automatic responses
+\citep{Puryear22}, it is essential that students use these tools within an
 active learning process to enhance their learning outcomes. This active
-learning process can be further understood through the lens of Metacognition,
-which focuses on the awareness of one’s mental processes. \cite{flavell79}
+learning process can be further understood through the lens of metacognition,
+which focuses on awareness of one's mental processes. \cite{flavell79}
 proposed a model of metacognitive monitoring that includes four interrelated
@@
 
-Spacing and Interleaving are two metacognitive strategies that have been shown
-to enhance learning outcomes and supports the research questions of this study.
+Spacing and interleaving are two metacognitive strategies that have been shown
+to enhance learning outcomes and support the research questions of this study.
 Spacing refers to the practice of distributing study sessions over time, which
 has been shown to improve long-term retention and understanding of the material
-\citep{Carvalho20}. Interleaving involve mixing different topics or problems
+\citep{Carvalho20}. Interleaving involves mixing different topics or problems
 within a single learning session, which has been shown to also enhance long-term
@@
 situations \citep{Rivers21}.
-
-Since students are often dealing with a single problem during an interaction
-session with the bot, in this study we propose a specialized form of the
-Interleaving strategy, termed the \textit{Perspective Interleaving Hypothesis}.
-Unlike traditional Interleaving, which typically involves alternating between
-different topics or problem types, Perspective Interleaving can be characterized
-by a metacognitive strategy that intentionally switches between angles of a
-study object.
-
-An important factor supporting the hypothesis is that when switching between
-categories, such as an abstraction and a concrete implementation, a student is
-engaging in Discriminative Contrasts \citep{kang12}. Furthermore, the hypothesis
-that this approach enhances student learning may be grounded on Kolb’s
-Experiential Learning Theory, which posits that learning involves a continuous
-and structured cycle of concrete experiences, reflective observation, abstract
-conceptualization, and active experimentation \citep{kolb14}.
 
@@
 relationship between AI usage and students’ self-regulation strategies,
-self-efficacy, and fear of failure in programming. Additionally, the study
-explored how these variables interacted with learner characteristics, perceived
-usefulness of AI, and academic performance. Findings revealed that students with
-higher self-efficacy, lower fear of failure, or higher prior grades tended to
-use AI less frequently or  later in the problem-solving process and perceived it
-as less useful compared to their peers. However, no significant relationship was
-found between students’ self-regulation strategies and their use of AI.
+self-efficacy, and fear of failure in programming. Furthermore, the study
+explored how these variables interacted with the characteristics of the learners,
+the perceived usefulness of AI, and academic performance. The findings revealed
+that students with higher self-efficacy, lower fear of failure or higher prior
+grades tended to use AI less frequently or later in the problem solving process
+and perceived it as less useful compared to their peers. However, no significant
+relationship was found between students’ self-regulation strategies and their
+use of AI.
 
@@
 was crucial for the effective use of AI. The results indicated that human
-workers needed to develop and apply metacognitive skills—such as planning,
-monitoring, and evaluating their thought processes—to mitigate cognitive biases
+workers needed to develop and apply metacognitive skills, such as planning,
+monitoring, and evaluating their thought processes, to mitigate cognitive biases
 and improve decision-making when working with AI. The study concluded by
@@
 positive impacts on metacognitive behavior and group performance without
-increasing cognitive load. \cite{LiWei23} proposed a metacognition-based
+increasing the cognitive load. \cite{LiWei23} proposed a metacognition-based
 collaborative approach to enhance performance in collaborative programming. The
@@
 elucidates the relationship between meta-learning and metacognition. This
-research suggests that meta-learning can improve metacognitive abilities,
+research suggests that metalearning can improve metacognitive abilities,
 providing valuable insights into educational technology and course design in
@@
 
+\begin{figure}[h!]
+  \centering
+  \includegraphics[scale=0.62]{img/figure1.png}
+  \caption{Categorization of messages.}
+  \label{fig:graph1}
+\end{figure}
+
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.45]{img/figure2.png}
+  \caption{Examples of conversations.}
+  \label{fig:graph2}
+\end{figure*}
+
 In the categorization phase, the students' messages were manually classified
-into eight distinct categories for analysis. Initially, \cite{Ghimire24}
-proposed four categories: Debugging Help (DH), Conceptual Question (CQ), Code
-Snippet (CS), and Complete Solution (CSO). However, the data collected indicated
-the need for additional categories, leading to the inclusion of four more:
-Multiple Question (MQ), Student Corrections (SC), Language Change (LC), and
-Uncategorized (U). In this study, students were free to use the bot as they
-wanted, which contributed to the emergence of these additional categories.
+into eight distinct categories for analysis, representing different types of
+prompt that students use when interacting with the bot. Initially,
+\cite{Ghimire24} proposed four categories: Debugging Help (DH), Conceptual
+Question (CQ), Code Snippet (CS), and Complete Solution (CSO). However, the data
+collected indicated the need for additional categories, leading to the inclusion
+of four more: Multiple Question (MQ), Student Corrections (SC), Language Change
+(LC), and Uncategorized (U). In this study, students were free to use the bot as
+they wanted, which contributed to the emergence of these additional categories.
 Table \ref{tab:categories} presents these categories along with their respective
-descriptions.
+descriptions. To facilitate the analysis of interleaving, the prompt topics were
+also categorized according to the course subjects, which are: (1) Java, (2)
+Object Orientation, (3) Testing, (4) Recursion, (5) Data Structures, and (6)
+Sorting and Searching Algorithms.
 
@@
 
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.39]{img/figure3.png}
+  \caption{Average interaction time per category.}
+  \label{fig:graph3}
+\end{figure*}
+
 \section{\uppercase{Results and Discussion}}
 
-A total of 343 student messages were categorized in 42 conversations, with
-an average of 8.17 messages per conversation. This data offers insights into how
-users engage with CharlieBot and the types of prompts they submit. The results of
-the analysis are presented in this section, addressing the research questions
-outlined in the introduction.
+A total of 625 student messages were categorized in 81 conversations, with
+an average of 7.72 messages per conversation. This data offers insights into how
+users engage with CharlieBot and the types of prompts they submit. The results
+of the analysis are presented in this section, addressing the research questions
+outlined in the introduction. Figure \ref{fig:graph1} shows the distribution of
+the messages in the categories.
 
-Figure \ref{fig:graph1} shows the distribution of the messages in the
-categories:
+\subsection{Categorization of messages}
 
-\begin{figure}[h!]
-    \centering
-    \includegraphics[scale=0.62]{img/figure1.png}
-    \caption{Classification of messages.}
-    \label{fig:graph1}
-\end{figure}
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.52]{img/figure2.png}
-  \caption{Examples of conversations.}
-  \label{fig:graph2}
-\end{figure*}
-
-\subsection{Classification of Messages}
-
-About the first research question: \textit{RQ1: What was the distribution
-pattern of the classified messages?}
+About the first research question: \textit{RQ1: What distribution patterns
+emerged in the classified student interaction prompts with the educational
+chatbot?}
 
 According to Figure \ref{fig:graph1} messages classified as
-Conceptual Question account for 32.1\% of the total. Often, the sequence of
-Conceptual Question begins with a request for a Complete Solution or Code
-Snippet, which then evolves into a series of Conceptual Question, as seen in
-conversation 2 of Figure \ref{fig:graph2}, which illustrates some examples of
-students' conversations with the bot. In other cases, the conversation
-consists entirely of messages classified as Conceptual Question, such as in
-conversation 10 of Figure \ref{fig:graph2}. This type of message indicates that
-students recognize that they have not yet fully mastered a concept and are
-seeking to better understand it. In this context, the student realizes that
-solving the problem requires understanding of concepts, thus applying an active
-learning strategy.
+Conceptual Question account for 35.8\% of the total. Sequences of Conceptual
+Questions often start from a practical perspective of a given code, as
+illustrated in conversation 4 of Figure \ref{fig:graph2}, which presents
+examples of students' interactions with the bot. In other cases, the
+conversation consists entirely of messages classified as Conceptual Question,
+such as in conversation 15 of Figure \ref{fig:graph2}.
 
-As shown in Figure \ref{fig:graph1}, approximately 30.9\% of the messages were
-classified as Debugging Help. These prompts often begin with a request for a
-Complete Solution or Code Snippet. Following this, students typically switch from
-passive to more active learning strategy, as illustrated by the start
-of conversation 12 in Figure \ref{fig:graph2}. Debugging Help messages represent
-a form of active learning, in which students likely focus on gaining a practical
-understanding of the code, identifying errors, and improving their debugging
-skills for future tasks.
+As shown in Figure \ref{fig:graph1}, approximately 29.9\% of the messages were
+classified as Debugging Help. Debugging help messages are likely to provide
+students with a practical approach to understanding code, identifying errors,
+and building skills to enhance their debugging abilities for future tasks.
 
-Since most of the messages are located in categories where students are
-notoriously engaged in being more active in their learning (Conceptual Question
-and Debugging Help), corroborating the findings of \cite{Ghimire24} which
-showed that the questions are also localized in the same categories.
+Most of the messages are located in categories Conceptual Question and Debugging
+Help corroborate the findings of \cite{Ghimire24} which showed that the
+questions are also localized in the same categories.
+
+Multiple Question exercise resolutions represented 14.9\% of the responses.
+These prompts typically involved students submitting one or more exercises to
+CharlieBot and requesting solutions. The conversation 13 in \ref{fig:graph2}
+illustrates a sequence of multiple choice questions.
 
 Requests for the bot to generate a Code Snippet or Complete Solution accounted
-for 10.5\% of student messages. These prompts reflected a desire for more
+for 8.2\% of student messages. These prompts reflected a desire for more
 direct answers; however, it was observed that, after receiving these solutions,
@@
 
-Multiple Choice exercise resolutions represented 16.9\% of the responses.
-These prompts typically involved students submitting one or more exercises to
-CharlieBot and requesting solutions. We identified sequences of multiple choice
-questions, as illustrated by conversations 8, 9, and 15 in \ref{fig:graph2}.
-This less active learning approach can occur for various reasons, ranging from
-the belief that reviewing solutions to similar questions enhances learning, to
-issues such as negligence, poor planning, or other possible factors.
+Student corrections to bot responses accounted for 4.2\% of the interactions.
+In most cases, these corrections occurred when the student already had the
+correct answer to an exercise and identified an error in the bot response, as
+illustrated in conversation 2 in Figure \ref{fig:graph2}. These corrections
+highlight that, although the bot provides quick feedback, it is not always
+accurate.
 
-Student corrections to bot responses accounted for 2.6\% of the
-interactions. In most cases, these corrections occurred when the student already
-had the correct answer to an exercise and identified an error in the bot
-response, as illustrated in conversations 4 and 9 in Figure \ref{fig:graph2}.
-These corrections highlight that, although the bot provides quick feedback, it
-is not always accurate and its errors can be difficult for students to
-identify.
+Figure \ref{fig:graph1} also indicates that approximately 5.8\% of the prompts
+written by students were classified as Uncategorized. These messages include
+expressions of gratitude toward the bot, such as a simple '\textit{thanks}',
+contextual statements like '\textit{I'm reviewing object orientation}', as well
+as requests for additional exercises and summaries. Depending on the type of
+investigation, there is opportunity for these messages to be further subdivided
+into new categories.
 
-Figure \ref{fig:graph1} also shows that around 5\% of the prompts written by
-students were not categorized. These messages range from greetings to the Bot,
-such as a simple "\textit{Hello}," to contextual statements like
-"\textit{I'm reviewing object orientation}"
+Finally, 1.3\% of the messages were classified as Language Change. These prompts
+were typically requests for the bot to switch languages, as seen in conversation
+9 in Figure \ref{fig:graph2}.
+
+\begin{table*}[htbp]
+  \caption{Average and Standard Deviation of Different Categories (in minutes)}
+  \begin{center}
+    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
+    \begin{tabular}{p{4cm} p{4cm} p{4cm}} % Remove vertical bars
+      \hline
+      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
+      \hline
+      Code Snippet & 1.26 & 1.74 \\
+      Uncategorized & 1.65 & 3.54 \\
+      Language Change & 1.89 & 3.28 \\
+      Multiple Question & 2.65 & 6.10 \\
+      Student Correction & 3.17 & 7.33 \\
+      \hdashline
+      \textit{Overall} & \textit{3.43} & \textit{6.60} \\
+      \hdashline
+      Debugging Help & 4.84 & 10.13 \\
+      Conceptual Questions & 5.41 & 10.54 \\
+      Complete Solution & 6.55 & 10.14 \\
+      \hline
+    \end{tabular}
+    \label{tab:averages}
+  \end{center}
+\end{table*}
 
@@
 
-% Spacing
+About the second research question: \textit{RQ2: Was there spacing between
+student prompts, and if so, which category led to the most rapid and consistent
+interactions with the bot?}
 
-About the second research question: \textit{RQ2: Was there a space between the
-student prompts? If so, which category promotes the largest spacing between
-submissions?}
+For the analysis of the spacing, we initially considered a minimum interval of
+60 minutes as a reference to define the significant spacing \citep{Gadella24}.
+Based on this, our results showed that, of the 81 conversations analyzed, only
+27 (33.3\%) had at least one message with an interval greater than 60 minutes.
+Figure 2 illustrates the elapsed time between messages (delta). For example,
+in conversation 15, the first message was sent at time zero, while the second
+message was sent by the student after a delta of 0.6 minutes. Additionally,
+message 4 in conversation 15, as shown in Figure \ref{fig:graph2}, presents a
+delta of 69.1 minutes, which, according to our reference, characterizes
+significant spacing. Figure \ref{fig:graph3} also shows the number of spacings
+in each conversation. Among conversations that exhibited spacing, the
+average number of subsections per conversation was 2.2. Therefore, a
+conversation with one instance of spacing is typically divided into two distinct
+study subsections.
 
-Approximately 26.2\% of the conversations, or roughly a quarter, included pauses
-longer than 60 minutes. However, this data only indicates that students paused
-their use of the AI, which could suggest that they were either reflecting on the
-topic independently or engaging in unrelated activities. Furthermore, 16.7\%
-of the conversations exhibited pauses of 12 hours and 9\% showed delays
-exceeding 24 hours between messages, indicating that some degree of spacing
-occurred. However, it is important to note that this behavior may have been
-influenced by the design of the initial version of CharlieBot, which only kept
-the chat session active as long as the student kept the browser tab open and
-avoided reloading the page. This adds uncertainty about the precise context in
-which students applied spacing.
+This data suggests that there is indeed little spacing between study sessions.
+It is worth noting that, in order to create a new section in CharlieBot,
+students must close the browser tab and start a new conversation, which results
+in losing the previous chat history.
 
-There is a noticeable difference in the interaction time with the bot when a
-student asks a conceptual question. The time gap creates an effect of
-information retrieval when resuming the activity. Therefore, we can consider
-this category as one in which students are more cognitively engaged. Thus, in
-this category, students are likely more focused on understanding, which aligns
-with the metacognitive goal of promoting comprehension and long-term retention.
-Figure \ref{fig:graph3} shows the average interaction time per category.
+\subsubsection{Interaction Time per Category}
 
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.55]{img/figure3.png}
-  \caption{Average interaction time per category.}
-  \label{fig:graph3}
-\end{figure*}
+When a student requests a code snippet (CS), it is generally predictable that
+they will resume interaction with the bot within a short period of time, as
+indicated by the mean (1.26) and standard deviation (1.74) of the Code Snippet
+category in Table \ref{tab:averages}. This result is noteworthy because,
+although some requests are simple and lead to quick responses, others may be
+more complex and would require more time for reflection. However, the low
+average response time and minimal variability may indicate that students adopt a
+more passive learning posture when requesting the bot to generate small pieces
+of code.
 
-\subsection{Perspective Interleaving}
+Despite some occasional variation (3.54), prompts classified as Uncategorized
+exhibit a relatively low average time (1.65) for students to resume interaction
+with the bot. This variation occurs, in part, due to situations where students
+end the conversation with an uncategorized message, such as a simple
+\textit{'thanks'} but later choose to resume the interaction.
 
-% Interleaving
+When a student requests the bot to change the language (LC) in a response, the
+new interaction happens quickly (1.89) with little variation in response time
+(3.28), meaning it is predictable that the student will continue interacting.
 
-Regarding the third research question: \textit{RQ3: Do students' conversations
-with the bot alternate between practical and theoretical approaches, thereby
-promoting interleaving?}
+Although still below the overall average, students take a moderate amount of
+time (2.65) to resume interaction with the bot after requesting the solution to
+a multiple-choice question (MQ). The variation (6.10) suggests that some
+responses may lead to a slower follow-up interaction, possibly indicating a
+moment of deeper reflection on the part of the students. However, since both the
+mean and standard deviation are below the overall average, this could indicate
+a more passive behavior, as students obtain ready-made answers from the bot.
 
-When analyzing the classification of messages in conversations, we identified
-the presence of sequences of messages belonging to the same category, which
-could suggest a low occurrence of interleaving. To illustrate the presence of
-blocks, Figure \ref{fig:graph2} presents examples of conversations and their
-classification. In Conversation 2, the student alternates messages among Code
-Snippet, Debugging Help, and Conceptual Question, showing a tendency to
-interleave from a practical to theoretical perspective. In contrast, in
-Conversation 10, the student focused on a single block of Conceptual Question.
-Although the focus was on understanding a theoretical concept, this conversation
-tended not to promote interleaving, due to the absence of alternation between
-different types of task.
+% This less active learning approach can occur for various reasons, ranging from
+% the belief that reviewing solutions to similar questions enhances learning, to
+% issues such as negligence, poor planning, or other possible factors.
 
-To better assess the extent of these sequences, we calculated the percentage of
-messages that were part of repeated blocks. Considering blocks with at least
-three messages, we observed that 56.2\% of the student's prompts were grouped
-into some sequence. When analyzing blocks with a minimum of four messages, we
-found that 47.5\% of the prompts were part of continuous sequences.
+The average time for students to resume interaction in the correction category
+(SC) is close to the overall average (3.17), although the high standard
+deviation (7.33) indicates significant variation. In some cases, students are
+already familiar with the answer and only seek a quick correction from the bot;
+in others, they detect flaws in the bot's responses and need to carefully
+examine the subsequent answer to identify potential errors, which suggests a
+more active behavior.
 
-However, if we consider that sequential messages from a certain point indicate
-that the student is \textit{blocked} (or focused) on a task, the remaining space
-for interleaving is 43.8\%, assuming a block starting from at least three
-consecutive messages, and 52.5\% for blocks starting from four consecutive
-messages. Thus, the data suggest that interleaving occurs on students'
-conversations with CharlieBot, but in which situations and with what type of
-student profile is a study that needs to be further investigated.
+After a question classified as Debugging Help (DH), there is both a
+longer-than-average response time (4.84) and a high variation in the time taken
+for a new interaction (10.13). This suggests that student prompts in this
+category tend to be more time-consuming and unpredictable. A possible
+explanation is that some DH issues are simple, such as '\textit{what does x++ do?}',
+while others involve requests for analysis of more complex code, like
+\textit{'explain this piece of code (accompanied by a complete method).'}
+
+Conceptual questions (CQ) generally result in a high average time (5.41) for
+students to re-engage with the bot. Moreover, the high standard deviation
+(10.54) indicates an unpredictable variation in response times. This
+unpredictability may be attributed to the fact that, in some cases, the
+complexity of the question and answer demands more reflection time from the
+student, whereas in others, lesser complexity allows for a quicker interaction
+with the bot.
+
+Finally, a message classified as a Complete Solution request (SCO) exhibits the
+longest average time (6.55) for students to resume interaction with the bot.
+Additionally, the high standard deviation (10.14) highlights significant
+unpredictability in response time. After requesting a complete solution,
+students tend to take longer to return, due to the complexity of the problems
+involved, such as the implementation of sorting algorithms like Heapsort. Table
+\ref{tab:averages} presents the means and standard deviations for each
+category.
+
+However, some categories suggest quicker interactions with less reflection on
+the part of students, such as 'Code Snippet' and 'Multiple Choice.' On the other
+hand, other categories indicate that students may be in a more reflective
+process, such as 'Debugging Help,' 'Conceptual Question,' and 'Complete
+Solution.' This demonstrates that the level of student engagement varies
+according to the nature of the requested interaction, reflecting different
+approaches and depths in the learning process.
+
+\subsection{Interleaving}
+
+About the third research question: \textit{RQ3: Do students' interactions with
+the bot alternate between study topics to suggest interleaving?}
+
+To analyze the interweaving of topics, sections were initially divided into
+those that exhibited spacing and those that did not.
+
+In conversations with the bot that did not include spacing (66.7\% of the total
+conversations), approximately 19.2\% of the messages resulted in a change of
+topic. Qualitative analysis of what led to this interweaving indicates that
+part of it was due to students' requests for assistance in solving exercises
+from different topics, suggesting a less active learning approach, as the time
+interval for this type of interaction tends to be shorter than average.
+Additionally, some instances of interweaving were observed after a prolonged
+period of 20 to 50 minutes, still below the 60-minute threshold established as
+the cutoff for considering spacing.
+
+In interactions with the bot that included spacing (33.3\% of the total
+conversations), approximately 31.4\% of the messages switched topics.
+Qualitative analysis of these interactions reveals that a significant portion
+of this interweaving was caused by spacing. In other words, some students use
+the bot for quick consultations, asking a question and then returning hours
+later with another question on a different subject.
+
+Thus, the data revealed that the existing interweaving is limited and does not
+suggest an effective study strategy, as it mainly arises from requests for
+exercise resolution and from the natural spacing pattern of a consultation
+interaction.
 
@@
 
-The results suggest that most student interactions with GenAI tools are centered
-on active learning strategies, although the deliberate use of interleaving and
-spacing techniques remains limited. Guiding students toward a more intentional
-application of these strategies could enhance their learning outcomes. The
-anonymity of interactions leaves it unclear whether these strategies were used
-deliberately or emerged naturally, highlighting the need for future research on
-students' conscious application of these methods. To support educators in
-analyzing student behavior and offering more targeted assistance, an AI model
-is being developed to automatically classify student messages, providing
-valuable insights into how these tools are being utilized.
+The objective of this work was to analyze the interaction of second-semester
+Java programming students with an educational bot developed using the RAG
+technique, employing two metacognitive strategies: spacing and interleaving.
 
-% TODO: Work in progress.
+Initially, the theory of metacognition, alongside the classification of student
+actions, has proven effective in understanding and revealing how students
+utilize the bot. Although the data shows that most messages fall within the
+categories of Conceptual Questions and Debugging Help categories identified as
+promoting slower but potentially deeper engagement with the bot, indicative of a
+more active form of learning—the findings also reveal that students' use of
+spacing and interleaving strategies remains quite limited. Encouraging students
+to apply these strategies more intentionally could significantly enhance their
+learning outcomes. To further explore this, a broader study is planned to
+examine, for instance, the relationship between student profiles and their
+strategies in using a Generative AI tool. Additionally, to support educators
+in analyzing student behavior and providing more targeted assistance, an AI
+model is being developed to automatically classify student messages, offering
+valuable insights into the ways these tools are being employed.
 
-% * The results indicate that most interactions are focused on active learning
-% strategies, although deliberate use of interleaving and spacing is limited. We
-% believe that students can be guided \cite{Zheng19} \cite{Wang23} \cite{LiWei23}
-% \cite{Khusnul24} to adopt a more deliberate approach in utilizing GenAI,
-% enabling them to more effectively harness these tools to enhance their learning.
-
-% * Due to the anonymity of the interactions, it's unclear whether
-% students were intentionally using interleaving and spacing strategies or if
-% these patterns emerged naturally. Future research should investigate whether
-% students were consciously applying.
-
-% * To provide teachers with valuable data for analyzing how students are using
-% the tools, we are developing an AI model to automatically classify student
-% messages. This will enable educators to better understand student behavior and
-% offer more targeted support where needed.
-
-\section*{ACKNOWLEDGEMENTS}
-
-The authors would like to thank the Federal Institute of Education, Science and
-Technology (IFRS) for the partial financial support provided for the execution
-of this research.
+% The results suggest that most student interactions with GenAI tools are centered
+% on active learning strategies, although the deliberate use of interleaving and
+% spacing techniques remains limited. Guiding students toward a more intentional
+% application of these strategies could enhance their learning outcomes. The
+% anonymity of interactions leaves it unclear whether these strategies were used
+% deliberately or emerged naturally, highlighting the need for future research on
+% students' conscious application of these methods. To support educators in
+% analyzing student behavior and offering more targeted assistance, an AI model
+% is being developed to automatically classify student messages, providing
+% valuable insights into how these tools are being utilized.
 
@@
 \bibliography{References}}
-
-\section*{\uppercase{\textit{Appendix}}}
-
-If any, the appendix should appear directly after the
-references without numbering, and not on a new page. To do so please use the
-following command: \textit{$\backslash$section*\{APPENDIX\}}
 

### 2024-10-30 CSEDU2025/Paper.tex
Update paper

@@
 \abstract{
-This study examines the interactions of programming students with a chatbot
-powered by OpenAI's GPT-3.5 and enhanced with the Retrieval-Augmented Generation
-(RAG) technique, within the context of a Java programming course. It
-investigates how students use the bot while employing two metacognitive
-strategies: interleaving and spacing. Using the logs of student interactions
-with the bot, the methodology involved classifying and analyzing student prompts
-into eight distinct categories. The findings reveal that while most interactions
-emphasize active learning strategies, students demonstrate limited conscious
-application of interleaving and spacing. This study underscores the need for
-more guidance on the strategic use of AI tools to improve learning outcomes.
+The aim of this study is to analyze the interactions of programming students
+with a chatbot powered by OpenAI's GPT-3.5, enhanced with the Retrieval
+Augmented Generation (RAG) technique, within the context of a Java programming
+course. The study investigates how students use the chatbot while employing
+two metacognitive strategies: interleaving and spacing. By analyzing the logs
+of student interactions with the chatbot, the methodology involved
+classifying and categorizing student prompts into eight distinct types. The
+findings reveal that, while most interactions focus on active learning
+strategies, students show limited conscious application of study strategies such
+as interleaving and spacing. The study highlights the need for more
+comprehensive guidance on the strategic use of AI tools to enhance learning
+outcomes.
 }
@@
   interaction prompts with the educational chatbot?
-  \item RQ2 - Was there spacing between student prompts, and if so, which
-  category led to the most rapid and consistent interactions with the bot?
+  \item RQ2 - Was there spacing between student prompts? Which category led to
+  the most rapid and consistent interactions with the bot?
   \item RQ3 - Do students' interactions with the bot alternate between study
@@
 paper and outlines future work.
-
-%% Are there any existing solutions?
-% To deal with this issue, some researchers have proposed different pedagogical
-% strategies \cite{Denny24} introduced the concept of \textit{Prompt
-% Problems}, where students solve programming exercises by formulating natural
-% language prompts. \cite{Prasad24} proposed a self-regulated learning framework
-% using GenAI to solve programming problems. \cite{Lauren23} explored the
-% integration of GenAI with evidence-based learning strategies in computer science
-% and engineering courses.
-
-%% Which is the best?
-
-% The study of effective educational strategies in use of GenAI tools in education
-% context are important for determining the best training for educators. However,
-% as these tools gain popularity among students \cite{Dickey24} the urgency to
-% equip educators with best practices may lag behind their rapid adoption.
-% Therefore, analyzing students' interaction patterns with GenAI tools is
-% important for understanding how these tools are being used and how educators can
-% leverage them to enhance learning. In addition, these interaction patterns can
-% create new opportunities to redesign GenAI tools to better support pedagogical
-% strategies without repressing the development of abstract, critical and creative
-% thinking.
 
@@
   \includegraphics[scale=0.39]{img/figure3.png}
-  \caption{Average interaction time per category.}
+  \caption{Subsections per conversation. \textit{A zero value indicates
+  no subsection and no spacing.}}
   \label{fig:graph3}
@@
 outlined in the introduction. Figure \ref{fig:graph1} shows the distribution of
-the messages in the categories.
+the messages in the categories. The following sections present the results and
+discussions of each research questions.
 
@@
 
-About the first research question: \textit{RQ1: What distribution patterns
+About the first research question: \textit{RQ1 - What distribution patterns
 emerged in the classified student interaction prompts with the educational
@@
 
-About the second research question: \textit{RQ2: Was there spacing between
-student prompts, and if so, which category led to the most rapid and consistent
-interactions with the bot?}
+About the first part of the second research question: \textit{RQ2 - Was there
+spacing between student prompts?}
 
@@
 \subsubsection{Interaction Time per Category}
+
+About the second part of the second research question: \textit{RQ2 - Which
+category led to the most rapid and consistent interactions with the bot?}
 
@@
 a more passive behavior, as students obtain ready-made answers from the bot.
-
-% This less active learning approach can occur for various reasons, ranging from
-% the belief that reviewing solutions to similar questions enhances learning, to
-% issues such as negligence, poor planning, or other possible factors.
 
@@
 
-About the third research question: \textit{RQ3: Do students' interactions with
+About the third research question: \textit{RQ3 - Do students' interactions with
 the bot alternate between study topics to suggest interleaving?}
@@
 
-The objective of this work was to analyze the interaction of second-semester
-Java programming students with an educational bot developed using the RAG
-technique, employing two metacognitive strategies: spacing and interleaving.
+The objective of this work was to analyze the interactions of Java programming
+students with an educational bot developed using the RAG technique, employing
+two metacognitive study strategies: spacing and interleaving.
 
-Initially, the theory of metacognition, alongside the classification of student
-actions, has proven effective in understanding and revealing how students
-utilize the bot. Although the data shows that most messages fall within the
-categories of Conceptual Questions and Debugging Help categories identified as
-promoting slower but potentially deeper engagement with the bot, indicative of a
-more active form of learning—the findings also reveal that students' use of
-spacing and interleaving strategies remains quite limited. Encouraging students
-to apply these strategies more intentionally could significantly enhance their
-learning outcomes. To further explore this, a broader study is planned to
-examine, for instance, the relationship between student profiles and their
-strategies in using a Generative AI tool. Additionally, to support educators
-in analyzing student behavior and providing more targeted assistance, an AI
-model is being developed to automatically classify student messages, offering
-valuable insights into the ways these tools are being employed.
+The theory of metacognition, along with the classification of student actions,
+proved effective in understanding and revealing how students utilize the bot.
+The results showed that most messages fall into the categories of Conceptual
+Questions and Debugging Help, both known for promoting slower engagement and
+potentially fostering more active learning. However, the results also indicated
+that students' use of study strategies such as spacing and interleaving remains
+limited, from which we can conclude that encouraging students to apply these
+strategies more intentionally could enhance their learning while using GenAI.
 
-% The results suggest that most student interactions with GenAI tools are centered
-% on active learning strategies, although the deliberate use of interleaving and
-% spacing techniques remains limited. Guiding students toward a more intentional
-% application of these strategies could enhance their learning outcomes. The
-% anonymity of interactions leaves it unclear whether these strategies were used
-% deliberately or emerged naturally, highlighting the need for future research on
-% students' conscious application of these methods. To support educators in
-% analyzing student behavior and offering more targeted assistance, an AI model
-% is being developed to automatically classify student messages, providing
-% valuable insights into how these tools are being utilized.
+To further explore this, a broader study is planned to examine, for instance,
+the relationship between student profiles, use of metacognitive scaffolds, their
+effective study strategies in using a GenAI and, learning outcomes.
+Additionally, to support educators in analyzing student behavior and providing
+more targeted assistance, an AI model is being developed to automatically
+classify student messages, offering valuable insights into how these tools are
+being used.
 
@@
 \end{document}
-
-% % Conceptual Question
-% \subsubsection*{Sequence of Conceptual Question}
-
-% Messages classified as conceptual question represent around 32\% of all
-% messages. We have found a pattern of sequences of Conceptual Question in some
-% conversations. Some conversations are entirely composed of messages classified
-% as Conceptual Question. In other cases, the sequence of Conceptual question
-% starts with a request for a Complete Solution or a Code Snippet, followed by
-% a progression of Conceptual Question. In both cases, the student is actively
-% engaging with the bot to understand the concepts and compromised with more
-% deep learning.
-
-% Regarding Metacognitive Knowledge, these sequences of prompts classified as
-% Conceptual Question, can be interpreted that the students are aware they have
-% not fully mastered a concept and are trying to understand it, demonstrating a
-% good level of self-awareness. In this sense, the student realizes that the
-% problem requires an understanding of concepts in order to advance in their
-% studies. In this way, the student applies an active learning strategy.
-
-% About Metacognitive Experience, several types of feelings may be associated at
-% this moment, such as actively reflecting on the answers, uncertainty and/or
-% confusion when trying to understand the concept, frustration for not
-% progressing, and satisfaction as their questions are answered in a way that
-% facilitates understanding.
-
-% Regarding the type of Metacognition related to objective of the Task, the
-% students may recognize the difficult and trying to achieve a deeper
-% understanding of the evolved concepts to later apply in practice.
-
-% With respect to Metacognitive Strategies, the student demonstrates a sence of
-% planning to identify their knowledge gaps. They may start with more general
-% questions to obtain an overview of the concept and then ask more specific
-% questions as they gain more clarity. Because the questions are in sequence, the
-% student probably can adequately monitor their progress.
-
-% % Multiple Choice
-% \subsubsection*{Sequence of Multiple Choice}
-
-% Around 16.9\% of the messages were classified as Multiple Choice. These prompts
-% typically involve students sending one or more exercises to CharlieBot, asking
-% for the solution.
-
-% We identified a sequence of Multiple Choice patterns in some of the
-% conversations. From the point of view of Metacognitive Knowledge, a possible
-% interpretation of this complex phenomenon is that students may believe in their
-% inability to answer several questions. They may also assume that the questions
-% share a common theme, and solving them together may help consolidate knowledge.
-% Additionally, they might understand that obtaining the answers quickly is more
-% efficient.
-
-% Regarding Metacognitive Experiences, students may experience feelings such as
-% cognitive overload, a sense of urgency, and/or confidence in ready-made answers,
-% which can often lead to relief and a reduction of frustration.
-
-% Concerning the Metacognitive about the Task, one interpretation may be that the
-% student simply wants to complete the task quickly to alleviate the workload,
-% and/or obtain a correct answer without focusing on deeper learning at
-% this moment.
-
-% In relation to Metacognitive Strategies, students are likely focusing on saving
-% time and/or possibly lacking planning skills. They may also perceive that
-% solving the questions on their own requires too much cognitive effort, but this
-% pattern might reveal a lack of progress monitoring and understanding of
-% concepts. After obtaining the answers, the metacognitive evaluation process
-% could lead to feelings of satisfaction, but further reflection may indicate that
-% the strategy was not effective in supporting deep learning. Depending on this
-% evaluation, it may influence future decision-making to change the strategy.
-
-% % Debugging help
-% \subsubsection*{Sequence of Debugging Help}
-
-% Around 30.9\% of the messages were classified as Debugging Help. Besides of
-% that, we identified a pattern of a sequences of Debugging Help in the
-% conversations. These prompts typically start with a message requesting a
-% Complete Solution or a Code Snippet. After that, the students normally chance
-% from passive to more active learning strategy.
-
-% In terms of Metacognitive Knowledge, the student probably recognizes that they
-% do not have the necessary tools or knowledge to debug the code on their own.
-% Concerning the understanding of the task, they may know that debugging is not
-% just a matter of superficial corrections but requires understanding the
-% underlying logic. Thus, the student apparently believes that the sequence of
-% questions related to debugging will help them understand the codes.
-
-% In terms of Metacognitive Experiences, some feelings may arise at this moment,
-% such as frustration or uncertainty in the face of errors, confusion during the
-% process, and also growing confidence or satisfaction as they better understand
-% the code and possible corrections.
-
-% With respect to Metacognitive related with the objective of the Task, the
-% student is probably focused on understanding the code in practical terms,
-% identifying errors, and learning how to better debug in the future.
-
-% As for Metacognitive associated with the Strategy, the student has requested
-% external help and realizes that dividing the debugging task into several
-% questions can result in a better understanding of the code. They can monitor
-% their progress throughout the process, adjusting their questions as necessary,
-% evaluate the effectiveness of the responses received, and reflect on what was
-% learned during the process.

### 2024-11-08 CSEDU2025/Paper.tex
update paper

@@
 \author{\authorname{Rodrigo Prestes Machado\sup{1}\orcidAuthor{0000-0003-0428-6387},
-Carlos Alario Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814},
-Patricia Callejo Pinardo\sup{2}\orcidAuthor{0000-0001-6124-6213},
+Carlos Alario-Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814},
+Patricia Callejo\sup{2}\orcidAuthor{0000-0001-6124-6213},
 Iria Estévez-Ayres\sup{2}\orcidAuthor{0000-0002-1047-5398},
@@
 de Madrid, Leganés (Madrid), Spain}
-\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario,pcallejo, ayres,cdk\}@it.uc3m.es}
+\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario, pcallejo, ayres, cdk\}@it.uc3m.es}
 }
@@
 course. The study investigates how students use the chatbot while employing
-two metacognitive strategies: interleaving and spacing. By analyzing the logs
-of student interactions with the chatbot, the methodology involved
-classifying and categorizing student prompts into eight distinct types. The
-findings reveal that, while most interactions focus on active learning
-strategies, students show limited conscious application of study strategies such
-as interleaving and spacing. The study highlights the need for more
-comprehensive guidance on the strategic use of AI tools to enhance learning
-outcomes.
+two metacognitive strategies: interleaving and spacing. The methodology involved
+classifying and categorizing the students' interaction prompts with the chatbot
+into eight distinct categories. The findings reveal that, while most
+interactions focus on active learning strategies, students demonstrate limited
+conscious application of metacognitive strategies. The study highlights the need
+for more comprehensive guidance on the strategic use of AI tools to enhance
+learning outcomes.
 }
@@
 postgraduate students have positive attitudes towards the use of GenAI.
-Furthermore, a systematic review organized by \cite{Lo24} demonstrated that
-students can effectively learn from ChatGPT, resulting in improved comprehension
-and academic achievement \citep{Callejo24}. Furthermore, it was observed that
-ChatGPT allows students to regulate their learning pace \citep{Baha24} and can
+A systematic review organized by \cite{Lo24} demonstrated that students can
+effectively learn from ChatGPT, resulting in improved comprehension and academic
+achievement \citep{Callejo24}. Furthermore, it was observed that ChatGPT may
+allow students to regulate their learning pace \citep{Baha24} and can
 potentially support self-regulated learning, particularly for students with
@@
 overdependence on ChatGPT could lead to a decrease in critical thinking, as
-students can make decisions based solely on the information provided by the tool.
-As a potential consequence, \cite{Bastani24} observed in a study that when
-programming students lost access to ChatGPT-4, those who had previously relied
+students can make decisions based solely on the information provided by the
+tool. As a potential consequence, \cite{Bastani24} observed in a study that when
+programming students lost access to ChatGPT, those who had previously relied
 on it saw their performance drop by 17\%. In contrast, students who had never
@@
 Regardless of teachers' preferences or beliefs, preliminary surveys conducted
-by \cite{Dickey24} indicate that more than 54,5\% of students are already using
+by \cite{Dickey24} indicate that more than 54.5\% of students are already using
 GenAI for homework, likely due to their perception of this technology as
@@
 the need to increase the understanding of how students interact with these tools
-and how they can be used to improve learning, as emphasized \cite{Lo24}.
+and how they can be used to improve learning, as emphasized by \cite{Lo24}.
 
@@
 the interactions between students in a Java programming course and an
-educational bot powered by OpenAI gpt-3.5, enhanced with the
+educational chatbot powered by OpenAI gpt-3.5, enhanced with the
 Retrieval-Augmented Generation (RAG) technique. Specifically, it focuses on
@@
   \item RQ2 - Was there spacing between student prompts? Which category led to
-  the most rapid and consistent interactions with the bot?
-  \item RQ3 - Do students' interactions with the bot alternate between study
+  the most rapid and consistent interactions with the chatbot?
+  \item RQ3 - Do students' interactions with the chatbot alternate between study
   topics to suggest interleaving?
 \end{itemize}
-
-% What is its main limitation?
-
-It is important to consider how GenAI tools impact students in
-various demographic groups, academic disciplines, cultural backgrounds, and
-levels of previous experience \citep{catalan21} \citep{neo22}. Consequently,
-this study is limited to a specific group of students and focuses on the use of
-a single GenAI tool. As a result, the findings may not be generalizable to other
-populations or tools.
 
@@
       Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. \\
-      Student Correction (SC) & Prompts where the student corrects the bot. \\
+      Student Correction (SC) & Prompts where the student corrects the chatbot. \\
       Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. \\
@@
 
-Given that generative AI often produces highly accurate and automatic responses
+Given that GenAI often produces highly accurate and automatic responses
 \citep{Puryear22}, it is essential that students use these tools within an
@@
 
-CharlieBot is an educational bot built on ChatGPT 3.5 and enhanced with
+CharlieBot is an educational chatbot built on ChatGPT 3.5 and enhanced with
 Retrieval-Augmented Generation (RAG) technique to improve the precision of
@@
 into eight distinct categories for analysis, representing different types of
-prompt that students use when interacting with the bot. Initially,
+prompt that students use when interacting with the chatbot. Initially,
 \cite{Ghimire24} proposed four categories: Debugging Help (DH), Conceptual
@@
 of four more: Multiple Question (MQ), Student Corrections (SC), Language Change
-(LC), and Uncategorized (U). In this study, students were free to use the bot as
-they wanted, which contributed to the emergence of these additional categories.
-Table \ref{tab:categories} presents these categories along with their respective
-descriptions. To facilitate the analysis of interleaving, the prompt topics were
-also categorized according to the course subjects, which are: (1) Java, (2)
-Object Orientation, (3) Testing, (4) Recursion, (5) Data Structures, and (6)
-Sorting and Searching Algorithms.
+(LC), and Uncategorized (U). In this study, students were free to use the
+chatbot as they wanted, which contributed to the emergence of these additional
+categories. Table \ref{tab:categories} presents these categories along with
+their respective descriptions. To facilitate the analysis of interleaving, the
+prompt topics were also categorized according to the course subjects, which are:
+(1) Java, (2) Object Orientation, (3) Testing, (4) Recursion, (5) Data
+Structures, and (6) Sorting and Searching Algorithms.
 
@@
 available at: https://github.com/rodrigoprestesmachado/uc3m.
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.39]{img/figure3.png}
-  \caption{Subsections per conversation. \textit{A zero value indicates
-  no subsection and no spacing.}}
-  \label{fig:graph3}
-\end{figure*}
 
@@
 
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.39]{img/figure3.png}
+  \caption{Subsections per conversation. \textit{A zero value indicates
+  no subsection and no spacing.}}
+  \label{fig:graph3}
+\end{figure*}
+
 \subsection{Categorization of messages}
@@
 illustrated in conversation 4 of Figure \ref{fig:graph2}, which presents
-examples of students' interactions with the bot. In other cases, the
+examples of students' interactions with the chatbot. In other cases, the
 conversation consists entirely of messages classified as Conceptual Question,
@@
 
-Requests for the bot to generate a Code Snippet or Complete Solution accounted
+Requests for the chatbot to generate a Code Snippet or Complete Solution accounted
 for 8.2\% of student messages. These prompts reflected a desire for more
@@
 
-Student corrections to bot responses accounted for 4.2\% of the interactions.
+Student corrections to chatbot responses accounted for 4.2\% of the interactions.
 In most cases, these corrections occurred when the student already had the
-correct answer to an exercise and identified an error in the bot response, as
-illustrated in conversation 2 in Figure \ref{fig:graph2}. These corrections
-highlight that, although the bot provides quick feedback, it is not always
+correct answer to an exercise and identified an error in the chatbot response,
+as illustrated in conversation 2 in Figure \ref{fig:graph2}. These corrections
+highlight that, although the chatbot provides quick feedback, it is not always
 accurate.
@@
 written by students were classified as Uncategorized. These messages include
-expressions of gratitude toward the bot, such as a simple '\textit{thanks}',
+expressions of gratitude toward the chatbot, such as a simple '\textit{thanks}',
 contextual statements like '\textit{I'm reviewing object orientation}', as well
@@
 Finally, 1.3\% of the messages were classified as Language Change. These prompts
-were typically requests for the bot to switch languages, as seen in conversation
-9 in Figure \ref{fig:graph2}.
+were typically requests for the chatbot to switch languages, as seen in
+conversation 9 in Figure \ref{fig:graph2}.
 
@@
 About the second part of the second research question: \textit{RQ2 - Which
-category led to the most rapid and consistent interactions with the bot?}
+category led to the most rapid and consistent interactions with the chatbot?}
 
 When a student requests a code snippet (CS), it is generally predictable that
-they will resume interaction with the bot within a short period of time, as
+they will resume interaction with the chatbot within a short period of time, as
 indicated by the mean (1.26) and standard deviation (1.74) of the Code Snippet
@@
 average response time and minimal variability may indicate that students adopt a
-more passive learning posture when requesting the bot to generate small pieces
+more passive learning posture when requesting the chatbot to generate small pieces
 of code.
@@
 exhibit a relatively low average time (1.65) for students to resume interaction
-with the bot. This variation occurs, in part, due to situations where students
+with the chatbot. This variation occurs, in part, due to situations where students
 end the conversation with an uncategorized message, such as a simple
@@
 
-When a student requests the bot to change the language (LC) in a response, the
+When a student requests the chatbot to change the language (LC) in a response, the
 new interaction happens quickly (1.89) with little variation in response time
@@
 Although still below the overall average, students take a moderate amount of
-time (2.65) to resume interaction with the bot after requesting the solution to
+time (2.65) to resume interaction with the chatbot after requesting the solution to
 a multiple-choice question (MQ). The variation (6.10) suggests that some
@@
 mean and standard deviation are below the overall average, this could indicate
-a more passive behavior, as students obtain ready-made answers from the bot.
+a more passive behavior, as students obtain ready-made answers from the chatbot.
 
@@
 deviation (7.33) indicates significant variation. In some cases, students are
-already familiar with the answer and only seek a quick correction from the bot;
-in others, they detect flaws in the bot's responses and need to carefully
+already familiar with the answer and only seek a quick correction from the chatbot;
+in others, they detect flaws in the chatbot's responses and need to carefully
 examine the subsequent answer to identify potential errors, which suggests a
@@
 Conceptual questions (CQ) generally result in a high average time (5.41) for
-students to re-engage with the bot. Moreover, the high standard deviation
+students to re-engage with the chatbot. Moreover, the high standard deviation
 (10.54) indicates an unpredictable variation in response times. This
@@
 student, whereas in others, lesser complexity allows for a quicker interaction
-with the bot.
+with the chatbot.
 
 Finally, a message classified as a Complete Solution request (SCO) exhibits the
-longest average time (6.55) for students to resume interaction with the bot.
+longest average time (6.55) for students to resume interaction with the chatbot.
 Additionally, the high standard deviation (10.14) highlights significant
@@
 About the third research question: \textit{RQ3 - Do students' interactions with
-the bot alternate between study topics to suggest interleaving?}
+the chatbot alternate between study topics to suggest interleaving?}
 
@@
 
-In conversations with the bot that did not include spacing (66.7\% of the total
+In conversations with the chatbot that did not include spacing (66.7\% of the total
 conversations), approximately 19.2\% of the messages resulted in a change of
@@
 
-In interactions with the bot that included spacing (33.3\% of the total
+In interactions with the chatbot that included spacing (33.3\% of the total
 conversations), approximately 31.4\% of the messages switched topics.
@@
 of this interweaving was caused by spacing. In other words, some students use
-the bot for quick consultations, asking a question and then returning hours
+the chatbot for quick consultations, asking a question and then returning hours
 later with another question on a different subject.
@@
 The objective of this work was to analyze the interactions of Java programming
-students with an educational bot developed using the RAG technique, employing
+students with an educational chatbot developed using the RAG technique, employing
 two metacognitive study strategies: spacing and interleaving.
@@
 The theory of metacognition, along with the classification of student actions,
-proved effective in understanding and revealing how students utilize the bot.
+proved effective in understanding and revealing how students utilize the chatbot.
 The results showed that most messages fall into the categories of Conceptual
@@
 strategies more intentionally could enhance their learning while using GenAI.
+
+It is important to consider how GenAI tools impact students in various
+demographic groups, academic disciplines, cultural backgrounds, and levels of
+previous experience \citep{catalan21} \citep{neo22}. Consequently, this study is
+limited to a specific group of students and focuses on the use of a single GenAI
+tool. As a result, the findings may not be generalizable to other populations or
+tools.
 

### 2024-11-08 code/rag/digital_culture/dc_create.py

@@ -1,0 +1,28 @@
+import sys
+import os
+from orion_pdf2text import OrionPdf2Text
+
+# Add the parent directory to sys.path
+sys.path.insert(0, os.path.abspath(os.path.join(os.path.dirname(__file__), '..')))
+
+# Now you can import the class from orion_text_chunker
+from orion_text_chunker import TextChunker
+from orion_rag_chroma_client import OrionChromaDBClient
+
+
+if __name__ == "__main__":
+    # Convert the manuscript to txt using the OrionPdf2Text class
+    PATH = "manuscript.pdf"
+    converter = OrionPdf2Text(PATH)
+    text = converter.convert_to_text()
+    if text:
+        # Save the wrapped text to a new .txt file
+        converter.save_text_to_file("output.txt", text)
+
+        # create chunks of text
+        chunker = TextChunker("output.txt")
+        chunker.split_into_chunks()
+        chunks = chunker.get_chunks()
+
+        client = OrionChromaDBClient("digital_culture")
+        client.add_chunks_to_db(chunks)

### 2024-11-08 code/rag/digital_culture/dc_search.py

@@ -1,0 +1,33 @@
+# Add the parent directory to sys.path
+import sys
+import os
+sys.path.insert(0, os.path.abspath(os.path.join(os.path.dirname(__file__), '..')))
+
+from orion_rag_chroma_client import OrionChromaDBClient
+
+if __name__ == "__main__":
+    # Specify the ChromaDB collection name
+    collection_name = "digital_culture"
+
+    # Create an instance of the OrionChromaDBClient class
+    client = OrionChromaDBClient(collection_name)
+
+    # Search for a term in the collection
+    search_results = client.search("competências digitais")
+    #print(f"Search results: {search_results}")
+
+    print(search_results["ids"][0][0])
+    print(search_results["distances"][0][0])
+    print(search_results["documents"][0][0])
+
+    print("-------------------")
+
+    print(search_results["ids"][0][1])
+    print(search_results["distances"][0][1])
+    print(search_results["documents"][0][1])
+
+    print("-------------------")
+
+    print(search_results["ids"][0][2])
+    print(search_results["distances"][0][2])
+    print(search_results["documents"][0][2])

### 2024-11-13 CSEDU2025/Paper.tex
update paper

@@
 \usepackage{arydshln}
+\widowpenalty=10000
+\clubpenalty=10000
 
@@
 the perceived usefulness of AI, and academic performance. The findings revealed
-that students with higher self-efficacy, lower fear of failure or higher prior
-grades tended to use AI less frequently or later in the problem solving process
+that students with higher self-efficacy, lower fear of failure, or higher prior
+grades tended to use AI less frequently or later in the problem-solving process
 and perceived it as less useful compared to their peers. However, no significant
@@
 of access to GPT Base led to a 17\% decrease in student performance on exams.
-This indicated that unrestricted use of *GPT Base* could hinder learning.
+This indicated that the unrestricted use of *GPT Base* could hinder learning.
 However, the GPT Tutor was able to mitigate this negative effect.
@@
 influence how students interacted with programming tasks, their learning
-outcomes, and their attitudes toward the technology. The study employed a
-quasi-experimental design involving 82 college students divided into two
-groups: one using ChatGPT-assisted programming (CFP) and the other using
-traditional self-directed programming (SDP). The method combined mixed data
-collection techniques, including behavioral logs, programming performance
-evaluations, and interviews. The results showed that students using the CFP mode
-exhibited more frequent behaviors such as debugging and reading feedback.
-Although CFP students had a higher average score in programming performance,
-there was no statistically significant difference compared to the SDP group
-However, students’ perceptions of ChatGPT improved significantly after the
-intervention, with increased perceived usefulness, ease of use, and intention to
-continue using the tool in the future.
+outcomes, and their attitudes towards technology. The study employed a
+quasi-experimental design involving 82 college students divided into two groups:
+one using ChatGPT-assisted programming (CFP) and the other using traditional
+self-directed programming (SDP). The method combined mixed data collection
+techniques, including behavioral logs, programming performance evaluations, and
+interviews. The results showed that students using the CFP mode exhibited more
+frequent behaviors such as debugging and reading feedback. Although CFP students
+had a higher average score in programming performance, there was no
+statistically significant difference compared to the SDP group. However,
+students’ perceptions of ChatGPT improved significantly after the intervention,
+with increased perceived usefulness, ease of use, and intention to continue
+using the tool in the future.
 
@@
 AI-enabled workforce. The problem discussed was how AI systems, especially
-collaborative tools like large language models (LLMs), impacted human
+collaborative tools like Large Language Models (LLMs), impacted human
 decision-making and the challenges they posed due to the lack of human-like
@@
 and improve decision-making when working with AI. The study concluded by
-recommending training and AI design improvements to enhance workers’
+recommending improvements in AI design and training to improve workers'
 metacognitive abilities, ensuring more effective collaboration with AI systems.
@@
 the efficacy of a meta-learning approach in improving metacognitive and creative
-skills. This study confirms the effectiveness of meta-learning strategies and
+skills. This study confirms the effectiveness of metacognition strategies and
 elucidates the relationship between meta-learning and metacognition. This
-research suggests that metalearning can improve metacognitive abilities,
+research suggests that metacognition can improve metacognitive abilities,
 providing valuable insights into educational technology and course design in
@@
 Retrieval-Augmented Generation (RAG) technique to improve the precision of
-responses provided to students \cite{Sun24}. The RAG approach was developed
+responses provided to students \citep{Sun24}. The RAG approach was developed
 using materials from a MOOC, including text, exercises, video transcripts, and
@@
 
-The study was conducted in three phases: data collection,
-categorization, and analysis. During the data collection phase, students
-enrolled in a second semester Java course at a public university in Spain were
-introduced to CharlieBot and allowed to use it freely, without being subjected
-to a specific methodology for its use. Data were collected anonymously, ensuring
-that the records of interactions could not be linked to any individual student
-or their academic performance.
+The study comprised three phases: data collection, categorization, and analysis.
+During the data collection phase, students enrolled in a second-semester Java
+course at a public university in Spain were introduced to CharlieBot and given
+the freedom to use it without following any prescribed educational methodology.
+All data were collected anonymously to ensure that interactions could not be
+traced back to individual students or linked to their academic performance.
+
+In the categorization phase, the students' messages were manually classified
+into eight distinct categories for analysis, representing different types of
+prompt that students use when interacting with the chatbot. Initially,
+\cite{Ghimire24} proposed four categories: Debugging Help (DH), Conceptual
+Question (CQ), Code Snippet (CS), and Complete Solution (CSO). However, the data
+collected indicated the need for additional categories, leading to the inclusion
+of four more: Multiple Question (MQ), Student Corrections (SC), Language Change
+(LC), and Uncategorized (U). In this study, students were free to use the
+chatbot as they wanted, which contributed to the emergence of these additional
+categories. Table \ref{tab:categories} presents these categories along with
+their respective descriptions. To facilitate the analysis, the prompt topics
+were also categorized according to the course subjects, which are: (1) Java, (2)
+Object Orientation, (3) Testing, (4) Recursion, (5) Data Structures, and (6)
+Sorting and Searching Algorithms.
+
+The analysis phase involved the use of Python scripts to extract information
+from previously classified data. The datasets and code used in this process are
+available at: https://github.com/rodrigoprestesmachado/uc3m.
+
+\section{\uppercase{Results and Discussion}}
+
+A total of 625 student messages were categorized in 81 conversations, with
+an average of 7.72 messages per conversation. This data offers insights into how
+users engage with CharlieBot and the types of prompts they submit. The results
+of the analysis are presented in this section, addressing the research questions
+outlined in the introduction. Figure \ref{fig:graph1} shows the distribution of
+the messages in the categories. The following sections present the results and
+discussions of each research question.
+
+\subsection{Categorization of messages}
+
+About the first research question: \textit{RQ1 - What distribution patterns
+emerged in the classified student interaction prompts with the educational
+chatbot?}
+
+According to Figure \ref{fig:graph1} messages classified as
+Conceptual Question account for 35.8\% of the total. Sequences of Conceptual
+Questions often start from a practical perspective of a given code, as
+illustrated in conversation 4 of Figure \ref{fig:graph2}, which presents
+examples of students' interactions with the chatbot. In other cases, the
+conversation consists entirely of messages classified as Conceptual Question,
+such as in conversation 15 of Figure \ref{fig:graph2}.
 
@@
 
-In the categorization phase, the students' messages were manually classified
-into eight distinct categories for analysis, representing different types of
-prompt that students use when interacting with the chatbot. Initially,
-\cite{Ghimire24} proposed four categories: Debugging Help (DH), Conceptual
-Question (CQ), Code Snippet (CS), and Complete Solution (CSO). However, the data
-collected indicated the need for additional categories, leading to the inclusion
-of four more: Multiple Question (MQ), Student Corrections (SC), Language Change
-(LC), and Uncategorized (U). In this study, students were free to use the
-chatbot as they wanted, which contributed to the emergence of these additional
-categories. Table \ref{tab:categories} presents these categories along with
-their respective descriptions. To facilitate the analysis of interleaving, the
-prompt topics were also categorized according to the course subjects, which are:
-(1) Java, (2) Object Orientation, (3) Testing, (4) Recursion, (5) Data
-Structures, and (6) Sorting and Searching Algorithms.
-
-The analysis phase involved the use of Python scripts to extract information
-from previously classified data. The datasets and code used in this process are
-available at: https://github.com/rodrigoprestesmachado/uc3m.
-
-\section{\uppercase{Results and Discussion}}
-
-A total of 625 student messages were categorized in 81 conversations, with
-an average of 7.72 messages per conversation. This data offers insights into how
-users engage with CharlieBot and the types of prompts they submit. The results
-of the analysis are presented in this section, addressing the research questions
-outlined in the introduction. Figure \ref{fig:graph1} shows the distribution of
-the messages in the categories. The following sections present the results and
-discussions of each research questions.
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.39]{img/figure3.png}
-  \caption{Subsections per conversation. \textit{A zero value indicates
-  no subsection and no spacing.}}
-  \label{fig:graph3}
-\end{figure*}
-
-\subsection{Categorization of messages}
-
-About the first research question: \textit{RQ1 - What distribution patterns
-emerged in the classified student interaction prompts with the educational
-chatbot?}
-
-According to Figure \ref{fig:graph1} messages classified as
-Conceptual Question account for 35.8\% of the total. Sequences of Conceptual
-Questions often start from a practical perspective of a given code, as
-illustrated in conversation 4 of Figure \ref{fig:graph2}, which presents
-examples of students' interactions with the chatbot. In other cases, the
-conversation consists entirely of messages classified as Conceptual Question,
-such as in conversation 15 of Figure \ref{fig:graph2}.
-
 As shown in Figure \ref{fig:graph1}, approximately 29.9\% of the messages were
@@
 
-Most of the messages are located in categories Conceptual Question and Debugging
-Help corroborate the findings of \cite{Ghimire24} which showed that the
-questions are also localized in the same categories.
+Most of the messages are located in the categories Conceptual Question and
+Debugging Help corroborate the findings of \cite{Ghimire24} which showed that
+the questions are also localized in the same categories.
 
@@
 These prompts typically involved students submitting one or more exercises to
-CharlieBot and requesting solutions. The conversation 13 in \ref{fig:graph2}
-illustrates a sequence of multiple choice questions.
+CharlieBot and requesting solutions. The conversation 13 in Figure
+\ref{fig:graph2} illustrates a sequence of multiple choice questions.
 
@@
 Figure \ref{fig:graph1} also indicates that approximately 5.8\% of the prompts
-written by students were classified as Uncategorized. These messages include
+written by the students were classified as Uncategorized. These messages include
 expressions of gratitude toward the chatbot, such as a simple '\textit{thanks}',
@@
 conversation 9 in Figure \ref{fig:graph2}.
-
-\begin{table*}[htbp]
-  \caption{Average and Standard Deviation of Different Categories (in minutes)}
-  \begin{center}
-    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{4cm} p{4cm} p{4cm}} % Remove vertical bars
-      \hline
-      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
-      \hline
-      Code Snippet & 1.26 & 1.74 \\
-      Uncategorized & 1.65 & 3.54 \\
-      Language Change & 1.89 & 3.28 \\
-      Multiple Question & 2.65 & 6.10 \\
-      Student Correction & 3.17 & 7.33 \\
-      \hdashline
-      \textit{Overall} & \textit{3.43} & \textit{6.60} \\
-      \hdashline
-      Debugging Help & 4.84 & 10.13 \\
-      Conceptual Questions & 5.41 & 10.54 \\
-      Complete Solution & 6.55 & 10.14 \\
-      \hline
-    \end{tabular}
-    \label{tab:averages}
-  \end{center}
-\end{table*}
 
@@
 
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.39]{img/figure3.png}
+  \caption{Subsections per conversation. \textit{A zero value indicates
+  no subsection and no spacing.}}
+  \label{fig:graph3}
+\end{figure*}
+
 Despite some occasional variation (3.54), prompts classified as Uncategorized
@@
 end the conversation with an uncategorized message, such as a simple
-\textit{'thanks'} but later choose to resume the interaction.
+\textit{"thanks"} but later choose to resume the interaction.
 
@@
 
+\begin{table*}[htbp]
+  \caption{Average and Standard Deviation of Different Categories (in minutes)}
+  \begin{center}
+    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
+    \begin{tabular}{p{4cm} p{4cm} p{4cm}} % Remove vertical bars
+      \hline
+      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
+      \hline
+      Code Snippet & 1.26 & 1.74 \\
+      Uncategorized & 1.65 & 3.54 \\
+      Language Change & 1.89 & 3.28 \\
+      Multiple Question & 2.65 & 6.10 \\
+      Student Correction & 3.17 & 7.33 \\
+      \hdashline
+      \textit{Overall} & \textit{3.43} & \textit{6.60} \\
+      \hdashline
+      Debugging Help & 4.84 & 10.13 \\
+      Conceptual Questions & 5.41 & 10.54 \\
+      Complete Solution & 6.55 & 10.14 \\
+      \hline
+    \end{tabular}
+    \label{tab:averages}
+  \end{center}
+\end{table*}
+
 \section{Conclusion and Future Work}

### 2024-11-14 CSEDU2025/Paper.tex
update paper

@@
 
-This section presents the theoretical framework that underpins the study. First,
-we introduce the concept of metacognition used in this study. Next, we discuss
-related work exploring the use of metacognitive strategies and scaffolds
-in educational settings.
-
-\begin{table*}[htbp]
-  \caption{Categories - adapted from \cite{Ghimire24}}
-  \begin{center}
-    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{4cm} p{12cm}} % Remove vertical bars
-      \hline
-      \textbf{Category} & \textbf{Description} \\
-      \hline
-      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
-      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. \\
-      Student Correction (SC) & Prompts where the student corrects the chatbot. \\
-      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. \\
-      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. \\
-      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
-      Language Change (LC) & Prompts that request a change of idiom. \\
-      Uncategorized (U) & Prompts that do not fit into any of the above categories. \\
-      \hline
-    \end{tabular}
-    \label{tab:categories}
-  \end{center}
-\end{table*}
+This section outlines the theoretical framework that grounds the study. We
+begin by introducing the concept of metacognition, along with the spacing and
+interweaving strategies applied in this research. Next, we examine related works
+ organized into two subsections: the use of GenAI in programming and recent
+ studies emphasizing the importance of metacognition in the use of GenAI.
 
@@
 
-\subsection{Related Work}
+\subsection{GenAI and Programming}
 
@@
 
-\cite{Bastani24} investigated the impact of generative artificial intelligence
-(AI), specifically GPT-4, on human learning, with a focus on mathematics
-education in a high school. The problem addressed was how the use of generative
-AI could affect the acquisition of new skills, which is crucial for long-term
-productivity. The method involved a controlled experiment with approximately
-one thousand students, who were exposed to two GPT-4-based tutors: a simple
-tutor (GPT Base) and another with safeguards designed to promote learning
-(GPT Tutor). The results showed that while access to GPT-4 improved performance
-on practice exercises (48\% with GPT Base and 127\% with GPT Tutor), the removal
-of access to GPT Base led to a 17\% decrease in student performance on exams.
-This indicated that the unrestricted use of *GPT Base* could hinder learning.
-However, the GPT Tutor was able to mitigate this negative effect.
+\cite{Bastani24} investigated the impact of GenAI, specifically GPT-4, on human
+learning, with a focus on mathematics education in a high school. The problem
+addressed was how the use of generative AI could affect the acquisition of new
+skills, which is crucial for long-term productivity. The method involved a
+controlled experiment with approximately one thousand students, who were exposed
+to two GPT-4-based tutors: a simple tutor (GPT Base) and another with safeguards
+designed to promote learning (GPT Tutor). The results showed that while access
+to GPT-4 improved performance on practice exercises (48\% with GPT Base and
+127\% with GPT Tutor), the removal of access to GPT Base led to a 17\% decrease
+in student performance on exams. This suggests that unrestricted use of
+\textit{GPT Base} could hinder learning. However, the GPT Tutor was able to
+mitigate this negative effect.
 
@@
 quasi-experimental design involving 82 college students divided into two groups:
-one using ChatGPT-assisted programming (CFP) and the other using traditional
-self-directed programming (SDP). The method combined mixed data collection
+one using ChatGPT-assisted Programming (CAP) and the other using traditional
+self-directed Programming (SDP). The method combined mixed data collection
 techniques, including behavioral logs, programming performance evaluations, and
-interviews. The results showed that students using the CFP mode exhibited more
-frequent behaviors such as debugging and reading feedback. Although CFP students
+interviews. The results showed that students using the CAP mode exhibited more
+frequent behaviors such as debugging and reading feedback. Although CAP students
 had a higher average score in programming performance, there was no
@@
 using the tool in the future.
+
+\begin{table*}[htbp]
+  \caption{Categories - adapted from \cite{Ghimire24}}
+  \begin{center}
+    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
+    \begin{tabular}{p{4cm} p{12cm}} % Remove vertical bars
+      \hline
+      \textbf{Category} & \textbf{Description} \\
+      \hline
+      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
+      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. \\
+      Student Correction (SC) & Prompts where the student corrects the chatbot. \\
+      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. \\
+      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. \\
+      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
+      Language Change (LC) & Prompts that request a change of idiom. \\
+      Uncategorized (U) & Prompts that do not fit into any of the above categories. \\
+      \hline
+    \end{tabular}
+    \label{tab:categories}
+  \end{center}
+\end{table*}
+
+\subsection{GenAI and Metacognition}
 
@@
 The analysis phase involved the use of Python scripts to extract information
-from previously classified data. The datasets and code used in this process are
+from previously classified data. The data sets and code used in this process are
 available at: https://github.com/rodrigoprestesmachado/uc3m.
@@
 
-Student corrections to chatbot responses accounted for 4.2\% of the interactions.
+Student corrections to chatbot responses represented 4.2\% of the interactions.
 In most cases, these corrections occurred when the student already had the
@@
 exhibit a relatively low average time (1.65) for students to resume interaction
-with the chatbot. This variation occurs, in part, due to situations where students
+with the chatbot. This variation occurs, in part, because of situations where students
 end the conversation with an uncategorized message, such as a simple
-\textit{"thanks"} but later choose to resume the interaction.
+\textit{"thanks"}, but later choose to resume the interaction.
 
 When a student requests the chatbot to change the language (LC) in a response, the
-new interaction happens quickly (1.89) with little variation in response time
-(3.28), meaning it is predictable that the student will continue interacting.
+new interaction occurs quickly (1.89) with little variation in response time
+(3.28), meaning that it is predictable that the student will continue interacting.
 
@@
 Conceptual questions (CQ) generally result in a high average time (5.41) for
-students to re-engage with the chatbot. Moreover, the high standard deviation
+students to reengage with the chatbot. Moreover, the high standard deviation
 (10.54) indicates an unpredictable variation in response times. This
@@
 unpredictability in response time. After requesting a complete solution,
-students tend to take longer to return, due to the complexity of the problems
-involved, such as the implementation of sorting algorithms like Heapsort. Table
+students tend to take longer to return due to the complexity of the problems
+involved, such as the implementation of sorting algorithms such as Heapsort. Table
 \ref{tab:averages} presents the means and standard deviations for each
@@
 approaches and depths in the learning process.
+
+\begin{table*}[htbp]
+  \caption{Average and Standard Deviation of Different Categories (in minutes)}
+  \begin{center}
+    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
+    \begin{tabular}{p{4cm} p{4cm} p{4cm}} % Remove vertical bars
+      \hline
+      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
+      \hline
+      Code Snippet & 1.26 & 1.74 \\
+      Uncategorized & 1.65 & 3.54 \\
+      Language Change & 1.89 & 3.28 \\
+      Multiple Question & 2.65 & 6.10 \\
+      Student Correction & 3.17 & 7.33 \\
+      \hdashline
+      \textit{Overall} & \textit{3.43} & \textit{6.60} \\
+      \hdashline
+      Debugging Help & 4.84 & 10.13 \\
+      Conceptual Questions & 5.41 & 10.54 \\
+      Complete Solution & 6.55 & 10.14 \\
+      \hline
+    \end{tabular}
+    \label{tab:averages}
+  \end{center}
+\end{table*}
 
@@
 
-\begin{table*}[htbp]
-  \caption{Average and Standard Deviation of Different Categories (in minutes)}
-  \begin{center}
-    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{4cm} p{4cm} p{4cm}} % Remove vertical bars
-      \hline
-      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
-      \hline
-      Code Snippet & 1.26 & 1.74 \\
-      Uncategorized & 1.65 & 3.54 \\
-      Language Change & 1.89 & 3.28 \\
-      Multiple Question & 2.65 & 6.10 \\
-      Student Correction & 3.17 & 7.33 \\
-      \hdashline
-      \textit{Overall} & \textit{3.43} & \textit{6.60} \\
-      \hdashline
-      Debugging Help & 4.84 & 10.13 \\
-      Conceptual Questions & 5.41 & 10.54 \\
-      Complete Solution & 6.55 & 10.14 \\
-      \hline
-    \end{tabular}
-    \label{tab:averages}
-  \end{center}
-\end{table*}
-
 \section{Conclusion and Future Work}
@@
 
-The theory of metacognition, along with the classification of student actions,
-proved effective in understanding and revealing how students utilize the chatbot.
+The metacognition theory, along with the classification of student actions,
+proved to be effective in understanding and revealing how students use the chatbot.
 The results showed that most messages fall into the categories of Conceptual
@@
 the relationship between student profiles, use of metacognitive scaffolds, their
-effective study strategies in using a GenAI and, learning outcomes.
-Additionally, to support educators in analyzing student behavior and providing
+effective study strategies in using a GenAI, and learning outcomes.
+In addition, to help educators analyze student behavior and provide
 more targeted assistance, an AI model is being developed to automatically
-classify student messages, offering valuable insights into how these tools are
-being used.
+classify student messages, offering valuable information on how these tools are used.
 

### 2024-11-14 IEEETransactions/PaperIEEE.tex
update paper

\documentclass[lettersize,journal]{IEEEtran}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=normalsize,labelfont=sf,textfont=sf]{subfig}
\usepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\hyphenation{op-tical net-works semi-conduc-tor IEEE-Xplore}
\def\BibTeX{{\rm B\kern-.05em{\sc i\kern-.025em b}\kern-.08em
    T\kern-.1667em\lower.7ex\hbox{E}\kern-.125emX}}
\usepackage{balance}
\begin{document}
\title{Analyzing Student Use of Spacing and Interleaving Strategies in
Interactions with GenAI-Powered Chatbots in Programming Courses}
\author{IEEE Publication Technology Department
\thanks{Manuscript created May, 2025; This work was developed by the IEEE Publication Technology Department. This work is distributed under the \LaTeX \ Project Public License (LPPL) ( http://www.latex-project.org/ ) version 1.3. A copy of the LPPL, version 1.3, is included in the base \LaTeX \ documentation of all distributions of \LaTeX \ released 2003/12/01 or later. The opinions expressed here are entirely that of the author. No warranty is expressed or implied. User assumes all risk.}}

\markboth{Journal of \LaTeX\ Class Files,~Vol.~18, No.~9, September~2020}%
{How to Use the IEEEtran \LaTeX \ Templates}

\maketitle

\begin{abstract}
The aim of this study is to analyze the interactions of programming students
with a chatbot powered by OpenAI's GPT-3.5, enhanced with the Retrieval
Augmented Generation (RAG) technique, within the context of a Java programming
course. The study investigates how students use the chatbot while employing
two metacognitive strategies: interleaving and spacing. The methodology involved
classifying and categorizing the students' interaction prompts with the chatbot
into eight distinct categories. The findings reveal that, while most
interactions focus on active learning strategies, students demonstrate limited
conscious application of metacognitive strategies. The study highlights the need
for more comprehensive guidance on the strategic use of AI tools to enhance
learning outcomes.
\end{abstract}

\begin{IEEEkeywords}
Class, IEEEtran, \LaTeX, paper, style, template, typesetting.
\end{IEEEkeywords}

\section{Introduction}
\IEEEPARstart{W}{elcome} to the updated and simplified documentation to using the IEEEtran \LaTeX \ class file. The IEEE has examined hundreds of author submissions using this package to help formulate this easy to follow guide. We will cover the most commonly used elements of a journal article. For less common elements we will refer back to the ``IEEEtran\_HOWTO.pdf''.

This document applies to version 1.8b of IEEEtran.

The IEEEtran template package contains the following example files: 
\begin{list}{}{}
\item{bare\_jrnl.tex}
\item{bare\_conf.tex}
\item{bare\_jrnl\_compsoc.tex}
\item{bare\_conf\_compsoc.tex}
\item{bare\_jrnl\_comsoc.tex}
\end{list}
These are ``bare bones" templates to quickly understand the document structure.

It is assumed that the reader has a basic working knowledge of \LaTeX. Those who are new to \LaTeX \ are encouraged to read Tobias Oetiker's ``The Not So Short Introduction to \LaTeX '', available at: \url{http://tug.ctan.org/info/lshort/english/lshort.pdf} which provides an overview of working with \LaTeX.

\section{The Design, Intent and \\ Limitations of the Templates}
\noindent The templates are intended to {\bf{approximate the final look and page length of the articles/papers}}. Therefore, {\bf{they are NOT intended to be the final produced work that is displayed in print or on IEEEXplore\textsuperscript{\textregistered}}}. They will help to give the authors an approximation of the number of pages that will be in the final version. The structure of the \LaTeX files, as designed, enable easy conversion to XML for the composition systems used by the IEEE's outsource vendors. The XML files are used to produce the final print/IEEEXplore\textsuperscript{\textregistered} pdf and then converted to HTML for IEEEXplore\textsuperscript{\textregistered}. Have you looked at your article/paper in the HTML version?

\section{\LaTeX \ Distributions: Where to Get Them}
\noindent IEEE recommends using the distribution from the \TeX User Group at \url{http://www.tug.org}. You can join TUG and obtain a DVD distribution or download for free  from the links provided on their website: \url{http://www.tug.org/texlive/}. The DVD includes distributions for Windows, Mac OS X and Linux operating systems.

\section{Where to get the IEEEtran Templates}
\noindent The {\bf{IEEE Template Selector}} will always have the most up-to-date versions of the \LaTeX\ and MSWord templates. Please see: \url{https://template-selector.ieee.org/} and follow the steps to find the correct template for your intended publication. Many publications use the IEEETran LaTeX templates, however, some publications have their own special templates. Many of these are  based on IEEEtran, but may have special instructions that vary slightly from those in this document.

\section{Where to get \LaTeX \ help - user groups}
\noindent The following on-line groups are very helpful to beginning and experienced \LaTeX\ users. A search through their archives can provide many answers to common questions.
\begin{list}{}{}
\item{\url{http://www.latex-community.org/}} 
\item{\url{https://tex.stackexchange.com/} }
\end{list}

\section{Document Class Options in IEEEtran}
\noindent At the beginning of your \LaTeX\ file you will need to establish what type of publication style you intend to use. The following list shows appropriate documentclass options for each of the types covered by IEEEtran.

\begin{list}{}{}
\item{Regular Journal Article}
\item{{\tt{$\backslash$documentclass[journal]{IEEEtran}}}}\\
\item{{Conference Paper}}
\item{{\tt{$\backslash$documentclass[conference]{IEEEtran}}}}\\
\item{Computer Society Journal Article}
\item{{\tt{$\backslash$documentclass[10pt,journal,compsoc]{IEEEtran}}}}\\
\item{Computer Society Conference Paper}
\item{{\tt{$\backslash$documentclass[conference,compsoc]{IEEEtran}}}}\\
\item{{Communications Society Journal Article}}
\item{{\tt{$\backslash$documentclass[journal,comsoc]{IEEEtran}}}}\\
\item{{Brief, Correspondence or Technote}}
\item{{\tt{$\backslash$documentclass[9pt,technote]{IEEEtran}}}}
\end{list}

There are other options available for each of these when submitting for peer review or other special requirements. IEEE recommends to compose your article in the base 2-column format to make sure all your equations, tables and graphics will fit the final 2-column format. Please refer to the document ``IEEEtran\_HOWTO.pdf'' for more information on settings for peer review submission if required by your EIC.

\section{How to Create Common Front Matter}
\noindent The following sections describe general coding for these common elements. Computer Society publications and Conferences may have their own special variations and will be noted below.
\subsection{Paper Title}
\noindent The title of your paper is coded as:

\begin{verbatim}
\title{The Title of Your Paper}
\end{verbatim}

\noindent Please try to avoid the use of math or chemical formulas in your title if possible.

\subsection{Author Names and Affiliations}
\noindent The author section should be coded as follows:
\begin{verbatim}
\author{Masahito Hayashi 
\IEEEmembership{Fellow, IEEE}, Masaki Owari
\thanks{M. Hayashi is with Graduate School 
of Mathematics, Nagoya University, Nagoya, 
Japan}
\thanks{M. Owari is with the Faculty of 
Informatics, Shizuoka University, 
Hamamatsu, Shizuoka, Japan.}
}
\end{verbatim}
Be sure to use the $\backslash$IEEEmembership command to identify IEEE membership status.
Please see the ``IEEEtran\_HOWTO.pdf'' for specific information on coding authors for Conferences and Computer Society publications. Note that the closing curly brace for the author group comes at the end of the thanks group. This will prevent you from creating a blank first page.

\subsection{Running Heads}
\noindent The running heads are declared by using the $\backslash${\tt{markboth}} command. There are two arguments to this command: the first contains the journal name information and the second contains the author names and paper title.
\begin{verbatim}
\markboth{Journal of Quantum Electronics, 
Vol. 1, No. 1, January 2021}
{Author1, Author2, 
\MakeLowercase{\textit{(et al.)}: 
Paper Title}
\end{verbatim}

\subsection{Copyright Line}
\noindent For Transactions and Journals papers, this is not necessary to use at the submission stage of your paper. The IEEE production process will add the appropriate copyright line. If you are writing a conference paper, please see the ``IEEEtran\_HOWTO.pdf'' for specific information on how to code "Publication ID Marks".

\subsection{Abstracts}
\noindent The abstract is the first element of a paper after the $\backslash${\tt{maketitle}} macro is invoked.  The coding is simply:
\begin{verbatim}
\begin{abstract}
Text of your abstract.
\end{abstract}
\end{verbatim}
Please try to avoid mathematical and chemical formulas in the abstract.

\subsection{Index Terms}
\noindent The index terms are used to help other researchers discover your paper. Each society may have it's own keyword set. Contact the EIC of your intended publication for this list.
\begin{verbatim}
\begin{IEEEkeywords}
Broad band networks, quality of service
\end{IEEEkeywords}
\end{verbatim}
\section{How to Create Common Body Elements}
\noindent The following sections describe common body text elements and how to code them.

\subsection{Initial Drop Cap Letter}
\noindent The first text paragraph uses a ``drop cap'' followed by the first word in ALL CAPS. This is accomplished by using the $\backslash${\tt{IEEEPARstart}} command as follows:
\begin{verbatim}
\IEEEPARstart{T}{his} is the first paragraph 
of your paper. . .
\end{verbatim}

\subsection{Sections and Subsections}
\noindent Section headings use standard \LaTeX\ commands: $\backslash${\tt{section}}, $\backslash${\tt{subsection}} and $\backslash${\tt{subsubsection}}. Numbering is handled automatically for you and varies according to type of publication. It is common to not indent the first paragraph following a section head by using $\backslash${\tt{noindent}} as follows:
\begin{verbatim}
\section{Section Head}
\noindent The text of your paragraph . . .
\end{verbatim}

\subsection{Citations to the Bibliography}
\noindent The coding for the citations are made with the \LaTeX\ $\backslash${\tt{cite}} command. This will produce individual bracketed reference numbers in the IEEE style. At the top of your \LaTeX\ file you should include:
\begin{verbatim}
\usepackage{cite}
\end{verbatim}
For a single citation code as follows:
\begin{verbatim}
see \cite{ams}
\end{verbatim}
This will display as: see \cite{ams}\\

For multiple citations code as follows:
\begin{verbatim}
\cite{ams,oxford,lacomp}
\end{verbatim}

This will display as \cite{ams,oxford,lacomp}

\subsection{Figures}
\noindent Figures are coded with the standard \LaTeX\ commands as follows:
\begin{verbatim}
\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{fig1}
\caption{This is the caption for one fig.}
\label{fig1}
\end{figure}
\end{verbatim}
The [!t] argument enables floats to the top of the page to follow IEEE style. Make sure you include:
\begin{verbatim}
\usepackage{graphicx}
\end{verbatim}

\noindent at the top of your \LaTeX file with the other package declarations.

To cross-reference your figures in the text use the following code example:
\begin{verbatim}
See figure \ref{fig1} ...
\end{verbatim}
This will produce:\\
See figure \ref{fig1} . . .

\begin{figure}[!t]
\centering
\includegraphics[width=2.5in]{fig1}
\caption{This is the caption for one fig.}
\label{fig1}
\end{figure}

\subsection{Tables}
\noindent Tables should be coded with the standard \LaTeX\ coding. The following example shows a simple table.

\begin{verbatim}
\begin{table}
\begin{center}
\caption{Filter design equations  ...}
\label{tab1}
\begin{tabular}{| c | c | c |}
\hline
Order & Arbitrary coefficients & 
coefficients\\
of filter & $e_m$ &   $b_{ij}$ \\
\hline
1& $b_{ij}=\hat{e}.\hat{\beta_{ij}}$, 
& $b_{00}=0$\\
\hline
2&$\beta_{22}=(~1,-1,-1,~~1,~~1,~~1)$ &\\ 
\hline
3& $b_{ij}=\hat{e}.\hat{\beta_{ij}}$, 
& $b_{00}=0$,\\
\hline 
\end{tabular}
\end{center}
\end{table}
\end{verbatim}
To reference the table in the text, code as follows:
\begin{verbatim}Table~\ref{tab1} lists the closed-form...\end{verbatim}
to produce:

Table~\ref{tab1} lists the closed-form . . .

%moved here for pagination purposes
\begin{table}
\begin{center}
\caption{A Simple Table Example.}
\label{tab1}
\begin{tabular}{| c | c | c |}
\hline
Order & Arbitrary coefficients & coefficients\\
of filter & $e_m$ &   $b_{ij}$ \\
\hline
1& $b_{ij}=\hat{e}.\hat{\beta_{ij}}$, & $b_{00}=0$\\
\hline
2&$\beta_{22}=(~1,-1,-1,~~1,~~1,~~1)$ &\\ 
\hline
3& $b_{ij}=\hat{e}.\hat{\beta_{ij}}$, & $b_{00}=0$,\\
\hline 
\end{tabular}
\end{center}
\end{table}

\subsection{Lists}
\noindent In this section, we will consider three types of lists: simple unnumbered, numbered and bulleted. There have been numerous options added to IEEEtran to enhance the creation of lists. If your lists are more complex than those shown below, please refer to the  ``IEEEtran\_HOWTO.pdf'' for additional options.\\

\noindent{\bf A plain  unnumbered list}

\begin{list}{}{}
\item{bare\_jrnl.tex}
\item{bare\_conf.tex}
\item{bare\_jrnl\_compsoc.tex}
\item{bare\_conf\_compsoc.tex}
\item{bare\_jrnl\_comsoc.tex}
\end{list}

\noindent coded as:
\begin{verbatim}
\begin{list}{}{}
\item{bare\_jrnl.tex}
\item{bare\_conf.tex}
\item{bare\_jrnl\_compsoc.tex}
\item{bare\_conf\_compsoc.tex}
\item{bare\_jrnl\_comsoc.tex}
\end{list}
\end{verbatim}
\noindent{\bf A simple numbered list}

\begin{enumerate}
\item{bare\_jrnl.tex}
\item{bare\_conf.tex}
\item{bare\_jrnl\_compsoc.tex}
\item{bare\_conf\_compsoc.tex}
\item{bare\_jrnl\_comsoc.tex}
\end{enumerate}
\noindent coded as: 
\begin{verbatim}
\begin{enumerate}
\item{bare\_jrnl.tex}
\item{bare\_conf.tex}
\item{bare\_jrnl\_compsoc.tex}
\item{bare\_conf\_compsoc.tex}
\item{bare\_jrnl\_comsoc.tex}
\end{enumerate}
\end{verbatim}

\noindent{\bf A simple bulleted list}

\begin{itemize}
\item{bare\_jrnl.tex}
\item{bare\_conf.tex}
\item{bare\_jrnl\_compsoc.tex}
\item{bare\_conf\_compsoc.tex}
\item{bare\_jrnl\_comsoc.tex}
\end{itemize}

\noindent coded as:

\begin{verbatim}
\begin{itemize}
\item{bare\_jrnl.tex}
\item{bare\_conf.tex}
\item{bare\_jrnl\_compsoc.tex}
\item{bare\_conf\_compsoc.tex}
\item{bare\_jrnl\_comsoc.tex}
\end{itemize}
\end{verbatim}

\subsection{Other Elements}
\noindent For other less common elements such as Algorithms, Theorems and Proofs, and Floating Structures such as page-wide tables, figures or equations, please refer to the ``IEEEtran\_HOWTO.pdf'' section on ``Double Column Floats.''

\section{How to Create Common Back Matter Elements}
\noindent The following sections demonstrate common back matter elements such as Acknowledgments, Bibliographies, Appendicies and Author Biographies.

\subsection{Acknowledgments}
\noindent This should be a simple paragraph before the bibliography to thank those individuals and institutions who have supported your work on this article.

\begin{verbatim}
\section{Acknowledgments}
\noindent Text describing those who 
supported your paper.
\end{verbatim}

\subsection{Bibliographies}
\noindent {\bf{References Simplified:}} A simple way of composing references is to use the $\backslash${\tt{bibitem}} macro to define the beginning of a reference as in the following examples:\\

\noindent [6] H. Sira-Ramirez. ``On the sliding mode control of nonlinear systems,'' \textit{Systems \& Control Letters}, vol. 19, pp. 303--312, 1992.

\noindent coded as:
\begin{verbatim}
\bibitem{Sira3}
H. Sira-Ramirez. ``On the sliding mode 
control of nonlinear systems,'' 
\textit{Systems \& Control Letters}, 
vol. 19, pp. 303--312, 1992.
\end{verbatim}

\noindent [7] A. Levant.``Exact differentiation of signals with unbounded higher derivatives,''  in \textit{Proceedings of the 45th IEEE Conference on Decision and Control}, San Diego, California, USA, pp. 5585--5590, 2006.

\noindent coded as:
\begin{verbatim}\bibitem{Levant}
A. Levant. ``Exact differentiation of 
signals with unbounded higher 
derivatives,''  in \textit{Proceedings 
of the 45th IEEE Conference on 
Decision and Control}, San Diego, 
California, USA, pp. 5585--5590, 2006.
\end{verbatim}

\noindent [8] M. Fliess, C. Join, and H. Sira-Ramirez. ``Non-linear estimation is easy,'' \textit{International Journal of Modelling, Identification and Control}, vol. 4, no. 1, pp. 12--27, 2008.

\noindent coded as:
\begin{verbatim}
\bibitem{Cedric}
M. Fliess, C. Join, and H. Sira-Ramirez. 
``Non-linear estimation is easy,'' 
\textit{International Journal of Modelling, 
Identification and Control}, vol. 4, 
no. 1, pp. 12--27, 2008.
\end{verbatim}

\noindent [9] R. Ortega, A. Astolfi, G. Bastin, and H. Rodriguez. ``Stabilization of food-chain systems using a port-controlled Hamiltonian description,'' in \textit{Proceedings of the American Control Conference}, Chicago, Illinois, USA, pp. 2245--2249, 2000.

\noindent coded as:
\begin{verbatim}
\bibitem{Ortega}
R. Ortega, A. Astolfi, G. Bastin, and H. 
Rodriguez. ``Stabilization of food-chain 
systems using a port-controlled Hamiltonian 
description,'' in \textit{Proceedings of the 
American Control Conference}, Chicago, 
Illinois, USA, pp. 2245--2249, 2000.
\end{verbatim}

\subsection{Accented Characters in References}
\noindent When using accented characters in references, please use the standard LaTeX coding for accents. {\bf{Do not use math coding for character accents}}. For example:
\begin{verbatim}
\'e, \"o, \`a, \~e 
\end{verbatim}
will produce: \'e, \"o, \`a, \~e

\subsection{Use of BibTeX}
\noindent If you wish to use BibTeX, please see the documentation that accompanies the IEEEtran Bibliography package.

\subsection{Biographies and Author Photos}
\noindent Authors may have options to include their photo or not. Photos should be a bit-map graphic (.tif or .jpg) and sized to fit in the space allowed. Please see the coding samples below:
\begin{verbatim}
\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here without a photo.
\end{IEEEbiographynophoto}
\end{verbatim}
or a biography with a photo

\begin{verbatim}
\begin{IEEEbiography}[{\includegraphics
[width=1in,height=1.25in,clip,
keepaspectratio]{fig1.png}}]
{IEEE Publications Technology Team} 
In this paragraph you can place 
your educational, professional background 
and research and other interests.
\end{IEEEbiography}
\end{verbatim}

Please see the end of this document to see the output of these coding examples.

\section{Mathematical Typography \\ and Why It Matters}

\noindent Typographical conventions for mathematical formulas have been developed to {\bf provide uniformity and clarity of presentation across mathematical texts}. This enables the readers of those texts to both understand the author's ideas and to grasp new concepts quickly. While software such as \LaTeX \ and MathType\textsuperscript{\textregistered} can produce aesthetically pleasing math when used properly, it is also very easy to misuse the software, potentially resulting in incorrect math display.

IEEE aims to provide authors with the proper guidance on mathematical typesetting style and assist them in writing the best possible article.

As such, IEEE has assembled a set of examples of good and bad mathematical typesetting. You will see how various issues are dealt with. The following publications have been referenced in preparing this material:

\begin{list}{}{}
\item{\emph{Mathematics into Type}, published by the American Mathematical Society}
\item{\emph{The Printing of Mathematics}, published by Oxford University Press}
\item{\emph{The \LaTeX Companion}, by F. Mittelbach and M. Goossens}
\item{\emph{More Math into LaTeX}, by G. Gr\"atzer}
\item{AMS-StyleGuide-online.pdf, published by the American Mathematical Society}
\end{list}

Further examples can be seen at \url{http://journals.ieeeauthorcenter.ieee.org/wp-content/uploads/sites/7/IEEE-Math-Typesetting-Guide.pdf}

\subsection{Display Equations}
\noindent A simple display equation example shown below uses the ``equation'' environment. To number the equations, use the $\backslash${\tt{label}} macro to create an identifier for the equation. LaTeX will automatically number the equation for you.
\begin{equation}
\label{deqn_ex1}
x = \sum_{i=0}^{n} 2{i} Q.
\end{equation}

\noindent is coded as follows:
\begin{verbatim}
\begin{equation}
\label{deqn_ex1}
x = \sum_{i=0}^{n} 2{i} Q.
\end{equation}
\end{verbatim}

To reference this equation in the text use the $\backslash${\tt{ref}} macro. 
Please see (\ref{deqn_ex1})\\
\noindent is coded as follows:
\begin{verbatim}
Please see (\ref{deqn_ex1})\end{verbatim}

\subsection{Equation Numbering}
\noindent {\bf{Consecutive Numbering:}} Equations within an article are numbered consecutively from the beginning of the
article to the end, i.e., (1), (2), (3), (4), (5), etc. Do not use roman numerals or section numbers for equation numbering.\\

\noindent {\bf{Appendix Equations:}} The continuation of consecutively numbered equations is best in the Appendix, but numbering
 as (A1), (A2), etc., is permissible.\\

\noindent {\bf{Hyphens and Periods}}: Hyphens and periods should not be used in equation numbers, i.e., use (1a) rather than
(1-a) and (2a) rather than (2.a) for sub-equations. This should be consistent throughout the article.

\subsection{Multi-line equations and alignment}
\noindent Here we show several examples of multi-line equations and proper alignments.

\noindent {\bf{A single equation that must break over multiple lines due to length with no specific alignment.}}
\begin{multline}
\text{The first line of this example}\\
\text{The second line of this example}\\
\text{The third line of this example}
\end{multline}

\noindent is coded as:
\begin{verbatim}
\begin{multline}
\text{The first line of this example}\\
\text{The second line of this example}\\
\text{The third line of this example}
\end{multline}
\end{verbatim}

\noindent {\bf{A single equation with multiple lines aligned at the = signs}}
\begin{align}
a &= c+d \\
b &= e+f
\end{align}
\noindent is coded as:
\begin{verbatim}
\begin{align}
a &= c+d \\
b &= e+f
\end{align}
\end{verbatim}

The {\tt{align}} environment can align on multiple  points as shown in the following example:
\begin{align}
x &= y & X & =Y & a &=bc\\
x' &= y' & X' &=Y' &a' &=bz
\end{align}
\noindent is coded as:
\begin{verbatim}
\begin{align}
x &= y & X & =Y & a &=bc\\
x' &= y' & X' &=Y' &a' &=bz
\end{align}
\end{verbatim}

\subsection{Subnumbering}
\noindent The amsmath package provides a {\tt{subequations}} environment to facilitate subnumbering. An example:

\begin{subequations}\label{eq:2}
\begin{align}
f&=g \label{eq:2A}\\
f' &=g' \label{eq:2B}\\
\mathcal{L}f &= \mathcal{L}g \label{eq:2c}
\end{align}
\end{subequations}

\noindent is coded as:
\begin{verbatim}
\begin{subequations}\label{eq:2}
\begin{align}
f&=g \label{eq:2A}\\
f' &=g' \label{eq:2B}\\
\mathcal{L}f &= \mathcal{L}g \label{eq:2c}
\end{align}
\end{subequations}

\end{verbatim}

\subsection{Matrices}
\noindent There are several useful matrix environments that can save you some keystrokes. See the example coding below and the output.

\noindent {\bf{A simple matrix:}}
\begin{equation}
\begin{matrix}  0 &  1 \\ 
1 &  0 \end{matrix}
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\begin{matrix}  0 &  1 \\ 
1 &  0 \end{matrix}
\end{equation}
\end{verbatim}

\noindent {\bf{A matrix with parenthesis}}
\begin{equation}
\begin{pmatrix} 0 & -i \\
 i &  0 \end{pmatrix}
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\begin{pmatrix} 0 & -i \\
 i &  0 \end{pmatrix}
\end{equation}
\end{verbatim}

\noindent {\bf{A matrix with square brackets}}
\begin{equation}
\begin{bmatrix} 0 & -1 \\ 
1 &  0 \end{bmatrix}
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\begin{bmatrix} 0 & -1 \\ 
1 &  0 \end{bmatrix}
\end{equation}
\end{verbatim}

\noindent {\bf{A matrix with curly braces}}
\begin{equation}
\begin{Bmatrix} 1 &  0 \\ 
0 & -1 \end{Bmatrix}
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\begin{Bmatrix} 1 &  0 \\ 
0 & -1 \end{Bmatrix}
\end{equation}\end{verbatim}

\noindent {\bf{A matrix with single verticals}}
\begin{equation}
\begin{vmatrix} a &  b \\ 
c &  d \end{vmatrix}
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\begin{vmatrix} a &  b \\ 
c &  d \end{vmatrix}
\end{equation}\end{verbatim}

\noindent {\bf{A matrix with double verticals}}
\begin{equation}
\begin{Vmatrix} i &  0 \\ 
0 & -i \end{Vmatrix}
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\begin{Vmatrix} i &  0 \\ 
0 & -i \end{Vmatrix}
\end{equation}\end{verbatim}

\subsection{Arrays}
\noindent The {\tt{array}} environment allows you some options for matrix-like equations. You will have to manually key the fences, but you'll have options for alignment of the columns and for setting horizontal and vertical rules. The argument to {\tt{array}} controls alignment and placement of vertical rules.

A simple array
\begin{equation}
\left(
\begin{array}{cccc}
a+b+c & uv & x-y & 27\\
a+b & u+v & z & 134
\end{array}\right)
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\left(
\begin{array}{cccc}
a+b+c & uv & x-y & 27\\
a+b & u+v & z & 134
\end{array} \right)
\end{equation}
\end{verbatim}

A slight variation on this to better align the numbers in the last column
\begin{equation}
\left(
\begin{array}{cccr}
a+b+c & uv & x-y & 27\\
a+b & u+v & z & 134
\end{array}\right)
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\left(
\begin{array}{cccr}
a+b+c & uv & x-y & 27\\
a+b & u+v & z & 134
\end{array} \right)
\end{equation}
\end{verbatim}

An array with vertical and horizontal rules
\begin{equation}
\left( \begin{array}{c|c|c|r}
a+b+c & uv & x-y & 27\\ \hline
a+b & u+v & z & 134
\end{array}\right)
\end{equation}
is coded as:
\begin{verbatim}
\begin{equation}
\left(
\begin{array}{c|c|c|r}
a+b+c & uv & x-y & 27\\
a+b & u+v & z & 134
\end{array} \right)
\end{equation}
\end{verbatim}
Note the argument now has the pipe "$\vert$" included to indicate the placement of the vertical rules.

\subsection{Cases Structures}
\noindent Many times we find cases coded using the wrong environment, i.e., {\tt{array}}. Using the {\tt{cases}} environment will save keystrokes (from not having to type the $\backslash${\tt{left}}$\backslash${\tt{lbrace}}) and automatically provide the correct column alignment.
\begin{equation*}
{z_m(t)} = \begin{cases}
1,&{\text{if}}\ {\beta }_m(t) \\ 
{0,}&{\text{otherwise.}} 
\end{cases}
\end{equation*}
\noindent is coded as follows:
\begin{verbatim}
\begin{equation*}
{z_m(t)} = 
\begin{cases}
1,&{\text{if}}\ {\beta }_m(t),\\ 
{0,}&{\text{otherwise.}} 
\end{cases}
\end{equation*}
\end{verbatim}
\noindent Note that the ``\&'' is used to mark the tabular alignment. This is important to get  proper column alignment. Do not use $\backslash${\tt{quad}} or other fixed spaces to try and align the columns. Also, note the use of the $\backslash${\tt{text}} macro for text elements such as ``if'' and ``otherwise''.

\subsection{Function Formatting in Equations}
In many cases there is an easy way to properly format most common functions. Use of the $\backslash$ in front of the function name will in most cases, provide the correct formatting. When this does not work, the following example provides a solution using the $\backslash${\tt{text}} macro.

\begin{equation*} 
  d_{R}^{KM} = \underset {d_{l}^{KM}} {\text{arg min}} \{ d_{1}^{KM},\ldots,d_{6}^{KM}\}.
\end{equation*}

\noindent is coded as follows:
\begin{verbatim}
\begin{equation*} 
 d_{R}^{KM} = \underset {d_{l}^{KM}} 
 {\text{arg min}} \{ d_{1}^{KM},
 \ldots,d_{6}^{KM}\}.
\end{equation*}
\end{verbatim}

\subsection{ Text Acronyms inside equations}
\noindent This example shows where the acronym ``MSE" is coded using $\backslash${\tt{text\{\}}} to match how it appears in the text.

\begin{equation*}
 \text{MSE} = \frac {1}{n}\sum _{i=1}^{n}(Y_{i} - \hat {Y_{i}})^{2}
\end{equation*}

\begin{verbatim}
\begin{equation*}
 \text{MSE} = \frac {1}{n}\sum _{i=1}^{n}
(Y_{i} - \hat {Y_{i}})^{2}
\end{equation*}
\end{verbatim}

\subsection{Obsolete Coding}
\noindent Avoid the use of outdated environments, such as {\tt{eqnarray}} and \$\$ math delimiters, for display equations. The \$\$ display math delimiters are left over from PlainTeX and should not be used in \LaTeX, ever. Poor vertical spacing will result.
\subsection{Use Appropriate Delimiters for Display Equations}
\noindent Some improper mathematical coding advice has been given in various YouTube\textsuperscript{TM} videos on how to write scholarly articles, so please follow these good examples:\\

For {\bf{single-line unnumbered display equations}}, please use the following delimiters: 
\begin{verbatim}\[ . . . \] or \end{verbatim} 
\begin{verbatim}\begin{equation*} . . . \end{equation*}\end{verbatim}
Note that the * in the environment name turns off equation numbering.\\

For {\bf{multiline unnumbered display equations}} that have alignment requirements, please use the following delimiters: 
\begin{verbatim}
\begin{align*} . . . \end{align*}
\end{verbatim}

For {\bf{single-line numbered display equations}}, please use the following delimiters: 
\begin{verbatim}
\begin{equation} . . . \end{equation}
\end{verbatim}

For {\bf{multiline numbered display equations}}, please use the following delimiters: 
\begin{verbatim}
\begin{align} . . . \end{align}
\end{verbatim}

\section{LaTeX Package Suggestions}
\noindent Immediately after your documenttype declaration at the top of your \LaTeX\ file is the place where you should declare any packages that are being used. The following packages were used in the production of this document.
\begin{verbatim}
\usepackage{amsmath,amsfonts}
\usepackage{algorithmic}
\usepackage{array}
\usepackage[caption=false,font=normalsize,
   labelfont=sf,textfont=sf]{subfig}
\u00sepackage{textcomp}
\usepackage{stfloats}
\usepackage{url}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{balance}
\end{verbatim}

\section{Additional Advice}

Please use ``soft'' (e.g., \verb|\eqref{Eq}|) or \verb|(\ref{Eq})|
cross references instead of ``hard'' references (e.g., \verb|(1)|).
That will make it possible to combine sections, add equations, or
change the order of figures or citations without having to go through
the file line by line.

Please note that the \verb|{subequations}| environment in {\LaTeX}
will increment the main equation counter even when there are no
equation numbers displayed. If you forget that, you might write an
article in which the equation numbers skip from (17) to (20), causing
the copy editors to wonder if you've discovered a new method of
counting.

{\BibTeX} does not work by magic. It doesn't get the bibliographic
data from thin air but from .bib files. If you use {\BibTeX} to produce a
bibliography you must send the .bib files.

{\LaTeX} can't read your mind. If you assign the same label to a
subsubsection and a table, you might find that Table I has been cross
referenced as Table IV-B3.

{\LaTeX} does not have precognitive abilities. If you put a
\verb|\label| command before the command that updates the counter it's
supposed to be using, the label will pick up the last counter to be
cross referenced instead. In particular, a \verb|\label| command
should not go before the caption of a figure or a table.

Please do not use \verb|\nonumber| or \verb|\notag| inside the
\verb|{array}| environment. It will not stop equation numbers inside
\verb|{array}| (there won't be any anyway) and it might stop a wanted
equation number in the surrounding equation.

\balance

\section{A Final Checklist}
\begin{enumerate}{}{}
\item{Make sure that your equations are numbered sequentially and there are no equation numbers missing or duplicated. Avoid hyphens and periods in your equation numbering. Stay with IEEE style, i.e., (1), (2), (3) or for sub-equations (1a), (1b). For equations in the appendix (A1), (A2), etc.}. 
\item{Are your equations properly formatted? Text, functions, alignment points in cases and arrays, etc. }
\item{Make sure all graphics are included.}
\item{Make sure your references are included either in your main LaTeX file or a separate .bib file if calling the external file.}
\end{enumerate}

\begin{thebibliography}{1}

\bibitem{ams}
{\it{Mathematics into Type}}, American Mathematical Society. Online available:

\bibitem{oxford}
T.W. Chaundy, P.R. Barrett and C. Batey, {\it{The Printing of Mathematics}}, Oxford University Press. London, 1954.

\bibitem{lacomp}{\it{The \LaTeX Companion}}, by F. Mittelbach and M. Goossens

\bibitem{mmt}{\it{More Math into LaTeX}}, by G. Gr\"atzer

\bibitem{amstyle}{\it{AMS-StyleGuide-online.pdf,}} published by the American Mathematical Society

\bibitem{Sira3}
H. Sira-Ramirez. ``On the sliding mode control of nonlinear systems,'' \textit{Systems \& Control Letters}, vol. 19, pp. 303--312, 1992.

\bibitem{Levant}
A. Levant. ``Exact differentiation of signals with unbounded higher derivatives,''  in \textit{Proceedings of the 45th IEEE Conference on Decision and Control}, San Diego, California, USA, pp. 5585--5590, 2006.

\bibitem{Cedric}
M. Fliess, C. Join, and H. Sira-Ramirez. ``Non-linear estimation is easy,'' \textit{International Journal of Modelling, Identification and Control}, vol. 4, no. 1, pp. 12--27, 2008.

\bibitem{Ortega}
R. Ortega, A. Astolfi, G. Bastin, and H. Rodriguez. ``Stabilization of food-chain systems using a port-controlled Hamiltonian description,'' in \textit{Proceedings of the American Control Conference}, Chicago, Illinois, USA, pp. 2245--2249, 2000.

\end{thebibliography}

\begin{IEEEbiographynophoto}{Jane Doe}
Biography text here without a photo.
\end{IEEEbiographynophoto}

\begin{IEEEbiography}[{\includegraphics[width=1in,height=1.25in,clip,keepaspectratio]{fig1.png}}]{IEEE Publications Technology Team}
In this paragraph you can place your educational, professional background and research and other interests.\end{IEEEbiography}

\end{document}

### 2024-11-18 AEID2025/samplepaper.tex
update

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.21 of 2022/01/12
%
\documentclass[runningheads]{llncs}
%
\usepackage[T1]{fontenc}
% T1 fonts will be used to generate the final print and online PDFs,
% so please use T1 fonts in your manuscript whenever possible.
% Other font encondings may result in incorrect characters.
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following two lines
% to display URLs in blue roman font according to Springer's eBook style:
%\usepackage{color}
%\renewcommand\UrlFont{\color{blue}\rmfamily}
%\urlstyle{rm}
%
\begin{document}
%
\title{Contribution Title}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Rodrigo Prestes Machado\inst{1}\orcidID{0000-0003-0428-6387} \and
Second Author\inst{2,3}\orcidID{1111-2222-3333-4444} \and
Third Author\inst{3}\orcidID{2222--3333-4444-5555}}
%
\authorrunning{F. Author et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{Princeton University, Princeton NJ 08544, USA \and
Springer Heidelberg, Tiergartenstr. 17, 69121 Heidelberg, Germany
\email{lncs@springer.com}\\
\url{http://www.springer.com/gp/computer-science/lncs} \and
ABC Institute, Rupert-Karls-University Heidelberg, Heidelberg, Germany\\
\email{\{abc,lncs\}@uni-heidelberg.de}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}
The aim of this study is to analyze the interactions of programming students
with a chatbot powered by OpenAI's GPT-3.5, enhanced with the Retrieval
Augmented Generation (RAG) technique, within the context of a Java programming
course. The study investigates how students use the chatbot while employing
two metacognitive strategies: interleaving and spacing. The methodology involved
classifying and categorizing the students' interaction prompts with the chatbot
into eight distinct categories. The findings reveal that, while most
interactions focus on active learning strategies, students demonstrate limited
conscious application of metacognitive strategies. The study highlights the need
for more comprehensive guidance on the strategic use of AI tools to enhance
learning outcomes.

\keywords{Programming \and Generative Artificial Intelligence \and Chatbots \and
Spacing \and Interleaving \and Metacognition.}
\end{abstract}
%
%
%
\section{Theoretical Framework}
\subsection{Metacognition}
Given that GenAI often produces highly accurate and automatic responses
\cite{Puryear22}, it is essential that students use these tools within an
active learning process to enhance their learning outcomes. This active
learning process can be further understood through the lens of metacognition,
which focuses on awareness of one's mental processes. \cite{flavell79}
proposed a model of metacognitive monitoring that includes four interrelated
phenomena: Metacognitive Knowledge, Metacognitive Experience, Metacognitive
Goals, and Metacognitive Actions. These processes do not occur in isolation, but
they influence each other, altering cognitive progress over time.

Metacognitive knowledge includes beliefs about variables that affect the
outcomes of cognitive activities. It is divided into three types: beliefs about
personal abilities, perceived difficulty in the task, and previously used
strategies. Metacognitive Experience refers to the feelings that arise before,
during, and after cognitive activity, such as frustration, confusion,
satisfaction, and others. Metacognitive Goals are the key to regulating thought
as they relate to the goals the individual seeks to achieve, directly
influencing the actions taken. For example, if a student’s goal is to complete
a task quickly, they may adopt a more passive approach to learning. Lastly,
Metacognitive Actions involve the planning, monitoring, and evaluation of
strategies used to achieve the goals. In terms of planning, students can
determine how to approach a task, such as spacing their study sessions
\cite{Ouhao18,Carvalho20} interleaving topics \cite{Rivers21}, utilizing
retrieval practice \cite{larsen18}, and other possible strategies.

Spacing and interleaving are two metacognitive strategies that have been shown
to enhance learning outcomes and support the research questions of this study.
Spacing refers to the practice of distributing study sessions over time, which
has been shown to improve long-term retention and understanding of the material
\cite{Carvalho20}. Interleaving involves mixing different topics or problems
within a single learning session, which has been shown to also enhance long-term
learning and the application of student's knowledge in other contexts and
situations \cite{Rivers21}.

\subsubsection{Sample Heading (Third Level)} Only two levels of
headings should be numbered. Lower level headings remain unnumbered;
they are formatted as run-in headings.

\paragraph{Sample Heading (Fourth Level)}
The contribution should contain no more than four levels of
headings. Table~\ref{tab1} gives a summary of all heading levels.

\begin{table}
\caption{Table captions should be placed above the
tables.}\label{tab1}
\begin{tabular}{|l|l|l|}
\hline
Heading level &  Example & Font size and style\\
\hline
Title (centered) &  {\Large\bfseries Lecture Notes} & 14 point, bold\\
1st-level heading &  {\large\bfseries 1 Introduction} & 12 point, bold\\
2nd-level heading & {\bfseries 2.1 Printing Area} & 10 point, bold\\
3rd-level heading & {\bfseries Run-in Heading in Bold.} Text follows & 10 point, bold\\
4th-level heading & {\itshape Lowest Level Heading.} Text follows & 10 point, italic\\
\hline
\end{tabular}
\end{table}

\noindent Displayed equations are centered and set on a separate
line.
\begin{equation}
x + y = z
\end{equation}
Please try to avoid rasterized images for line-art diagrams and
schemas. Whenever possible, use vector graphics instead (see
Fig.~\ref{fig1}).

\begin{figure}
\includegraphics[width=\textwidth]{fig1.eps}
\caption{A figure caption is always placed below the illustration.
Please note that short captions are centered, while long ones are
justified by the macro package automatically.} \label{fig1}
\end{figure}

\begin{theorem}
This is a sample theorem. The run-in heading is set in bold, while
the following text appears in italics. Definitions, lemmas,
propositions, and corollaries are styled the same way.
\end{theorem}
%
% the environments 'definition', 'lemma', 'proposition', 'corollary',
% 'remark', and 'example' are defined in the LLNCS documentclass as well.
%
\begin{proof}
Proofs, examples, and remarks have the initial word in italics,
while the following text appears in normal font.
\end{proof}
For citations of references, we prefer the use of square brackets
and consecutive numbers. Citations using labels or the author/year
convention are also acceptable. The following bibliography provides
a sample reference list with entries for journal
articles~\cite{ref_article1}, an LNCS chapter~\cite{ref_lncs1}, a
book~\cite{ref_book1}, proceedings without editors~\cite{ref_proc1},
and a homepage~\cite{ref_url1}. Multiple citations are grouped
\cite{ref_article1,ref_lncs1,ref_book1},
\cite{ref_article1,ref_book1,ref_proc1,ref_url1}.

\begin{credits}
\subsubsection{\ackname} A bold run-in heading in small font size at the end of the paper is
used for general acknowledgments, for example: This study was funded
by X (grant number Y).

\subsubsection{\discintname}
It is now necessary to declare any competing interests or to specifically
state that the authors have no competing interests. Please place the
statement with a bold run-in heading in small font size beneath the
(optional) acknowledgments\footnote{If EquinOCS, our proceedings submission
system, is used, then the disclaimer can be provided directly in the system.},
for example: The authors have no competing interests to declare that are
relevant to the content of this article. Or: Author A has received research
grants from Company W. Author B has received a speaker honorarium from
Company X and owns stock in Company Y. Author C is a member of committee Z.
\end{credits}
%
% ---- Bibliography ----
%
% BibTeX users should specify bibliography style 'splncs04'.
% References will then be sorted and formatted in the correct style.
%
\bibliographystyle{splncs04}
\bibliography{References}
\end{document}

### 2024-11-18 CSEDU2025/Paper.tex
update

@@
 examining these interactions within the context of two metacognitive strategies:
-spacing and interleaving. To achieve this objective, we formulate three research
-questions:
+spacing \citep{Carvalho20} and interleaving \citep{Rivers21}. To achieve this
+objective, we formulate three research questions:
 
@@
 
-This section outlines the theoretical framework that grounds the study. We
-begin by introducing the concept of metacognition, along with the spacing and
-interweaving strategies applied in this research. Next, we examine related works
- organized into two subsections: the use of GenAI in programming and recent
- studies emphasizing the importance of metacognition in the use of GenAI.
+This section outlines the theoretical framework that is the basis of the study.
+We begin by introducing the concept of metacognition, along with the spacing and
+interweaving strategies applied in this research. Next, we present related works
+organized into two subsections: research involving GenAI and studies focusing
+on metacognition.
 
@@
 
-\subsection{GenAI and Programming}
+\subsection{GenAI Related Works}
 
@@
 use of AI.
-
-\cite{Bastani24} investigated the impact of GenAI, specifically GPT-4, on human
-learning, with a focus on mathematics education in a high school. The problem
-addressed was how the use of generative AI could affect the acquisition of new
-skills, which is crucial for long-term productivity. The method involved a
-controlled experiment with approximately one thousand students, who were exposed
-to two GPT-4-based tutors: a simple tutor (GPT Base) and another with safeguards
-designed to promote learning (GPT Tutor). The results showed that while access
-to GPT-4 improved performance on practice exercises (48\% with GPT Base and
-127\% with GPT Tutor), the removal of access to GPT Base led to a 17\% decrease
-in student performance on exams. This suggests that unrestricted use of
-\textit{GPT Base} could hinder learning. However, the GPT Tutor was able to
-mitigate this negative effect.
 
@@
 using the tool in the future.
+
+\cite{Bastani24} investigated the impact of GenAI, specifically GPT-4, on human
+learning, with a focus on mathematics education in a high school. The problem
+addressed was how the use of generative AI could affect the acquisition of new
+skills, which is crucial for long-term productivity. The method involved a
+controlled experiment with approximately one thousand students, who were exposed
+to two GPT-4-based tutors: a simple tutor (GPT Base) and another with safeguards
+designed to promote learning (GPT Tutor). The results showed that while access
+to GPT-4 improved performance on practice exercises (48\% with GPT Base and
+127\% with GPT Tutor), the removal of access to GPT Base led to a 17\% decrease
+in student performance on exams. This suggests that unrestricted use of
+\textit{GPT Base} could hinder learning. However, the GPT Tutor was able to
+mitigate this negative effect.
 
@@
 
-\subsection{GenAI and Metacognition}
-
-\cite{Sidra24} addressed the growing importance of metacognitive skills in an
-AI-enabled workforce. The problem discussed was how AI systems, especially
-collaborative tools like Large Language Models (LLMs), impacted human
-decision-making and the challenges they posed due to the lack of human-like
-metacognition in AI. The method used identified four main characteristics that
-differentiated human-AI interaction from human-human collaboration and proposed,
-based on the dual-process theory of human thought, that metacognitive thinking
-was crucial for the effective use of AI. The results indicated that human
-workers needed to develop and apply metacognitive skills, such as planning,
-monitoring, and evaluating their thought processes, to mitigate cognitive biases
-and improve decision-making when working with AI. The study concluded by
-recommending improvements in AI design and training to improve workers'
-metacognitive abilities, ensuring more effective collaboration with AI systems.
+\subsection{Metacognition Related Works}
 
@@
 elucidates the relationship between meta-learning and metacognition. This
-research suggests that metacognition can improve metacognitive abilities,
+research suggests that meta-learning can improve metacognitive abilities,
 providing valuable insights into educational technology and course design in
 higher education settings.
+
+\cite{Sidra24} addressed the growing importance of metacognitive skills in an
+AI-enabled workforce. The problem discussed was how AI systems, especially in
+collaborative tools based on Large Language Models (LLMs), impacted human
+decision-making and the challenges they posed due to the lack of human-like
+metacognition in AI. The method used identified four main characteristics that
+differentiated human-AI interaction from human-human collaboration and proposed,
+based on the dual-process theory of human thought, that metacognitive thinking
+was crucial for the effective use of AI. The results indicated that human
+workers needed to develop and apply metacognitive skills, such as planning,
+monitoring, and evaluating their thought processes, to mitigate cognitive biases
+and improve decision-making when working with AI. The study concluded by
+recommending improvements in AI design and training to improve workers'
+metacognitive abilities, ensuring more effective collaboration with AI systems.
 

### 2024-11-19 AEID2025/PaperAEID25.tex
PaperAEID25

--- 2024-11-18 CSEDU2025/Paper.tex
@@
-\documentclass[a4paper,twoside]{article}
-
-\usepackage{epsfig}
-\usepackage{subcaption}
-\usepackage{calc}
-\usepackage{amssymb}
-\usepackage{amstext}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{multicol}
-\usepackage{pslatex}
-\usepackage{algorithm2e}
-\usepackage[bottom]{footmisc}
-% Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
-\usepackage{natbib}
-\usepackage{multirow}
-\usepackage{SCITEPRESS}
-\usepackage{arydshln}
-\widowpenalty=10000
-\clubpenalty=10000
+% This is samplepaper.tex, a sample chapter demonstrating the
+% LLNCS macro package for Springer Computer Science proceedings;
+% Version 2.21 of 2022/01/12
+%
+\documentclass[runningheads]{llncs}
+%
+\usepackage[T1]{fontenc}
+% T1 fonts will be used to generate the final print and online PDFs,
+% so please use T1 fonts in your manuscript whenever possible.
+% Other font encondings may result in incorrect characters.
+%
+\usepackage{graphicx}
+% Used for displaying a sample figure. If possible, figure files should
+% be included in EPS format.
+%
+% If you use the hyperref package, please uncomment the following two lines
+% to display URLs in blue roman font according to Springer's eBook style:
+%\usepackage{color}
+%\renewcommand\UrlFont{\color{blue}\rmfamily}
+%\urlstyle{rm}
+%
+\usepackage{arydshln} % Add this line to use the \hdashline command
 
 \begin{document}
-
-\title{
-  Analyzing Student Use of Spacing and Interleaving Strategies in Interactions
-  with GenAI-Powered Chatbots in Programming Courses
+%
+\title{Analyzing Student Interaction Patterns with GenAI-Powered Chatbots in
+Programming Courses through Metacognitive Strategies of Spacing and Interleaving
 }
-
-\author{\authorname{Rodrigo Prestes Machado\sup{1}\orcidAuthor{0000-0003-0428-6387},
-Carlos Alario-Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814},
-Patricia Callejo\sup{2}\orcidAuthor{0000-0001-6124-6213},
-Iria Estévez-Ayres\sup{2}\orcidAuthor{0000-0002-1047-5398},
-Carlos Delgado Kloos\sup{2}\orcidAuthor{0000-0003-4093-3705}}
-\affiliation{\sup{1}Department of Informatics, Instituto Federal de Educação,
-Ciência e Tecnologia, Porto Alegre, Brazil}
-\affiliation{\sup{2}Department of Telematics Engineering, Universidad Carlos III
-de Madrid, Leganés (Madrid), Spain}
-\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario, pcallejo, ayres, cdk\}@it.uc3m.es}
+%
+\titlerunning{Analyzing Student Interaction with Chatbots through Metacognition}
+% If the paper title is too long for the running head, you can set
+% an abbreviated paper title here
+%
+\author{
+Rodrigo Prestes Machado\inst{1}\orcidID{0000-0003-0428-6387} \and
+Carlos Alario-Hoyos\inst{2}\orcidID{0000-0002-3082-0814} \and
+Patricia Callejo\inst{2}\orcidID{0000-0001-6124-6213} \and
+Iria Estévez-Ayres\inst{2}\orcidID{0000-0002-1047-5398} \and
+Carlos Delgado Kloos\inst{2}\orcidID{0000-0003-4093-3705}}
+%
+\authorrunning{Machado, R. P., et al.}
+% First names are abbreviated in the running head.
+% If there are more than two authors, 'et al.' is used.
+%
+\institute{Instituto Federal de Educação, Ciência e Tecnologia, Porto Alegre, Brazil \\
+\url{http://poa.ifrs.edu.br} \\
+\email{rodrigo.prestes@poa.ifrs.edu.br} \and
+Universidad Carlos III de Madrid, Leganés, Spain \\
+\url{https://www.uc3m.es} \\
+\email{\{calario, pcallejo, ayres, cdk\}@it.uc3m.es}
 }
-
-\keywords{
-  Programming, Generative Artificial Intelligence, Chatbots, Spacing,
-  Interleaving and Metacognition
-}
-
-\abstract{
+%
+\maketitle              % typeset the header of the contribution
+%
+\begin{abstract}
 The aim of this study is to analyze the interactions of programming students
@@
 learning outcomes.
-}
 
-\twocolumn\maketitle\normalsize\setcounter{footnote}{0}
-
-\section{\uppercase{Introduction}}
-\label{sec:introduction}
+\keywords{Programming \and Generative Artificial Intelligence \and Chatbots \and
+Spacing \and Interleaving \and Metacognition.}
+\end{abstract}
+%
+%
+\section{Introduction}
 
 Recent advances in Generative Artificial Intelligence (GenAI) have created
-new opportunities in both professional programming \citep{Peng23}
-\citep{Pandey24} and education \citep{Puryear22}. In programming courses,
+new opportunities in both professional programming \cite{Peng23}
+\cite{Pandey24} and education \cite{Puryear22}. In programming courses,
 students can employ GenAI tools to improve their understanding, receive
@@
 providing real-time suggestions and accelerating code writing, which could help
-streamline the learning process \citep{Denny23}. Educational chatbots, on the
+streamline the learning process \cite{Denny23}. Educational chatbots, on the
 other hand, try to guide students through coding challenges, answer questions,
 and try to foster a deeper understanding of programming concepts
-\citep{Labadze23}.
+\cite{Labadze23}.
 
@@
 effectively learn from ChatGPT, resulting in improved comprehension and academic
-achievement \citep{Callejo24}. Furthermore, it was observed that ChatGPT may
-allow students to regulate their learning pace \citep{Baha24} and can
+achievement \cite{Callejo24}. Furthermore, it was observed that ChatGPT may
+allow students to regulate their learning pace \cite{Baha24} and can
 potentially support self-regulated learning, particularly for students with
-prior technical and disciplinary knowledge \citep{Xia23}.
+prior technical and disciplinary knowledge \cite{Xia23}.
 
@@
 GenAI for homework, likely due to their perception of this technology as
-advantageous and intuitive \citep{Boudouaia24}. All of these studies highlight
+advantageous and intuitive \cite{Boudouaia24}. All of these studies highlight
 the need to increase the understanding of how students interact with these tools
@@
 examining these interactions within the context of two metacognitive strategies:
-spacing \citep{Carvalho20} and interleaving \citep{Rivers21}. To achieve this
+spacing \cite{Carvalho20} and interleaving \cite{Rivers21}.
+
+To achieve this
 objective, we formulate three research questions:
@@
 
-\section{\uppercase{Theoretical Framework}}
+\section{Theoretical Framework}
 
@@
 Given that GenAI often produces highly accurate and automatic responses
-\citep{Puryear22}, it is essential that students use these tools within an
+\cite{Puryear22}, it is essential that students use these tools within an
 active learning process to enhance their learning outcomes. This active
@@
 determine how to approach a task, such as spacing their study sessions
-\citep{Ouhao18, Carvalho20}, interleaving topics \citep{Rivers21}, utilizing
-retrieval practice \citep{larsen18}, and other possible strategies.
+\cite{Ouhao18,Carvalho20} interleaving topics \cite{Rivers21}, utilizing
+retrieval practice \cite{larsen18}, and other possible strategies.
 
@@
 has been shown to improve long-term retention and understanding of the material
-\citep{Carvalho20}. Interleaving involves mixing different topics or problems
+\cite{Carvalho20}. Interleaving involves mixing different topics or problems
 within a single learning session, which has been shown to also enhance long-term
 learning and the application of student's knowledge in other contexts and
-situations \citep{Rivers21}.
+situations \cite{Rivers21}.
 
@@
 
-\begin{table*}[htbp]
-  \caption{Categories - adapted from \cite{Ghimire24}}
-  \begin{center}
-    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{4cm} p{12cm}} % Remove vertical bars
-      \hline
-      \textbf{Category} & \textbf{Description} \\
-      \hline
-      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
-      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. \\
-      Student Correction (SC) & Prompts where the student corrects the chatbot. \\
-      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. \\
-      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. \\
-      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
-      Language Change (LC) & Prompts that request a change of idiom. \\
-      Uncategorized (U) & Prompts that do not fit into any of the above categories. \\
-      \hline
-    \end{tabular}
-    \label{tab:categories}
-  \end{center}
-\end{table*}
-
 \subsection{Metacognition Related Works}
@@
 
-\section{\uppercase{Material and Method}}
+\section{Material and Method}
 
@@
 Retrieval-Augmented Generation (RAG) technique to improve the precision of
-responses provided to students \citep{Sun24}. The RAG approach was developed
+responses provided to students \cite{Sun24}. The RAG approach was developed
 using materials from a MOOC, including text, exercises, video transcripts, and
@@
 
-\section{\uppercase{Results and Discussion}}
+\begin{table}[htbp]
+\caption{Categories - adapted from \cite{Ghimire24}}
+\centering
+\renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
+\begin{tabular}{|p{4cm}|p{8cm}|} % Add vertical bars
+\hline
+\textbf{Category} & \textbf{Description} \\
+\hline
+Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
+Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. \\
+Student Correction (SC) & Prompts where the student corrects the chatbot. \\
+Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. \\
+Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. \\
+Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
+Language Change (LC) & Prompts that request a change of idiom. \\
+Uncategorized (U) & Prompts that do not fit into any of the above categories. \\
+\hline
+\end{tabular}
+\label{tab:categories}
+\end{table}
+
+\section{Results and Discussion}
 
@@
   \centering
-  \includegraphics[scale=0.62]{img/figure1.png}
+  \includegraphics[scale=0.8]{img/figure1.png}
   \caption{Categorization of messages.}
@@
 \end{figure}
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.45]{img/figure2.png}
-  \caption{Examples of conversations.}
-  \label{fig:graph2}
-\end{figure*}
 
@@
 For the analysis of the spacing, we initially considered a minimum interval of
-60 minutes as a reference to define the significant spacing \citep{Gadella24}.
+60 minutes as a reference to define the significant spacing \cite{Gadella24}.
 Based on this, our results showed that, of the 81 conversations analyzed, only
@@
 
+\begin{figure*}[htbp]
+    \centering
+    \includegraphics[scale=0.35]{img/figure2.png}
+    \caption{Examples of conversations.}
+    \label{fig:graph2}
+\end{figure*}
+
 This data suggests that there is indeed little spacing between study sessions.
@@
 
+\begin{figure*}[htbp]
+  \centering
+  \includegraphics[scale=0.3]{img/figure3.png}
+  \caption{Subsections per conversation. \textit{A zero value indicates no
+  subsection and no spacing.}}
+  \label{fig:graph3}
+\end{figure*}
+
 When a student requests a code snippet (CS), it is generally predictable that
@@
 of code.
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.39]{img/figure3.png}
-  \caption{Subsections per conversation. \textit{A zero value indicates
-  no subsection and no spacing.}}
-  \label{fig:graph3}
-\end{figure*}
 
@@
 more active behavior.
+
+\begin{table}[htbp]
+\caption{Average and Standard Deviation of Different Categories (in minutes)}
+\centering
+\renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
+    \begin{tabular}{|p{4cm}|p{4cm}|p{4cm}|} % Add vertical bars
+        \hline
+        \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
+        \hline
+        Code Snippet & 1.26 & 1.74 \\
+        Uncategorized & 1.65 & 3.54 \\
+        Language Change & 1.89 & 3.28 \\
+        Multiple Question & 2.65 & 6.10 \\
+        Student Correction & 3.17 & 7.33 \\
+        \hdashline
+        \textit{Overall} & \textit{3.43} & \textit{6.60} \\
+        \hdashline
+        Debugging Help & 4.84 & 10.13 \\
+        Conceptual Questions & 5.41 & 10.54 \\
+        Complete Solution & 6.55 & 10.14 \\
+        \hline
+    \end{tabular}
+\label{tab:averages}
+\end{table}
 
@@
 approaches and depths in the learning process.
-
-\begin{table*}[htbp]
-  \caption{Average and Standard Deviation of Different Categories (in minutes)}
-  \begin{center}
-    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{4cm} p{4cm} p{4cm}} % Remove vertical bars
-      \hline
-      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
-      \hline
-      Code Snippet & 1.26 & 1.74 \\
-      Uncategorized & 1.65 & 3.54 \\
-      Language Change & 1.89 & 3.28 \\
-      Multiple Question & 2.65 & 6.10 \\
-      Student Correction & 3.17 & 7.33 \\
-      \hdashline
-      \textit{Overall} & \textit{3.43} & \textit{6.60} \\
-      \hdashline
-      Debugging Help & 4.84 & 10.13 \\
-      Conceptual Questions & 5.41 & 10.54 \\
-      Complete Solution & 6.55 & 10.14 \\
-      \hline
-    \end{tabular}
-    \label{tab:averages}
-  \end{center}
-\end{table*}
 
@@
 demographic groups, academic disciplines, cultural backgrounds, and levels of
-previous experience \citep{catalan21} \citep{neo22}. Consequently, this study is
+previous experience \cite{catalan21} \cite{neo22}. Consequently, this study is
 limited to a specific group of students and focuses on the use of a single GenAI
@@
 more targeted assistance, an AI model is being developed to automatically
-classify student messages, offering valuable information on how these tools are used.
+classify student messages, offering valuable information on how these tools are
+used.
 
-\bibliographystyle{apalike}
-{\small
-\bibliography{References}}
-
+\begin{credits}
+\subsubsection{\ackname}
+Thanks to Instituto Federal de Educação, Ciência e Tecnologia do Rio Grande do
+Sul (IFRS) for partially funding this research
+\end{credits}
+%
+% ---- Bibliography ----
+%
+% BibTeX users should specify bibliography style 'splncs04'.
+% References will then be sorted and formatted in the correct style.
+%
+\bibliographystyle{splncs04}
+\bibliography{References}
 \end{document}

### 2024-11-19 CodeExamples/states/states.py
states

@@ -1,0 +1,97 @@
+import pandas as pd
+
+def generate_sequential_state_diagram(input_csv_path, output_puml_path):
+    """
+    Gera um diagrama de estados no formato .puml com momentos bloqueados e transições numerados sequencialmente.
+
+    Regras:
+    - Cada estado representa um tópico único.
+    - Momentos bloqueados dentro do estado são numerados sequencialmente.
+    - Transições entre estados são numeradas sequencialmente.
+    - Se o bloqueio tiver apenas uma interação, o tempo será "0 min".
+    - Transições podem incluir "interleaving" ou "interleaving and spacing" com o tempo de spacing.
+
+    Args:
+    - input_csv_path: Caminho para o arquivo CSV.
+    - output_puml_path: Caminho para salvar o arquivo .puml gerado.
+    """
+    # Carregar o CSV
+    data = pd.read_csv(input_csv_path, sep=';', encoding='utf-8')
+    # Converter colunas para os formatos corretos
+    data['Datetime'] = pd.to_datetime(data['Datetime'], format='%H:%M:%S - %d/%m/%Y', errors='coerce')
+    data = data.sort_values(by=['Id', 'Datetime'])  # Ordenar por estudante e tempo
+
+    # Calcular a duração entre interações em minutos
+    data['Time_Diff'] = data['Datetime'].diff().dt.total_seconds().fillna(0) / 60
+
+    # Identificar momentos bloqueados
+    data['Block_Number'] = (data['Time_Diff'] > 60).cumsum()
+
+    # Iniciar o conteúdo do PUML
+    puml_content = "@startuml\n\n"
+
+    # Definir estados por tópico
+    topics = data['Topic'].unique()
+    for topic in topics:
+        if pd.notna(topic):
+            alias = topic.replace(" ", "_")
+            puml_content += f'state "{topic}" as {alias}\n'
+
+    # Adicionar momentos bloqueados numerados dentro de cada estado
+    block_id = 1  # Numeração sequencial para bloqueios
+    grouped = data.groupby('Topic')
+    for topic, group in grouped:
+        if pd.isna(topic):
+            continue
+
+        alias = topic.replace(" ", "_")
+        group = group.sort_values(by=['Block_Number'])  # Ordenar por bloqueio
+        state_content = ""
+
+        for _, block_group in group.groupby('Block_Number'):
+            if len(block_group) > 1:
+                block_time = (block_group['Datetime'].iloc[-1] - block_group['Datetime'].iloc[0]).total_seconds() / 60
+            else:
[…]

### 2025-01-14 CSEDU2025/Paper.tex
paper update

@@
 \title{
-  Analyzing Student Use of Spacing and Interleaving Strategies in Interactions
+  AAnalyzing Student Use of Spacing and Interleaving Strategies in Interactions
   with GenAI-Powered Chatbots in Programming Courses
@@
 mitigate this negative effect.
-
 \begin{table*}[htbp]
@@
   \begin{center}
-    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{4cm} p{12cm}} % Remove vertical bars
+    \renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
+    \begin{tabular}{p{3.8cm} p{6cm} p{4.5cm}} % Remove vertical bars
       \hline
-      \textbf{Category} & \textbf{Description} \\
+      \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
       \hline
-      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. \\
-      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. \\
-      Student Correction (SC) & Prompts where the student corrects the chatbot. \\
-      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. \\
-      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. \\
-      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). \\
-      Language Change (LC) & Prompts that request a change of idiom. \\
-      Uncategorized (U) & Prompts that do not fit into any of the above categories. \\
+      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
+      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
+      Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
+      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
+      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
+      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
+      Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
+      Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
       \hline
@@
 
-When a student requests a code snippet (CS), it is generally predictable that
-they will resume interaction with the chatbot within a short period of time, as
-indicated by the mean (1.26) and standard deviation (1.74) of the Code Snippet
+The objective of this research question was to study the time a student takes to
+re-engage with the bot after sending a message for each category. This analysis
+can offer insights into students' behavior: faster interactions might reflect
+less reflection, whereas longer pauses could indicate the opposite.
+
+To calculate the average time and standard deviation for messages within a
+category initially was first observed conversations in which message timing was
+influenced by previous messages. Pearson's correlation analysis revealed that
+28.40\% of the conversations exhibited a strong correlation, which, in this
+study, was defined as values outside the range of -0.3 to 0.3. In this context,
+a strong correlation suggests that the response time for a message may depend
+on the time of the preceding message or multiple prior messages. Consequently,
+these 28.40\% of conversations were excluded from the average time and standard
+deviation calculation, leaving 56 out of the initial 86 conversations. Table
+\ref{tab:averages} presents the means and standard deviations for each category.
+
+When a student requests a Code Snippet (CS), it is generally predictable that
+they will summarize interaction with the chatbot within a short period, as
+indicated by the mean (1.34) and standard deviation (1.95) of the Code Snippet
 category in Table \ref{tab:averages}. This result is noteworthy because,
 although some requests are simple and lead to quick responses, others may be
-more complex and would require more time for reflection. However, the low
-average response time and minimal variability may indicate that students adopt a
-more passive learning posture when requesting the chatbot to generate small pieces
-of code.
+more complex and require more reflection time. However, the low average response
+time and minimal variability may indicate that students adopt a more passive
+learning posture when requesting the chatbot to generate pieces of code.
 
@@
 
-Despite some occasional variation (3.54), prompts classified as Uncategorized
-exhibit a relatively low average time (1.65) for students to resume interaction
-with the chatbot. This variation occurs, in part, because of situations where students
-end the conversation with an uncategorized message, such as a simple
-\textit{"thanks"}, but later choose to resume the interaction.
+Despite occasional variation (3.59), prompts classified as Uncategorized exhibit
+a relatively low average time (1.70) for students to resume interaction with the
+chatbot. This variation occurs partly because of situations where students end
+the conversation with an Uncategorized message like \textit{"thanks"} but later
+decide to resume the interaction.
 
-When a student requests the chatbot to change the language (LC) in a response, the
-new interaction occurs quickly (1.89) with little variation in response time
-(3.28), meaning that it is predictable that the student will continue interacting.
+When a student requests the chatbot to Change the Language (LC)  in a response,
+in our context English to Spanish, the new interaction occurs quickly (2.23)
+with slight variation in response time (3.71), meaning that it is predictable
+that the student will continue interacting.
 
-Although still below the overall average, students take a moderate amount of
-time (2.65) to resume interaction with the chatbot after requesting the solution to
-a multiple-choice question (MQ). The variation (6.10) suggests that some
-responses may lead to a slower follow-up interaction, possibly indicating a
-moment of deeper reflection on the part of the students. However, since both the
-mean and standard deviation are below the overall average, this could indicate
-a more passive behavior, as students obtain ready-made answers from the chatbot.
+Although still below the overall average, students take moderate time (2.47) to
+resume interaction with the chatbot after requesting the solution to a Multiple
+Question (MQ) exercise. The variation (5.56) suggests that some responses may
+lead to a slower follow-up interaction, possibly indicating a moment of deeper
+reflection on the part of the students. However, since the mean and standard
+deviation are below the overall average, this could indicate a more passive
+behavior, as students obtain ready-made answers from the chatbot.
 
-The average time for students to resume interaction in the correction category
-(SC) is close to the overall average (3.17), although the high standard
-deviation (7.33) indicates significant variation. In some cases, students are
-already familiar with the answer and only seek a quick correction from the chatbot;
-in others, they detect flaws in the chatbot's responses and need to carefully
-examine the subsequent answer to identify potential errors, which suggests a
-more active behavior.
+The average time of the Student Correction (SC) category is close to the overall
+average (4.07), although the high standard deviation (8.42) indicates
+significant variation. In some cases, students know the answer to an exercise
+and only accomplish a quick correction from the chatbot; in others, they detect
+flaws in the chatbot's responses and need to examine the subsequent answer to
+identify potential errors, which suggests a more active behavior.
 
-After a question classified as Debugging Help (DH), there is both a
-longer-than-average response time (4.84) and a high variation in the time taken
-for a new interaction (10.13). This suggests that student prompts in this
-category tend to be more time-consuming and unpredictable. A possible
-explanation is that some DH issues are simple, such as '\textit{what does x++ do?}',
-while others involve requests for analysis of more complex code, like
+A message classified as Debugging Help (DH) has both a long average response
+time (5.30) and high variation (10.80). Consequently, the behavior in this
+category is more time-consuming and unpredictable. A possible explanation is
+that some DH messages are simple, such as '\textit{what does x++ do?}'. In
+contrast, others involve requests for analysis of more complex code, like
 \textit{'explain this piece of code (accompanied by a complete method).'}
 
-Conceptual questions (CQ) generally result in a high average time (5.41) for
+Conceptual Questions (CQ) generally result in a high average time (5.47) for
 students to reengage with the chatbot. Moreover, the high standard deviation
-(10.54) indicates an unpredictable variation in response times. This
+(10.46) indicates an unpredictable variation in response times. This
 unpredictability may be attributed to the fact that, in some cases, the
@@
 Finally, a message classified as a Complete Solution request (SCO) exhibits the
-longest average time (6.55) for students to resume interaction with the chatbot.
-Additionally, the high standard deviation (10.14) highlights significant
+longest average time (7.65) for students to resume interaction with the chatbot.
+Additionally, the high standard deviation (11.03) highlights significant
 unpredictability in response time. After requesting a complete solution,
-students tend to take longer to return due to the complexity of the problems
-involved, such as the implementation of sorting algorithms such as Heapsort. Table
-\ref{tab:averages} presents the means and standard deviations for each
-category.
-
-However, some categories suggest quicker interactions with less reflection on
-the part of students, such as 'Code Snippet' and 'Multiple Choice.' On the other
-hand, other categories indicate that students may be in a more reflective
-process, such as 'Debugging Help,' 'Conceptual Question,' and 'Complete
-Solution.' This demonstrates that the level of student engagement varies
-according to the nature of the requested interaction, reflecting different
-approaches and depths in the learning process.
+students may take longer to return due to the complexity of the problems
+involved, such as the implementation of sorting algorithms such as Heapsort.
 
@@
   \begin{center}
-    \renewcommand{\arraystretch}{1.6} % Increase the spacing between rows
-    \begin{tabular}{p{4cm} p{4cm} p{4cm}} % Remove vertical bars
+    \renewcommand{\arraystretch}{1.2} % Adjust the spacing between rows
+    \begin{tabular}{p{4cm} p{3cm} p{3cm}} % Remove vertical bars
       \hline
@@
       \hline
-      Code Snippet & 1.26 & 1.74 \\
-      Uncategorized & 1.65 & 3.54 \\
-      Language Change & 1.89 & 3.28 \\
-      Multiple Question & 2.65 & 6.10 \\
-      Student Correction & 3.17 & 7.33 \\
+      Code Snippet & 1.34 & 1.95 \\
+      Uncategorized & 1.70 & 3.59 \\
+      Language Change & 2.23 & 3.71 \\
+      Multiple Question & 2.47 & 5.66 \\
       \hdashline
-      \textit{Overall} & \textit{3.43} & \textit{6.60} \\
+      \textit{Overall} & \textit{3.78} & \textit{6.95} \\
       \hdashline
-      Debugging Help & 4.84 & 10.13 \\
-      Conceptual Questions & 5.41 & 10.54 \\
-      Complete Solution & 6.55 & 10.14 \\
+      Student Correction & 4.07 & 8.42 \\
+      Debugging Help & 5.30 & 10.80 \\
+      Conceptual Questions & 5.47 & 10.46 \\
+      Complete Solution & 7.65 & 11.03 \\
       \hline
@@
 
-The objective of this work was to analyze the interactions of Java programming
-students with an educational chatbot developed using the RAG technique, employing
-two metacognitive study strategies: spacing and interleaving.
+This study aimed to analyze the interactions of Java programming students with
+an educational chatbot developed using the RAG technique, employing two
+metacognitive study strategies: Spacing and Interleaving.
 
-The metacognition theory, along with the classification of student actions,
-proved to be effective in understanding and revealing how students use the chatbot.
-The results showed that most messages fall into the categories of Conceptual
-Questions and Debugging Help, both known for promoting slower engagement and
-potentially fostering more active learning. However, the results also indicated
-that students' use of study strategies such as spacing and interleaving remains
-limited, from which we can conclude that encouraging students to apply these
-strategies more intentionally could enhance their learning while using GenAI.
+The results indicated that messages from the Code Snippets and Multiple Question
+(Quiz) category implied quick and consistent interactions with the bot. This
+pattern reflects a good level of engagement with the tool but also raises the
+possibility of a lack of more profound reflection on the received content, given
+that students' subsequent messages were sent consistently within short intervals.
 
-It is important to consider how GenAI tools impact students in various
+On the other hand, it was observed that most students' messages concentrated on
+categories such as Debugging Help and Conceptual Questions, which might suggest
+that students were committed to understanding the concepts. However, the results
+also indicated that students' use of study strategies such as Spacing and
+Interleaving remains limited. Therefore, encouraging students to apply these
+strategies intentionally could enhance their learning using GenAI.
+
+It is fundamental to consider how GenAI tools impact students in various
 demographic groups, academic disciplines, cultural backgrounds, and levels of
 previous experience \citep{catalan21} \citep{neo22}. Consequently, this study is
-limited to a specific group of students and focuses on the use of a single GenAI
+limited to a specific group of students and focuses on using a single GenAI
 tool. As a result, the findings may not be generalizable to other populations or
@@
 
-To further explore this, a broader study is planned to examine, for instance,
-the relationship between student profiles, use of metacognitive scaffolds, their
-effective study strategies in using a GenAI, and learning outcomes.
-In addition, to help educators analyze student behavior and provide
-more targeted assistance, an AI model is being developed to automatically
-classify student messages, offering valuable information on how these tools are used.
+There are ways to deepen the study; based on previous references, 60 minutes was
+the standard spacing measure. However, further studies will deepen this analysis
+using a quantitative variable to represent different spacing intervals better.
+Future research can also include additional metacognitive strategies, which will
+involve analyzing and interpreting user messages in depth. For instance, the
+Monitoring strategy was evident when students expressed that a concept remained
+unclear and requested further clarification. Another necessary analysis involves
+investigating the relationship between student profiles, metacognitive
+strategies for using GenAI, and learning outcomes.
 

### 2025-01-15 CSEDU2025/Paper.tex
update paper

@@
   \centering
-  \includegraphics[scale=0.45]{img/figure2.png}
+  \includegraphics[scale=0.42]{img/figure2.png}
   \caption{Examples of conversations.}
@@
 
-For the analysis of the spacing, we initially considered a minimum interval of
-60 minutes as a reference to define the significant spacing \citep{Gadella24}.
-Based on this, our results showed that, of the 81 conversations analyzed, only
-27 (33.3\%) had at least one message with an interval greater than 60 minutes.
-Figure 2 illustrates the elapsed time between messages (delta). For example,
-in conversation 15, the first message was sent at time zero, while the second
-message was sent by the student after a delta of 0.6 minutes. Additionally,
-message 4 in conversation 15, as shown in Figure \ref{fig:graph2}, presents a
-delta of 69.1 minutes, which, according to our reference, characterizes
-significant spacing. Figure \ref{fig:graph3} also shows the number of spacings
-in each conversation. Among conversations that exhibited spacing, the
-average number of subsections per conversation was 2.2. Therefore, a
-conversation with one instance of spacing is typically divided into two distinct
-study subsections.
+For the spacing analysis, we initially considered a minimum interval of 60
+minutes as a reference to define the significant spacing \citep{Gadella24}.
+Based on this, our results showed that, of the 81 conversations analyzed, just
+27, or around 34\%, had at least one message with an interval more significant
+than 60 minutes. Figure \ref{fig:graph2} illustrates the elapsed time between
+messages (delta). For example, in conversation 15, the first message was sent
+at time zero, while the student sent the second message after a delta of 0.6
+minutes. Additionally, message 4 in conversation 15, as shown in Figure
+\ref{fig:graph2}, presents a delta of 69.1 minutes, which, according to our
+reference, characterizes spacing. The average number of subsections per
+conversation was 2.2 among conversations that exhibited spacing. Therefore, a
+conversation with one spacing is divided into two distinct study sessions.
 
 This data suggests that there is indeed little spacing between study sessions.
-It is worth noting that, in order to create a new section in CharlieBot,
-students must close the browser tab and start a new conversation, which results
-in losing the previous chat history.
+To create a new section in CharlieBot; students must close the browser tab and
+start a new conversation, which results in losing the previous chat history.
 
@@
 re-engage with the bot after sending a message for each category. This analysis
-can offer insights into students' behavior: faster interactions might reflect
-less reflection, whereas longer pauses could indicate the opposite.
+can offer insights into students' behavior: faster interactions might with the
+chatbot might suggest a more passive learning posture.
 
@@
 learning posture when requesting the chatbot to generate pieces of code.
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.39]{img/figure3.png}
-  \caption{Subsections per conversation. \textit{A zero value indicates
-  no subsection and no spacing.}}
-  \label{fig:graph3}
-\end{figure*}
 
@@
 
-To analyze the interweaving of topics, sections were initially divided into
-those that exhibited spacing and those that did not.
+Switching between topics can help students identify distinct concepts inside
+problems. The results show that approximately 50.8\% of the conversations
+include interleaving; consequently, 49.2\% are related to a single study topic.
+To gain a deeper comprehension of this interleaving, conversations were analyzed
+ by separating those without spacing from those that included spacing.
 
-In conversations with the chatbot that did not include spacing (66.7\% of the total
-conversations), approximately 19.2\% of the messages resulted in a change of
-topic. Qualitative analysis of what led to this interweaving indicates that
-part of it was due to students' requests for assistance in solving exercises
-from different topics, suggesting a less active learning approach, as the time
-interval for this type of interaction tends to be shorter than average.
-Additionally, some instances of interweaving were observed after a prolonged
-period of 20 to 50 minutes, still below the 60-minute threshold established as
-the cutoff for considering spacing.
+Conversations without spacing comprise around 66\% of the total, with 34.2\%
+showing a topic change, indicating interleaving (equivalent to 22.8\% of all
+conversations). Qualitative analysis of these conversations without spacing
+revealed that about 7\% of the conversations presented interleaving caused by
+students' at least one request to solve quizzes. However, the speed with which
+students interact with the chatbot to solve quizzes may suggest a more
+superficial reflection on the answers received (see section 4.2.1). On the other
+hand, 8.8\% of the conversations showed no spacing but included topic changes,
+suggesting that students effectively might employ a metacognitive interleaving
+strategy to distinguish concepts around a problem. Finally, 7\% of the
+conversations featured interaction pauses lasting between 20 and 59 minutes but
+fell short of the study's definition of spacing, which required pauses of over
+60 minutes. This subset of conversations revealed a natural interleaving of
+topics, reflecting a chatbot usage pattern that often mirrors how users search
+Web pages.
 
-In interactions with the chatbot that included spacing (33.3\% of the total
-conversations), approximately 31.4\% of the messages switched topics.
-Qualitative analysis of these interactions reveals that a significant portion
-of this interweaving was caused by spacing. In other words, some students use
-the chatbot for quick consultations, asking a question and then returning hours
-later with another question on a different subject.
+Conversations with spacing comprise approximately 34\% of the total, with 84.2\%
+indicating interleaving (equivalent to 28\% of all conversations). Pauses longer
+than 60 minutes revealed similar behavior to those shorter than 60 minutes, with
+students using the chatbot primarily for quick consultations, asking questions,
+and returning at least one hour late with a query about a different study topic.
+The argument that some students use the chatbot primarily as a tool for quick
+queries is supported by data showing an average of 4.18 messages per study
+session. Interestingly, only 6\% of the conversations involved students pausing
+their interaction with the chatbot for more than 60 minutes before sending a
+message related to the exact topic of study.
 
-Thus, the data revealed that the existing interweaving is limited and does not
-suggest an effective study strategy, as it mainly arises from requests for
-exercise resolution and from the natural spacing pattern of a consultation
-interaction.
+Therefore, although approximately 50.8\% of the conversations contain
+interleaving, the data shows that this topic switching is limited in terms of
+using a conscious metacognitive strategy.
 

### 2025-01-15 CSEDU2025/Paper.tex
paper update

@@
   \begin{center}
-    \renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
+    \renewcommand{\arraystretch}{1.1} % Increase the spacing between rows
     \begin{tabular}{p{3.8cm} p{6cm} p{4.5cm}} % Remove vertical bars
@@
 contextual statements like '\textit{I'm reviewing object orientation}', as well
-as requests for additional exercises and summaries. Depending on the type of
-investigation, there is opportunity for these messages to be further subdivided
-into new categories.
+as requests for additional exercises and summaries. These Uncategorized messages
+can be subdivided into new categories depending on the investigation. For
+instance, requests for new exercises or summaries could be interpreted as a
+metacognitive monitoring strategy.
 
 Finally, 1.3\% of the messages were classified as Language Change. These prompts
-were typically requests for the chatbot to switch languages, as seen in
-conversation 9 in Figure \ref{fig:graph2}.
+were typically requests for the chatbot to switch languages (English to Spanish),
+as seen in Figure \ref{fig:graph2}, conversation 9, message 2.
 
@@
 
-On the other hand, it was observed that most students' messages concentrated on
-categories such as Debugging Help and Conceptual Questions, which might suggest
-that students were committed to understanding the concepts. However, the results
-also indicated that students' use of study strategies such as Spacing and
-Interleaving remains limited. Therefore, encouraging students to apply these
-strategies intentionally could enhance their learning using GenAI.
+The results also indicated that students' use of study strategies such
+as Spacing and Interleaving remains limited. Therefore, encouraging students to
+apply these strategies intentionally could enhance their learning using GenAI.
 

### 2025-01-21 CSEDU2025/Paper.tex
paper update

@@
 \title{
-  AAnalyzing Student Use of Spacing and Interleaving Strategies in Interactions
+  Analyzing Student Use of Spacing and Interleaving Strategies in Interactions
   with GenAI-Powered Chatbots in Programming Courses
@@
 \abstract{
-The aim of this study is to analyze the interactions of programming students
-with a chatbot powered by OpenAI's GPT-3.5, enhanced with the Retrieval
-Augmented Generation (RAG) technique, within the context of a Java programming
-course. The study investigates how students use the chatbot while employing
-two metacognitive strategies: interleaving and spacing. The methodology involved
-classifying and categorizing the students' interaction prompts with the chatbot
-into eight distinct categories. The findings reveal that, while most
-interactions focus on active learning strategies, students demonstrate limited
-conscious application of metacognitive strategies. The study highlights the need
-for more comprehensive guidance on the strategic use of AI tools to enhance
-learning outcomes.
+This study aimed to analyze prompts of programming students with a chatbot
+powered by OpenAI's GPT-3.5, enhanced with the Retrieval Augmented Generation
+(RAG) technique, within the context of a Java programming course. The focus was
+on students using two metacognitive strategies: interleaving and spacing.
+Student prompts were categorized into eight categories along with their
+respective study topics. Findings revealed that the categories and markers of
+spacing and interleaving were important in identifying study sessions with the
+chatbot. However, students showed limited intentional application of these
+learning strategies. These results highlight the need for more comprehensive
+guidance on leveraging AI tools to improve learning outcomes.
 }
@@
 learning and the application of student's knowledge in other contexts and
-situations \citep{Rivers21}.
+situations \citep{rohrer20} \citep{Rivers21} .
 
@@
 
-\cite{Boudouaia24} investigated the impact of ChatGPT-facilitated programming on
-college students’ programming behaviors, performance, and perceptions. The
-problem addressed was whether the use of ChatGPT in programming education could
-influence how students interacted with programming tasks, their learning
-outcomes, and their attitudes towards technology. The study employed a
-quasi-experimental design involving 82 college students divided into two groups:
-one using ChatGPT-assisted Programming (CAP) and the other using traditional
-self-directed Programming (SDP). The method combined mixed data collection
-techniques, including behavioral logs, programming performance evaluations, and
-interviews. The results showed that students using the CAP mode exhibited more
-frequent behaviors such as debugging and reading feedback. Although CAP students
-had a higher average score in programming performance, there was no
-statistically significant difference compared to the SDP group. However,
-students’ perceptions of ChatGPT improved significantly after the intervention,
-with increased perceived usefulness, ease of use, and intention to continue
-using the tool in the future.
+% \cite{Boudouaia24} investigated the impact of ChatGPT-facilitated programming on
+% college students’ programming behaviors, performance, and perceptions. The
+% problem addressed was whether the use of ChatGPT in programming education could
+% influence how students interacted with programming tasks, their learning
+% outcomes, and their attitudes towards technology. The study employed a
+% quasi-experimental design involving 82 college students divided into two groups:
+% one using ChatGPT-assisted Programming (CAP) and the other using traditional
+% self-directed Programming (SDP). The method combined mixed data collection
+% techniques, including behavioral logs, programming performance evaluations, and
+% interviews. The results showed that students using the CAP mode exhibited more
+% frequent behaviors such as debugging and reading feedback. Although CAP students
+% had a higher average score in programming performance, there was no
+% statistically significant difference compared to the SDP group. However,
+% students’ perceptions of ChatGPT improved significantly after the intervention,
+% with increased perceived usefulness, ease of use, and intention to continue
+% using the tool in the future.
+
+The study of \cite{Boudouaia24} examined the effects of ChatGPT-assisted
+programming on university students' behaviors, performance, and perceptions.
+A quasi-experimental research was conducted with 82 students divided into two
+groups: one with ChatGPT-assisted programming (CAP) and the other with
+self-directed programming (SDP). The analysis included behavioral logs,
+performance evaluations, and interviews. Students in the CAP group engaged more
+actively in debugging and feedback review activities. Although they achieved
+slightly higher scores, there was no statistically significant difference in
+performance compared to the SDP group. Nevertheless, perceptions of ChatGPT
+improved significantly, highlighting greater perceived usefulness, ease of use,
+and intention to use the tool in the future.
 
@@
 mitigate this negative effect.
+
 \begin{table*}[htbp]
@@
 thinking, and metacognitive skills in collaborative programming, observing
-improvements in problem-solving and collaboration. \cite{Khusnul24} investigates
-the efficacy of a meta-learning approach in improving metacognitive and creative
-skills. This study confirms the effectiveness of metacognition strategies and
-elucidates the relationship between meta-learning and metacognition. This
-research suggests that meta-learning can improve metacognitive abilities,
-providing valuable insights into educational technology and course design in
-higher education settings.
 
-\cite{Sidra24} addressed the growing importance of metacognitive skills in an
-AI-enabled workforce. The problem discussed was how AI systems, especially in
-collaborative tools based on Large Language Models (LLMs), impacted human
-decision-making and the challenges they posed due to the lack of human-like
-metacognition in AI. The method used identified four main characteristics that
-differentiated human-AI interaction from human-human collaboration and proposed,
-based on the dual-process theory of human thought, that metacognitive thinking
-was crucial for the effective use of AI. The results indicated that human
-workers needed to develop and apply metacognitive skills, such as planning,
-monitoring, and evaluating their thought processes, to mitigate cognitive biases
-and improve decision-making when working with AI. The study concluded by
-recommending improvements in AI design and training to improve workers'
-metacognitive abilities, ensuring more effective collaboration with AI systems.
+% \cite{Khusnul24} investigates
+% the efficacy of a meta-learning approach in improving metacognitive and creative
+% skills. This study confirms the effectiveness of metacognition strategies and
+% elucidates the relationship between meta-learning and metacognition. This
+% research suggests that meta-learning can improve metacognitive abilities,
+% providing valuable insights into educational technology and course design in
+% higher education settings.
+
+% \cite{Sidra24} addressed the growing importance of metacognitive skills in an
+% AI-enabled workforce. The problem discussed was how AI systems, especially in
+% collaborative tools based on Large Language Models (LLMs), impacted human
+% decision-making and the challenges they posed due to the lack of human-like
+% metacognition in AI. The method used identified four main characteristics that
+% differentiated human-AI interaction from human-human collaboration and proposed,
+% based on the dual-process theory of human thought, that metacognitive thinking
+% was crucial for the effective use of AI. The results indicated that human
+% workers needed to develop and apply metacognitive skills, such as planning,
+% monitoring, and evaluating their thought processes, to mitigate cognitive biases
+% and improve decision-making when working with AI. The study concluded by
+% recommending improvements in AI design and training to improve workers'
+% metacognitive abilities, ensuring more effective collaboration with AI systems.
+
+\cite{Sidra24} explored the importance of metacognitive skills in an AI-enabled
+workforce, focusing on how AI, particularly tools using Large Language Models,
+affects human decision-making. They highlighted challenges arising from AI's
+lack of human-like metacognition and identified key differences between human
+AI and human-human collaboration. Using dual-process theory, they emphasized the
+need for workers to develop metacognitive skills like planning, monitoring, and
+evaluating to counter cognitive biases and enhance decision-making. The study
+recommended improving AI design and training to boost workers' metacognitive
+abilities for more effective collaboration.
 
@@
 
-CharlieBot is an educational chatbot built on ChatGPT 3.5 and enhanced with
-Retrieval-Augmented Generation (RAG) technique to improve the precision of
-responses provided to students \citep{Sun24}. The RAG approach was developed
-using materials from a MOOC, including text, exercises, video transcripts, and
-other resources.
+CharlieBot is an educational chatbot powered by ChatGPT 3.5 and enhanced with
+Retrieval-Augmented Generation (RAG) \citep{Sun24}. RAG is an AI technique that
+integrates information retrieval with generative models. It first retrieves
+relevant documents or data from a knowledge base or external source using a
+retrieval model. Then, a generative model uses this retrieved information to
+generate more accurate, context-aware responses. This dual-step process enhances
+the quality and reliability of the chatbot's answers. A prior study on
+CharlieBot's performance revealed that most students found its responses
+appropriate for the Java programming course \citep{Hoyos24}.
 
@@
 During the data collection phase, students enrolled in a second-semester Java
-course at a public university in Spain were introduced to CharlieBot and given
-the freedom to use it without following any prescribed educational methodology.
-All data were collected anonymously to ensure that interactions could not be
-traced back to individual students or linked to their academic performance.
+course at the University Carlos III of Madrid (UC3M) were introduced to
+CharlieBot and allowed to use it without following any prescribed educational
+methodology. All data were collected anonymously to ensure that interactions
+could not be traced back to individual students or linked to their academic
+performance.
 
-In the categorization phase, the students' messages were manually classified
-into eight distinct categories for analysis, representing different types of
-prompt that students use when interacting with the chatbot. Initially,
-\cite{Ghimire24} proposed four categories: Debugging Help (DH), Conceptual
-Question (CQ), Code Snippet (CS), and Complete Solution (CSO). However, the data
-collected indicated the need for additional categories, leading to the inclusion
-of four more: Multiple Question (MQ), Student Corrections (SC), Language Change
-(LC), and Uncategorized (U). In this study, students were free to use the
-chatbot as they wanted, which contributed to the emergence of these additional
-categories. Table \ref{tab:categories} presents these categories along with
-their respective descriptions. To facilitate the analysis, the prompt topics
-were also categorized according to the course subjects, which are: (1) Java, (2)
-Object Orientation, (3) Testing, (4) Recursion, (5) Data Structures, and (6)
-Sorting and Searching Algorithms.
+During the categorization phase, the students' prompts were initially classified
+into eight distinct categories using the Claude.ai \citep{claude} tool.
+Subsequently, the authors of this study manually reviewed these classifications
+to ensure accuracy and alignment with the research objectives.
 
-The analysis phase involved the use of Python scripts to extract information
-from previously classified data. The data sets and code used in this process are
-available at: https://github.com/rodrigoprestesmachado/uc3m.
+Initially, \cite{Ghimire24} proposed four categories: Debugging Help (DH),
+Conceptual Question (CQ), Code Snippet (CS), and Complete Solution (CSO).
+However, the data collected indicated the need for additional categories,
+leading to the inclusion of four more: Multiple Questions (MQ), Student
+Corrections (SC), Language Change (LC), and Uncategorized (U). Unlike study
+\cite{Ghimire24}, which analyzed students' behavior using artificial
+intelligence for specific programming tasks, our study allowed participants to
+explore the chatbot as they preferred. This approach enabled the emergence of
+additional categories of analysis, broadening the understanding of the tool's
+uses and interactions. Table \ref{tab:categories} presents these categories,
+their descriptions, and an example. Besides that, the topics covered in the
+students' prompts were categorized based on the course issues, which include
+(1) Java, (2) Object Orientation, (3) Testing, (4) Recursion, (5) Data
+Structures, and (6) Sorting and Searching Algorithms.
+
+The analysis phase involved using Python/Pandas scripts to extract information
+from previously classified data. The data sets, and code used in this process
+are available at https://github.com/rodrigoprestesmachado/uc3m.
 
@@
   \centering
-  \includegraphics[scale=0.62]{img/figure1.png}
+  \includegraphics[scale=0.61]{img/figure1.png}
   \caption{Categorization of messages.}
@@
   \centering
-  \includegraphics[scale=0.42]{img/figure2.png}
+  \includegraphics[scale=0.41]{img/figure2.png}
   \caption{Examples of conversations.}
@@
   \begin{center}
-    \renewcommand{\arraystretch}{1.2} % Adjust the spacing between rows
+    \renewcommand{\arraystretch}{1.1} % Adjust the spacing between rows
     \begin{tabular}{p{4cm} p{3cm} p{3cm}} % Remove vertical bars
@@
 60 minutes. This subset of conversations revealed a natural interleaving of
-topics, reflecting a chatbot usage pattern that often mirrors how users search
-Web pages.
+topics, reflecting a chatbot usage pattern.
 
@@
 
-The results indicated that messages from the Code Snippets and Multiple Question
-(Quiz) category implied quick and consistent interactions with the bot. This
-pattern reflects a good level of engagement with the tool but also raises the
-possibility of a lack of more profound reflection on the received content, given
-that students' subsequent messages were sent consistently within short intervals.
+The analysis categories, spacing, and interleaving markers provided insights
+into students' interactions with the chatbot. With the concept of spacing, it
+was possible to define students' study sessions, which, in this work, consisted
+of a few interaction messages with the chatbot. Consequently, a significant
+portion of the observed interleaving occurred due to spacing, that is, in the
+transition between study sessions. These findings indicate that the intentional
+use of learning strategies, such as spacing and interleaving, is still limited.
+Therefore, encouraging the deliberate application of these practices can enhance
+students' learning outcomes using generative artificial intelligence (GenAI)
+tools.
 
-The results also indicated that students' use of study strategies such
-as Spacing and Interleaving remains limited. Therefore, encouraging students to
-apply these strategies intentionally could enhance their learning using GenAI.
+Messages from the Code Snippets and Multiple Question (Quiz) categories stood
+out for involving consistently quick interactions with the bot. This pattern
+suggests some level of engagement with the chatbot but also raises the
+possibility of less in-depth reflection on the content received, especially when
+students request small code snippets or the resolution of exercises.
 
@@
 
-There are ways to deepen the study; based on previous references, 60 minutes was
-the standard spacing measure. However, further studies will deepen this analysis
-using a quantitative variable to represent different spacing intervals better.
-Future research can also include additional metacognitive strategies, which will
-involve analyzing and interpreting user messages in depth. For instance, the
+There are ways to deepen this study; based on previous references, 60 minutes
+was the standard spacing measure. However, further studies will deepen this
+analysis using a quantitative variable to better represent different spacing
+intervals. Future research can also include additional metacognitive strategies,
+which will involve interpreting user messages in depth. For instance, the
 Monitoring strategy was evident when students expressed that a concept remained

### 2025-01-23 JavaLab/Recap/Section3.java
java classes

@@ -1,0 +1,42 @@
+package edu.uc3m;
+
+/**
+ * The Section3 program processes command-line arguments to calculate the sum of integers.
+ * If no arguments are provided, the program outputs Integer.MIN_VALUE.
+ * Invalid numeric inputs are ignored during the sum calculation.
+ *
+ * Usage examples:
+ * - java edu.uc3m.Section3 1 2 3 -> Output: 6
+ * - java edu.uc3m.Section3 a 5 b 3 -> Output: 8 (ignores invalid inputs "a" and "b")
+ * - java edu.uc3m.Section3 -> Output: -2147483648 (value of Integer.MIN_VALUE)
+ */
+public class Section3 {
+
+    /**
+     * Main method to process command-line arguments and calculate the sum of integers.
+     * If no arguments are provided, the program outputs Integer.MIN_VALUE.
+     * Invalid inputs that cannot be parsed as integers are ignored.
+     *
+     * @param args the command-line arguments. These are expected to be integers,
+     *             but non-numeric arguments are ignored.
+     */
+    public static void main(String[] args) {
+        if (args.length == 0) {
+            // Output Integer.MIN_VALUE if no arguments are provided
+            System.out.println(Integer.MIN_VALUE);
+        } else {
+            int sum = 0; // Stores the sum of valid integers
+            for (int i = 0; i < args.length; i++) {
+                try {
+                    // Parse each argument and add to the sum
+                    sum = sum + Integer.parseInt(args[i]);
+                } catch (NumberFormatException e) {
+                    // Ignore invalid inputs
+                    continue;
+                }
+            }
+            // Output the total sum of valid integers
+            System.out.println(sum);
+        }
+    }
+}

### 2025-01-23 JavaLab/Recap/Section4.java

@@ -1,0 +1,71 @@
+package edu.uc3m;
+
+/**
+ * This program processes command-line arguments to calculate either the sum or the maximum
+ * of a series of integers. The first argument specifies the operation ("sum" or "max"),
+ * and the subsequent arguments are the numbers to process. Non-numeric arguments are ignored.
+ *
+ * Usage examples:
+ * - java edu.uc3m.Section4 sum 1 2 3 -> Output: Sum: 6
+ * - java edu.uc3m.Section4 max 5 10 7 -> Output: Max: 10
+ * - java edu.uc3m.Section4 sum a 5 b 3 -> Output: Sum: 8 (ignores invalid inputs "a" and "b")
+ * - java edu.uc3m.Section4 -> Output: No parameters
+ */
+public class Section4 {
+
+    /**
+     * Main method to process command-line arguments and execute the specified operation.
+     *
+     * If the first argument is "sum", the program calculates the sum of all valid integers
+     * in the subsequent arguments. If the first argument is "max", the program finds the
+     * maximum value among all valid integers in the subsequent arguments. Non-numeric
+     * arguments are ignored.
+     *
+     * If no arguments are provided, the program outputs "No parameters".
+     *
+     * @param args the command-line arguments. The first argument specifies the operation
+     *             ("sum" or "max"), and the subsequent arguments are the integers to process.
+     */
+    public static void main(String[] args) {
+        if (args.length == 0) {
+            System.out.println("No parameters");
+        } else {
+            if (args[0].equals("sum")) {
+                int sum = 0;
+                for (int i = 1; i < args.length; i++) { // Starts at 1 to skip the operation argument
+                    int num = parseIntSafe(args[i]);
+                    if (num != Integer.MIN_VALUE) { // Ensures only valid numbers are processed
+                        sum += num;
+                    }
+                }
+                System.out.println("Sum: " + sum);
+            } else if (args[0].equals("max")) {
+                int max = Integer.MIN_VALUE;
+                for (int i = 1; i < args.length; i++) { // Starts at 1 to skip the operation argument
+                    int num = parseIntSafe(args[i]);
+                    if (num != Integer.MIN_VALUE) { // Ensures only valid numbers are processed
+                        max = Math.max(max, num);
+                    }
+                }
+                System.out.println("Max: " + (max == Integer.MIN_VALUE ? "No valid numbers" : max));
+            }
+        }
+    }
+
+    /**
[…]

### 2025-01-27 ETHE/sn-article.tex
Starting ETHE paper

--- 2025-01-21 CSEDU2025/Paper.tex
@@
-\documentclass[a4paper,twoside]{article}
+\documentclass[sn-apa]{sn-jnl} % APA Reference Style
 
-\usepackage{epsfig}
-\usepackage{subcaption}
-\usepackage{calc}
-\usepackage{amssymb}
-\usepackage{amstext}
-\usepackage{amsmath}
-\usepackage{amsthm}
-\usepackage{multicol}
-\usepackage{pslatex}
-\usepackage{algorithm2e}
-\usepackage[bottom]{footmisc}
-% Please add other packages that you may need BEFORE the SCITEPRESS.sty package.
-\usepackage{natbib}
-\usepackage{multirow}
-\usepackage{SCITEPRESS}
-\usepackage{arydshln}
-\widowpenalty=10000
-\clubpenalty=10000
+%%%% Standard Packages
+
+\usepackage{graphicx}%
+\usepackage{multirow}%
+\usepackage{amsmath,amssymb,amsfonts}%
+\usepackage{amsthm}%
+\usepackage{mathrsfs}%
+\usepackage[title]{appendix}%
+\usepackage{xcolor}%
+\usepackage{textcomp}%
+\usepackage{manyfoot}%
+\usepackage{booktabs}%
+\usepackage{algorithm}%
+\usepackage{algorithmicx}%
+\usepackage{algpseudocode}%
+\usepackage{listings}%
 
@@
 
-\title{
-  Analyzing Student Use of Spacing and Interleaving Strategies in Interactions
-  with GenAI-Powered Chatbots in Programming Courses
-}
+\title[Article Title]{Analyzing Students Use Metacognitive Strategies When
+Interacting with GenAI-Powered Chatbots in Programming Courses.}
 
-\author{\authorname{Rodrigo Prestes Machado\sup{1}\orcidAuthor{0000-0003-0428-6387},
-Carlos Alario-Hoyos\sup{2}\orcidAuthor{0000-0002-3082-0814},
-Patricia Callejo\sup{2}\orcidAuthor{0000-0001-6124-6213},
-Iria Estévez-Ayres\sup{2}\orcidAuthor{0000-0002-1047-5398},
-Carlos Delgado Kloos\sup{2}\orcidAuthor{0000-0003-4093-3705}}
-\affiliation{\sup{1}Department of Informatics, Instituto Federal de Educação,
-Ciência e Tecnologia, Porto Alegre, Brazil}
-\affiliation{\sup{2}Department of Telematics Engineering, Universidad Carlos III
-de Madrid, Leganés (Madrid), Spain}
-\email{\ rodrigo.prestes@poa.ifrs.edu.br, \{calario, pcallejo, ayres, cdk\}@it.uc3m.es}
-}
+%%=============================================================%%
+%% GivenName	-> \fnm{Joergen W.}
+%% Particle	-> \spfx{van der} -> surname prefix
+%% FamilyName	-> \sur{Ploeg}
+%% Suffix	-> \sfx{IV}
+%% \author*[1,2]{\fnm{Joergen W.} \spfx{van der} \sur{Ploeg} 
+%%  \sfx{IV}}\email{iauthor@gmail.com}
+%%=============================================================%%
 
-\keywords{
-  Programming, Generative Artificial Intelligence, Chatbots, Spacing,
-  Interleaving and Metacognition
-}
+\author*[1,2]{\fnm{Rodrigo} \sur{Prestes Machado}}\email{rodrigo.prestes@poa.ifrs.edu.br}
 
-\abstract{
-This study aimed to analyze prompts of programming students with a chatbot
-powered by OpenAI's GPT-3.5, enhanced with the Retrieval Augmented Generation
-(RAG) technique, within the context of a Java programming course. The focus was
-on students using two metacognitive strategies: interleaving and spacing.
-Student prompts were categorized into eight categories along with their
+\author[2,3]{\fnm{Second} \sur{Author}}\email{iiauthor@gmail.com}
+\equalcont{These authors contributed equally to this work.}
+
+\author[1,2]{\fnm{Third} \sur{Author}}\email{iiiauthor@gmail.com}
+\equalcont{These authors contributed equally to this work.}
+
+\affil*[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{100190}, \state{State}, \country{Country}}}
+
+\affil[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{10587}, \state{State}, \country{Country}}}
+
+\affil[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{610101}, \state{State}, \country{Country}}}
+
+%%==================================%%
+%% Sample for unstructured abstract %%
+%%==================================%%
+
+\abstract{ This study aimed to analyze prompts of programming students with a
+chatbot powered by OpenAI’s GPT-3.5, enhanced with the Retrieval Augmented
+Generation (RAG) technique, within the context of a Java programming course. The
+focus was on students using two metacognitive strategies: interleaving and
+spacing. Student prompts were categorized into eight categories along with their
 respective study topics. Findings revealed that the categories and markers of
@@
 
-\twocolumn\maketitle\normalsize\setcounter{footnote}{0}
+\keywords{Programming, Generative Artificial Intelligence, Chatbots, Spacing,
+Interleaving and Metacognition}
 
-\section{\uppercase{Introduction}}
-\label{sec:introduction}
+\maketitle
+
+\section{Introduction}\label{sec1}
 
 Recent advances in Generative Artificial Intelligence (GenAI) have created
-new opportunities in both professional programming \citep{Peng23}
-\citep{Pandey24} and education \citep{Puryear22}. In programming courses,
-students can employ GenAI tools to improve their understanding, receive
-personalized feedback, and access detailed explanations. For example, GitHub
-Copilot, a tool that is integrated into development environments, assists in
-providing real-time suggestions and accelerating code writing, which could help
-streamline the learning process \citep{Denny23}. Educational chatbots, on the
-other hand, try to guide students through coding challenges, answer questions,
-and try to foster a deeper understanding of programming concepts
-\citep{Labadze23}.
+new opportunities in both professional programming and education \citep{Puryear22}.
+In programming courses, students can employ GenAI tools to improve their
+understanding, receive personalized feedback, and access detailed explanations.
+For example, GitHub Copilot, a tool that is integrated into development
+environments, assists in providing real-time suggestions and accelerating code
+writing, which could help streamline the learning process. Educational chatbots,
+on the other hand, try to guide students through coding challenges, answer
+questions, and try to foster a deeper understanding of programming concepts.
 
@@
 achievement \citep{Callejo24}. Furthermore, it was observed that ChatGPT may
-allow students to regulate their learning pace \citep{Baha24} and can
-potentially support self-regulated learning, particularly for students with
-prior technical and disciplinary knowledge \citep{Xia23}.
+allow students to regulate their learning pace and can potentially support
+self-regulated learning, particularly for students with prior technical and
+disciplinary knowledge \citep{Xia23}.
 
@@
 
-\section{\uppercase{Theoretical Framework}}
+\section{Theoretical Framework}\label{sec1}
 
@@
 determine how to approach a task, such as spacing their study sessions
-\citep{Ouhao18, Carvalho20}, interleaving topics \citep{Rivers21}, utilizing
+\citep{Carvalho20}, interleaving topics \citep{Rivers21}, utilizing
 retrieval practice \citep{larsen18}, and other possible strategies.
@@
 learning and the application of student's knowledge in other contexts and
-situations \citep{rohrer20} \citep{Rivers21} .
+situations \citep{Rivers21} .
 
-\subsection{GenAI Related Works}
+\subsection{Related Work}
 
@@
 
-\begin{table*}[htbp]
-  \caption{Categories - adapted from \cite{Ghimire24}}
-  \begin{center}
-    \renewcommand{\arraystretch}{1.1} % Increase the spacing between rows
-    \begin{tabular}{p{3.8cm} p{6cm} p{4.5cm}} % Remove vertical bars
-      \hline
-      \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
-      \hline
-      Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
-      Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
-      Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
-      Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
-      Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
-      Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
-      Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
-      Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
-      \hline
-    \end{tabular}
-    \label{tab:categories}
-  \end{center}
-\end{table*}
-
 \subsection{Metacognition Related Works}
@@
 
-% \cite{Khusnul24} investigates
-% the efficacy of a meta-learning approach in improving metacognitive and creative
-% skills. This study confirms the effectiveness of metacognition strategies and
-% elucidates the relationship between meta-learning and metacognition. This
-% research suggests that meta-learning can improve metacognitive abilities,
-% providing valuable insights into educational technology and course design in
-% higher education settings.
+\cite{Khusnul24} investigates
+the efficacy of a meta-learning approach in improving metacognitive and creative
+skills. This study confirms the effectiveness of metacognition strategies and
+elucidates the relationship between meta-learning and metacognition. This
+research suggests that meta-learning can improve metacognitive abilities,
+providing valuable insights into educational technology and course design in
+higher education settings.
 
-% \cite{Sidra24} addressed the growing importance of metacognitive skills in an
-% AI-enabled workforce. The problem discussed was how AI systems, especially in
-% collaborative tools based on Large Language Models (LLMs), impacted human
-% decision-making and the challenges they posed due to the lack of human-like
-% metacognition in AI. The method used identified four main characteristics that
-% differentiated human-AI interaction from human-human collaboration and proposed,
-% based on the dual-process theory of human thought, that metacognitive thinking
-% was crucial for the effective use of AI. The results indicated that human
-% workers needed to develop and apply metacognitive skills, such as planning,
-% monitoring, and evaluating their thought processes, to mitigate cognitive biases
-% and improve decision-making when working with AI. The study concluded by
-% recommending improvements in AI design and training to improve workers'
-% metacognitive abilities, ensuring more effective collaboration with AI systems.
+\cite{Sidra24} addressed the growing importance of metacognitive skills in an
+AI-enabled workforce. The problem discussed was how AI systems, especially in
+collaborative tools based on Large Language Models (LLMs), impacted human
+decision-making and the challenges they posed due to the lack of human-like
+metacognition in AI. The method used identified four main characteristics that
+differentiated human-AI interaction from human-human collaboration and proposed,
+based on the dual-process theory of human thought, that metacognitive thinking
+was crucial for the effective use of AI. The results indicated that human
+workers needed to develop and apply metacognitive skills, such as planning,
+monitoring, and evaluating their thought processes, to mitigate cognitive biases
+and improve decision-making when working with AI. The study concluded by
+recommending improvements in AI design and training to improve workers'
+metacognitive abilities, ensuring more effective collaboration with AI systems.
 
@@
 
+\begin{table*}[htbp]
+    \caption{Categories - adapted from \cite{Ghimire24}}
+    \begin{center}
+      \renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
+      \begin{tabular}{p{3cm} p{4cm} p{3cm}} % Remove vertical bars
+        \hline
+        \textbf{Category} & \textbf{Description} & \textbf{Example of prompt} \\
+        \hline
+        Debugging Help (DH) & Prompts that seek help to identify, fix errors, or understand the provided code snippet. & \textit{Would this code be ok? \{code\}} \\
+        Conceptual Question (CQ) & Prompts that are more about understanding concepts than specific code. & \textit{What does it mean for a method to be static?} \\
+        Student Correction (SC) & Prompts where the student corrects the chatbot. & \textit{The correct answer is B} \\
+        Code Snippet  (CS) & Prompts that ask for a specific part of the code, like a function or a segment. & \textit{A class inherits from another write this code} \\
+        Complete Solution (CSO) & Prompts that request an entire solution or a complete code snippet. & \textit{Give me the code for a selection sort} \\
+        Multiple Question (MQ) & Prompts where the user wants to solve a multiple choice exercise (Quiz). & \textit{A heap is a data structure appropriate for: \{options\}} \\
+        Language Change (LC) & Prompts that request a change of idiom. & \textit{In Spanish} \\
+        Uncategorized (U) & Prompts that do not fit into any of the above categories. & \textit{Thanks} \\
+        \hline
+      \end{tabular}
+      \label{tab:categories}
+    \end{center}
+  \end{table*}
+
 Initially, \cite{Ghimire24} proposed four categories: Debugging Help (DH),
@@
 The analysis phase involved using Python/Pandas scripts to extract information
-from previously classified data. The data sets, and code used in this process
-are available at https://github.com/rodrigoprestesmachado/uc3m.
+from previously classified data.
 
-\section{\uppercase{Results and Discussion}}
-
-A total of 625 student messages were categorized in 81 conversations, with
-an average of 7.72 messages per conversation. This data offers insights into how
-users engage with CharlieBot and the types of prompts they submit. The results
-of the analysis are presented in this section, addressing the research questions
-outlined in the introduction. Figure \ref{fig:graph1} shows the distribution of
-the messages in the categories. The following sections present the results and
-discussions of each research question.
-
-\subsection{Categorization of messages}
-
-About the first research question: \textit{RQ1 - What distribution patterns
-emerged in the classified student interaction prompts with the educational
-chatbot?}
-
-According to Figure \ref{fig:graph1} messages classified as
-Conceptual Question account for 35.8\% of the total. Sequences of Conceptual
-Questions often start from a practical perspective of a given code, as
-illustrated in conversation 4 of Figure \ref{fig:graph2}, which presents
-examples of students' interactions with the chatbot. In other cases, the
-conversation consists entirely of messages classified as Conceptual Question,
-such as in conversation 15 of Figure \ref{fig:graph2}.
-
-\begin{figure}[h!]
-  \centering
-  \includegraphics[scale=0.61]{img/figure1.png}
-  \caption{Categorization of messages.}
-  \label{fig:graph1}
-\end{figure}
-
-\begin{figure*}[htbp]
-  \centering
-  \includegraphics[scale=0.41]{img/figure2.png}
-  \caption{Examples of conversations.}
-  \label{fig:graph2}
-\end{figure*}
-
-As shown in Figure \ref{fig:graph1}, approximately 29.9\% of the messages were
-classified as Debugging Help. Debugging help messages are likely to provide
-students with a practical approach to understanding code, identifying errors,
-and building skills to enhance their debugging abilities for future tasks.
-
-Most of the messages are located in the categories Conceptual Question and
-Debugging Help corroborate the findings of \cite{Ghimire24} which showed that
-the questions are also localized in the same categories.
-
-Multiple Question exercise resolutions represented 14.9\% of the responses.
-These prompts typically involved students submitting one or more exercises to
-CharlieBot and requesting solutions. The conversation 13 in Figure
-\ref{fig:graph2} illustrates a sequence of multiple choice questions.
-
-Requests for the chatbot to generate a Code Snippet or Complete Solution accounted
-for 8.2\% of student messages. These prompts reflected a desire for more
-direct answers; however, it was observed that, after receiving these solutions,
-many students transitioned to a more active approach in seeking to understand
-the code or the underlying concepts. Thus, it can be inferred that the prompts
-within the Code Snippet or Complete Solution category were often used as a
-starting point for a more in-depth study.
-
-Student corrections to chatbot responses represented 4.2\% of the interactions.
-In most cases, these corrections occurred when the student already had the
-correct answer to an exercise and identified an error in the chatbot response,
-as illustrated in conversation 2 in Figure \ref{fig:graph2}. These corrections
-highlight that, although the chatbot provides quick feedback, it is not always
-accurate.
-
-Figure \ref{fig:graph1} also indicates that approximately 5.8\% of the prompts
-written by the students were classified as Uncategorized. These messages include
-expressions of gratitude toward the chatbot, such as a simple '\textit{thanks}',
-contextual statements like '\textit{I'm reviewing object orientation}', as well
-as requests for additional exercises and summaries. These Uncategorized messages
-can be subdivided into new categories depending on the investigation. For
-instance, requests for new exercises or summaries could be interpreted as a
-metacognitive monitoring strategy.
-
-Finally, 1.3\% of the messages were classified as Language Change. These prompts
-were typically requests for the chatbot to switch languages (English to Spanish),
-as seen in Figure \ref{fig:graph2}, conversation 9, message 2.
-
-\subsection{Spacing}
-
-About the first part of the second research question: \textit{RQ2 - Was there
-spacing between student prompts?}
-
-For the spacing analysis, we initially considered a minimum interval of 60
-minutes as a reference to define the significant spacing \citep{Gadella24}.
-Based on this, our results showed that, of the 81 conversations analyzed, just
-27, or around 34\%, had at least one message with an interval more significant
-than 60 minutes. Figure \ref{fig:graph2} illustrates the elapsed time between
-messages (delta). For example, in conversation 15, the first message was sent
-at time zero, while the student sent the second message after a delta of 0.6
-minutes. Additionally, message 4 in conversation 15, as shown in Figure
-\ref{fig:graph2}, presents a delta of 69.1 minutes, which, according to our
-reference, characterizes spacing. The average number of subsections per
-conversation was 2.2 among conversations that exhibited spacing. Therefore, a
-conversation with one spacing is divided into two distinct study sessions.
-
-This data suggests that there is indeed little spacing between study sessions.
-To create a new section in CharlieBot; students must close the browser tab and
-start a new conversation, which results in losing the previous chat history.
-
-\subsubsection{Interaction Time per Category}
-
-About the second part of the second research question: \textit{RQ2 - Which
-category led to the most rapid and consistent interactions with the chatbot?}
-
-The objective of this research question was to study the time a student takes to
-re-engage with the bot after sending a message for each category. This analysis
-can offer insights into students' behavior: faster interactions might with the
-chatbot might suggest a more passive learning posture.
-
-To calculate the average time and standard deviation for messages within a
-category initially was first observed conversations in which message timing was
-influenced by previous messages. Pearson's correlation analysis revealed that
-28.40\% of the conversations exhibited a strong correlation, which, in this
-study, was defined as values outside the range of -0.3 to 0.3. In this context,
-a strong correlation suggests that the response time for a message may depend
-on the time of the preceding message or multiple prior messages. Consequently,
-these 28.40\% of conversations were excluded from the average time and standard
-deviation calculation, leaving 56 out of the initial 86 conversations. Table
-\ref{tab:averages} presents the means and standard deviations for each category.
-
-When a student requests a Code Snippet (CS), it is generally predictable that
-they will summarize interaction with the chatbot within a short period, as
-indicated by the mean (1.34) and standard deviation (1.95) of the Code Snippet
-category in Table \ref{tab:averages}. This result is noteworthy because,
-although some requests are simple and lead to quick responses, others may be
-more complex and require more reflection time. However, the low average response
-time and minimal variability may indicate that students adopt a more passive
-learning posture when requesting the chatbot to generate pieces of code.
-
-Despite occasional variation (3.59), prompts classified as Uncategorized exhibit
-a relatively low average time (1.70) for students to resume interaction with the
-chatbot. This variation occurs partly because of situations where students end
-the conversation with an Uncategorized message like \textit{"thanks"} but later
-decide to resume the interaction.
-
-When a student requests the chatbot to Change the Language (LC)  in a response,
-in our context English to Spanish, the new interaction occurs quickly (2.23)
-with slight variation in response time (3.71), meaning that it is predictable
-that the student will continue interacting.
-
-Although still below the overall average, students take moderate time (2.47) to
-resume interaction with the chatbot after requesting the solution to a Multiple
-Question (MQ) exercise. The variation (5.56) suggests that some responses may
-lead to a slower follow-up interaction, possibly indicating a moment of deeper
-reflection on the part of the students. However, since the mean and standard
-deviation are below the overall average, this could indicate a more passive
-behavior, as students obtain ready-made answers from the chatbot.
-
-The average time of the Student Correction (SC) category is close to the overall
-average (4.07), although the high standard deviation (8.42) indicates
-significant variation. In some cases, students know the answer to an exercise
-and only accomplish a quick correction from the chatbot; in others, they detect
-flaws in the chatbot's responses and need to examine the subsequent answer to
-identify potential errors, which suggests a more active behavior.
-
-A message classified as Debugging Help (DH) has both a long average response
-time (5.30) and high variation (10.80). Consequently, the behavior in this
-category is more time-consuming and unpredictable. A possible explanation is
-that some DH messages are simple, such as '\textit{what does x++ do?}'. In
-contrast, others involve requests for analysis of more complex code, like
-\textit{'explain this piece of code (accompanied by a complete method).'}
-
-Conceptual Questions (CQ) generally result in a high average time (5.47) for
-students to reengage with the chatbot. Moreover, the high standard deviation
-(10.46) indicates an unpredictable variation in response times. This
-unpredictability may be attributed to the fact that, in some cases, the
-complexity of the question and answer demands more reflection time from the
-student, whereas in others, lesser complexity allows for a quicker interaction
-with the chatbot.
-
-Finally, a message classified as a Complete Solution request (SCO) exhibits the
-longest average time (7.65) for students to resume interaction with the chatbot.
-Additionally, the high standard deviation (11.03) highlights significant
-unpredictability in response time. After requesting a complete solution,
-students may take longer to return due to the complexity of the problems
-involved, such as the implementation of sorting algorithms such as Heapsort.
-
-\begin{table*}[htbp]
-  \caption{Average and Standard Deviation of Different Categories (in minutes)}
-  \begin{center}
-    \renewcommand{\arraystretch}{1.1} % Adjust the spacing between rows
-    \begin{tabular}{p{4cm} p{3cm} p{3cm}} % Remove vertical bars
-      \hline
-      \textbf{Category} & \textbf{Average} & \textbf{Standard Deviation} \\
-      \hline
-      Code Snippet & 1.34 & 1.95 \\
-      Uncategorized & 1.70 & 3.59 \\
-      Language Change & 2.23 & 3.71 \\
-      Multiple Question & 2.47 & 5.66 \\
-      \hdashline
-      \textit{Overall} & \textit{3.78} & \textit{6.95} \\
-      \hdashline
-      Student Correction & 4.07 & 8.42 \\
-      Debugging Help & 5.30 & 10.80 \\
-      Conceptual Questions & 5.47 & 10.46 \\
-      Complete Solution & 7.65 & 11.03 \\
-      \hline
-    \end{tabular}
-    \label{tab:averages}
-  \end{center}
-\end{table*}
-
-\subsection{Interleaving}
-
-About the third research question: \textit{RQ3 - Do students' interactions with
-the chatbot alternate between study topics to suggest interleaving?}
-
-Switching between topics can help students identify distinct concepts inside
-problems. The results show that approximately 50.8\% of the conversations
-include interleaving; consequently, 49.2\% are related to a single study topic.
-To gain a deeper comprehension of this interleaving, conversations were analyzed
- by separating those without spacing from those that included spacing.
-
-Conversations without spacing comprise around 66\% of the total, with 34.2\%
-showing a topic change, indicating interleaving (equivalent to 22.8\% of all
-conversations). Qualitative analysis of these conversations without spacing
-revealed that about 7\% of the conversations presented interleaving caused by
-students' at least one request to solve quizzes. However, the speed with which
-students interact with the chatbot to solve quizzes may suggest a more
-superficial reflection on the answers received (see section 4.2.1). On the other
-hand, 8.8\% of the conversations showed no spacing but included topic changes,
-suggesting that students effectively might employ a metacognitive interleaving
-strategy to distinguish concepts around a problem. Finally, 7\% of the
-conversations featured interaction pauses lasting between 20 and 59 minutes but
-fell short of the study's definition of spacing, which required pauses of over
-60 minutes. This subset of conversations revealed a natural interleaving of
-topics, reflecting a chatbot usage pattern.
-
-Conversations with spacing comprise approximately 34\% of the total, with 84.2\%
-indicating interleaving (equivalent to 28\% of all conversations). Pauses longer
-than 60 minutes revealed similar behavior to those shorter than 60 minutes, with
-students using the chatbot primarily for quick consultations, asking questions,
-and returning at least one hour late with a query about a different study topic.
-The argument that some students use the chatbot primarily as a tool for quick
-queries is supported by data showing an average of 4.18 messages per study
-session. Interestingly, only 6\% of the conversations involved students pausing
-their interaction with the chatbot for more than 60 minutes before sending a
-message related to the exact topic of study.
-
-Therefore, although approximately 50.8\% of the conversations contain
-interleaving, the data shows that this topic switching is limited in terms of
-using a conscious metacognitive strategy.
-
-\section{Conclusion and Future Work}
-
-This study aimed to analyze the interactions of Java programming students with
-an educational chatbot developed using the RAG technique, employing two
-metacognitive study strategies: Spacing and Interleaving.
-
-The analysis categories, spacing, and interleaving markers provided insights
-into students' interactions with the chatbot. With the concept of spacing, it
-was possible to define students' study sessions, which, in this work, consisted
-of a few interaction messages with the chatbot. Consequently, a significant
-portion of the observed interleaving occurred due to spacing, that is, in the
-transition between study sessions. These findings indicate that the intentional
-use of learning strategies, such as spacing and interleaving, is still limited.
-Therefore, encouraging the deliberate application of these practices can enhance
-students' learning outcomes using generative artificial intelligence (GenAI)
-tools.
-
-Messages from the Code Snippets and Multiple Question (Quiz) categories stood
-out for involving consistently quick interactions with the bot. This pattern
-suggests some level of engagement with the chatbot but also raises the
-possibility of less in-depth reflection on the content received, especially when
-students request small code snippets or the resolution of exercises.
-
-It is fundamental to consider how GenAI tools impact students in various
-demographic groups, academic disciplines, cultural backgrounds, and levels of
-previous experience \citep{catalan21} \citep{neo22}. Consequently, this study is
-limited to a specific group of students and focuses on using a single GenAI
-tool. As a result, the findings may not be generalizable to other populations or
-tools.
-
-There are ways to deepen this study; based on previous references, 60 minutes
-was the standard spacing measure. However, further studies will deepen this
-analysis using a quantitative variable to better represent different spacing
-intervals. Future research can also include additional metacognitive strategies,
-which will involve interpreting user messages in depth. For instance, the
-Monitoring strategy was evident when students expressed that a concept remained
-unclear and requested further clarification. Another necessary analysis involves
-investigating the relationship between student profiles, metacognitive
-strategies for using GenAI, and learning outcomes.
-
-\bibliographystyle{apalike}
-{\small
-\bibliography{References}}
+% Ensure the bibliography style and data are correctly included
+\bibliography{sn-bibliography}
 

### 2025-01-27 CSEDU2025/Paper.tex

@@
 Recent advances in Generative Artificial Intelligence (GenAI) have created
-new opportunities in both professional programming \citep{Peng23}
-\citep{Pandey24} and education \citep{Puryear22}. In programming courses,
-students can employ GenAI tools to improve their understanding, receive
-personalized feedback, and access detailed explanations. For example, GitHub
-Copilot, a tool that is integrated into development environments, assists in
-providing real-time suggestions and accelerating code writing, which could help
-streamline the learning process \citep{Denny23}. Educational chatbots, on the
-other hand, try to guide students through coding challenges, answer questions,
-and try to foster a deeper understanding of programming concepts
-\citep{Labadze23}.
+new opportunities in both professional programming and education \citep{Puryear22}.
+In programming courses, students can employ GenAI tools to improve their
+understanding, receive personalized feedback, and access detailed explanations.
+For example, GitHub Copilot, a tool that is integrated into development
+environments, assists in providing real-time suggestions and accelerating code
+writing, which could help streamline the learning process. Educational chatbots,
+on the other hand, try to guide students through coding challenges, answer
+questions, and try to foster a deeper understanding of programming concepts.
 
@@
   \begin{center}
-    \renewcommand{\arraystretch}{1.1} % Increase the spacing between rows
+    \renewcommand{\arraystretch}{1} % Increase the spacing between rows
     \begin{tabular}{p{3.8cm} p{6cm} p{4.5cm}} % Remove vertical bars
@@
 
-\subsection{Metacognition Related Works}
+% \subsection{Metacognition Related Works}
 
-Learning based on metacognitive strategies has been shown to be effective in
-enhancing students’ performance and skills. \cite{Zheng19} investigated the
-effects of metacognitive scaffolds on group behavior, performance, and cognitive
-load in computer-supported collaborative environments. The results showed
-positive impacts on metacognitive behavior and group performance without
-increasing the cognitive load. \cite{LiWei23} proposed a metacognition-based
-collaborative approach to enhance performance in collaborative programming. The
-findings indicated that this approach influenced computational thinking,
-critical thinking, and metacognitive awareness. Similarly, \cite{Wang23}
-explored how metacognitive instruction affects computational thinking, critical
-thinking, and metacognitive skills in collaborative programming, observing
+% Learning based on metacognitive strategies has been shown to be effective in
+% enhancing students’ performance and skills. \cite{Zheng19} investigated the
+% effects of metacognitive scaffolds on group behavior, performance, and cognitive
+% load in computer-supported collaborative environments. The results showed
+% positive impacts on metacognitive behavior and group performance without
+% increasing the cognitive load. \cite{LiWei23} proposed a metacognition-based
+% collaborative approach to enhance performance in collaborative programming. The
+% findings indicated that this approach influenced computational thinking,
+% critical thinking, and metacognitive awareness. Similarly, \cite{Wang23}
+% explored how metacognitive instruction affects computational thinking, critical
+% thinking, and metacognitive skills in collaborative programming, observing
 
@@
 
-\cite{Sidra24} explored the importance of metacognitive skills in an AI-enabled
-workforce, focusing on how AI, particularly tools using Large Language Models,
-affects human decision-making. They highlighted challenges arising from AI's
-lack of human-like metacognition and identified key differences between human
-AI and human-human collaboration. Using dual-process theory, they emphasized the
-need for workers to develop metacognitive skills like planning, monitoring, and
-evaluating to counter cognitive biases and enhance decision-making. The study
-recommended improving AI design and training to boost workers' metacognitive
-abilities for more effective collaboration.
+% \cite{Sidra24} explored the importance of metacognitive skills in an AI-enabled
+% workforce, focusing on how AI, particularly tools using Large Language Models,
+% affects human decision-making. They highlighted challenges arising from AI's
+% lack of human-like metacognition and identified key differences between human
+% AI and human-human collaboration. Using dual-process theory, they emphasized the
+% need for workers to develop metacognitive skills like planning, monitoring, and
+% evaluating to counter cognitive biases and enhance decision-making. The study
+% recommended improving AI design and training to boost workers' metacognitive
+% abilities for more effective collaboration.
 
@@
 The analysis phase involved using Python/Pandas scripts to extract information
-from previously classified data. The data sets, and code used in this process
-are available at https://github.com/rodrigoprestesmachado/uc3m.
+from previously classified data.
 
@@
   \centering
-  \includegraphics[scale=0.61]{img/figure1.png}
+  \includegraphics[scale=0.6]{img/figure1.png}
   \caption{Categorization of messages.}
@@
   \centering
-  \includegraphics[scale=0.41]{img/figure2.png}
+  \includegraphics[scale=0.39]{img/figure2.png}
   \caption{Examples of conversations.}
@@
   \begin{center}
-    \renewcommand{\arraystretch}{1.1} % Adjust the spacing between rows
+    \renewcommand{\arraystretch}{1.0} % Adjust the spacing between rows
     \begin{tabular}{p{4cm} p{3cm} p{3cm}} % Remove vertical bars
@@
 
-There are ways to deepen this study; based on previous references, 60 minutes
-was the standard spacing measure. However, further studies will deepen this
-analysis using a quantitative variable to better represent different spacing
-intervals. Future research can also include additional metacognitive strategies,
-which will involve interpreting user messages in depth. For instance, the
-Monitoring strategy was evident when students expressed that a concept remained
-unclear and requested further clarification. Another necessary analysis involves
-investigating the relationship between student profiles, metacognitive
-strategies for using GenAI, and learning outcomes.
+There are ways to deepen this study. While 60 minutes was the standard spacing
+measure, future research will explore different intervals using quantitative
+variables. Additional metacognitive strategies, such as the Monitoring
+strategy—where students seek clarification on unclear concepts—will also be
+examined. Moreover, investigating the relationship between student profiles,
+GenAI metacognitive strategies, and learning outcomes is essential.
 

### 2025-01-27 ETHE/sn-article.tex
ETHE update

@@
 
-\author*[1,2]{\fnm{Rodrigo} \sur{Prestes Machado}}\email{rodrigo.prestes@poa.ifrs.edu.br}
+\author*[1]{\fnm{Rodrigo} \sur{Prestes Machado}}\email{rodrigo.prestes@poa.ifrs.edu.br}
+\author*[2]{\fnm{Carlos} \sur{Alario-Hoyos}}\email{calario@it.uc3m.es}
+\author[2]{\fnm{Iria} \sur{Estévez-Ayres}}\email{ayres@it.uc3m.es}
+\author[2]{\fnm{Patricia} \sur{Callejo}}\email{pcallejo@it.uc3m.es}
+\author[2]{\fnm{Carlos} \sur{Delgado Kloos}}\email{cdk@it.uc3m.es}
 
-\author[2,3]{\fnm{Second} \sur{Author}}\email{iiauthor@gmail.com}
 \equalcont{These authors contributed equally to this work.}
 
-\author[1,2]{\fnm{Third} \sur{Author}}\email{iiiauthor@gmail.com}
-\equalcont{These authors contributed equally to this work.}
+\affil*[1]{\orgdiv{Informatics}, \orgname{Federal Institute of Education,
+Science and Technology}, \orgaddress{\street{Cel. Vicente}, \city{Porto Alegre},
+\postcode{90.030-041}, \state{Rio Grande do Sul}, \country{Brazil}}}
 
-\affil*[1]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{100190}, \state{State}, \country{Country}}}
-
-\affil[2]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{10587}, \state{State}, \country{Country}}}
-
-\affil[3]{\orgdiv{Department}, \orgname{Organization}, \orgaddress{\street{Street}, \city{City}, \postcode{610101}, \state{State}, \country{Country}}}
+\affil[2]{\orgdiv{Department of Telematics Engineering},
+\orgname{Universidad Carlos III}, \orgaddress{ \street{Av. de la Universidad},
+\city{Leganés}, \postcode{28911}, \state{Madrid}, \country{Spain}}}
 

### 2025-01-30 ETHE/sn-article.tex
update paper and Java Labs

@@
 
-Recent advances in Generative Artificial Intelligence (GenAI) have created
-new opportunities in both professional programming and education \citep{Puryear22}.
-In programming courses, students can employ GenAI tools to improve their
+Recent advances in Generative Artificial Intelligence (GenAI) have created new
+opportunities in professional programming and education \citep{Puryear22}. In
+programming courses, students can employ GenAI tools to improve their
 understanding, receive personalized feedback, and access detailed explanations.
-For example, GitHub Copilot, a tool that is integrated into development
-environments, assists in providing real-time suggestions and accelerating code
-writing, which could help streamline the learning process. Educational chatbots,
-on the other hand, try to guide students through coding challenges, answer
-questions, and try to foster a deeper understanding of programming concepts.
+For example, GitHub Copilot, a tool integrated into development environments,
+assists in providing real-time suggestions and accelerating code writing, which
+could help streamline the learning process. On the other hand, educational
+chatbots try to guide students through coding challenges, answer questions, and
+foster a deeper understanding of programming concepts.
 
-The study conducted by \cite{chan23} revealed that both undergraduate and
-postgraduate students have positive attitudes towards the use of GenAI.
-A systematic review organized by \cite{Lo24} demonstrated that students can
-effectively learn from ChatGPT, resulting in improved comprehension and academic
-achievement \citep{Callejo24}. Furthermore, it was observed that ChatGPT may
-allow students to regulate their learning pace and can potentially support
-self-regulated learning, particularly for students with prior technical and
-disciplinary knowledge \citep{Xia23}.
+The study by \cite{chan23} revealed that undergraduate and postgraduate students
+have positive attitudes towards using GenAI. A systematic review organized by
+\cite{Lo24} demonstrated that students could effectively learn from ChatGPT,
+resulting in improved comprehension and academic achievement \citep{Callejo24}.
+Additionally, it was observed that ChatGPT can enable students to control their
+learning pace and may support self-regulated learning, especially for those with
+prior technical and disciplinary knowledge \citep{Xia23}.
 
@@
 
-\subsection{Method}
-
-The study comprised three phases: data collection, categorization, and analysis.
-During the data collection phase, students enrolled in a second-semester Java
-course at the University Carlos III of Madrid (UC3M) were introduced to
-CharlieBot and allowed to use it without following any prescribed educational
-methodology. All data were collected anonymously to ensure that interactions
-could not be traced back to individual students or linked to their academic
-performance.
-
-During the categorization phase, the students' prompts were initially classified
-into eight distinct categories using the Claude.ai \citep{claude} tool.
-Subsequently, the authors of this study manually reviewed these classifications
-to ensure accuracy and alignment with the research objectives.
-
 \begin{table*}[htbp]
@@
     \renewcommand{\arraystretch}{1.2} % Increase the spacing between rows
-      \begin{tabular}{p{3cm} p{4cm} p{3cm}} % Remove vertical bars
+    \begin{tabular}{p{4.1cm} p{4cm} p{3.7cm}} % Remove vertical bars
       \hline
@@
 
+\subsection{Method}
+
+The study comprised three phases: data collection, categorization, and analysis.
+During the data collection phase, students enrolled in a second-semester Java
+course at the University Carlos III of Madrid (UC3M) were introduced to
+CharlieBot and allowed to use it without following any prescribed educational
+methodology. All data were collected anonymously to ensure that interactions
+could not be traced back to individual students or linked to their academic
+performance.
+
+During the categorization phase, the students' prompts were initially classified
+into eight distinct categories using the Claude.ai \citep{claude} tool.
+Subsequently, the authors of this study manually reviewed these classifications
+to ensure accuracy and alignment with the research objectives.
 Initially, \cite{Ghimire24} proposed four categories: Debugging Help (DH),

### 2025-01-30 JavaLab/Recap2/Circle.java

@@ -1,0 +1,63 @@
+package edu.uc3m;
+
+import java.math.BigDecimal;
+import java.math.MathContext;
+
+/**
+ * The Circle class represents a circle with a specified precision for area
+ * calculation.
+ */
+public class Circle {
+
+    private int precision;
+
+    /**
+     * Constructs a Circle with the specified precision.
+     *
+     * @param precision the number of decimal places to round the area to
+     */
+    public Circle(int precision) {
+        this.precision = precision;
+    }
+
+    /**
+     * Calculates the area of the circle with the given radius and rounds it to
+     * the specified precision.
+     *
+     * @param radius the radius of the circle
+     * @return the area of the circle rounded to the specified precision
+     */
+    public double area(double radius) {
+        double area = Math.PI * (radius * radius);
+        BigDecimal roundedArea = new BigDecimal(area)
+                .round(new MathContext(this.precision));
+        return roundedArea.doubleValue();
+    }
+
+    /**
+     * The main method to run the Circle program. It expects command-line
+     * arguments for precision and radii.
+     *
+     * @param args the command-line arguments where the first argument is the
+     *             precision and the subsequent arguments are the radii
+     */
+    public static void main(String[] args) {
+
+        int precision = Integer.parseInt(args[0]);
+        if (args.length == 0) {
+            System.err.println(
+                    "Command­line arguments expected: precision and radii (1..n).");
+        } else if (precision < 0) {
+            System.err.println("The precision must be positive.");
+        } else {
+            Circle circle = new Circle(precision);
+
+            for (int i = 1; i < args.length; i++) {
+                double radius = Double.parseDouble(args[i]);
+                System.out.println(
+                        "Area of circle with radius " + radius + " is " +
+                                circle.area(radius));
+            }
+        }
+    }
+}
